---
title: "Chapter 7 - LME analysis of Mode and Nuclear Contours: l_t"
author: "Antoin Rodgers"
date: "5.6.2022"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE)
```

```{r message=FALSE, warning=FALSE, include=FALSE}

# Load Functions.
source("../../functions/myFunctions.R") 

# Load packages.
installMissingPackages(
  c(
    # Include statistical packages.
    "performance",
    "lmerTest",
    "lme4",
    "optimx",
    "MuMIn",
    # Include packages for tidy output.
    "tidyverse",
    "broomExtra",
    "sjPlot",
    "formattable",
    "knitr",
    "RColorBrewer",
    # Include packages for %in% %notin% syntactic notation
    "mefa4"
  )
)

# Set themes and colour schemes.
theme_set(theme_minimal(base_size = 14))

mode_colours <- c("WHQ" = "#a6dba0",
                  "MDC" = "#7b32942",
                  "MYN" = "#008837",
                  "MDQ" = "#c2a5cf")

pitch_accent_colours <- c("H*"     = brewer.pal(6, "Spectral")[4],
                          "L*H"    = brewer.pal(6, "Spectral")[6],
                          ">H*"    = brewer.pal(6, "Spectral")[5],
                          "^[L*]H" = brewer.pal(6, "Spectral")[3],
                          "L*^[H]" = brewer.pal(6, "Spectral")[2],
                          "^[L*H]" = brewer.pal(6, "Spectral")[1])

p_color <- all_models_tidy ~ style(color =
                                     if_else(as.double(all_models_tidy) < 0.05,
                                             "green",
                                             "red")
                                   )


```

```{r}
# Load Corpus
m_corpus <- get_m_corpus("../data/m_corpus.csv")  %>%
  # Consolidate nuc_contour based on observation and types of register shift.
  mutate(
    nuc_contour =
      if_else(
        nuc_contour == "^[L*H %]" | nuc_contour == "^[L*H L%]",
        "^[L*H] %",
        if_else(nuc_contour == "L*^[H L%]", "L*^[H] %", nuc_contour)
      ),
    acc_phon =
      factor(
        str_replace(nuc_contour,
                    "\\s%$|\\sL%$",
                    ""),
        levels = c("H*", "L*H", ">H*", "^[L*]H", "L*^[H]", "^[L*H]")
      ),
    fin_phon = str_replace_all(fin_phon, "L%]|%]", "%")
  ) %>%
  # Select only columns for analysis
  select(
    speaker,
    gender,
    mode,
    prompt,
    phr_phon,
    fin_phon,
    l_t
  )

```



```{r}
# set current equation
cur_equation <-
  "l_t ~ mode + fin_phon + gender + (1 + mode | speaker) + (1 | prompt)"

```

### LME model analysis using initial model.

Initial equation: ``r cur_equation``

Prompt is used as an intercepts only random effect since the model does not converge when intercept and slope effect are both included.

NB: Gender caused singularity issues for another f0 parameter test (l_f0). However, it is not excluded here, since time is not a characteristic of speaker.

```{r fig.height=3, fig.width=9, message = FALSE, tidy = TRUE}
temp_mdl = lmer(
  formula(cur_equation),
  data = m_corpus,
  control = lmerControl(
    optimizer = "optimx",
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)
model_after_step <- step(temp_mdl)
```


```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
print(summary(temp_mdl))

drawResiduals(temp_mdl)

anova(temp_mdl) %>% print()

cat("Check_singularity(temp_mdl, tolerance = 1e-05 -->", check_singularity(temp_mdl, tolerance = 1e-05, "\n"))
cat("\n")
print(model_after_step)

```

* Model found from step regression: 
    * `l_t ~ mode + gender + (1 + mode | speaker)`
* Prompt is non-significant.
    * NB: prompt retained as it is significant for h_t.
* fin_phon is non-significant.
    * NB: fin_phon retained as it is significant for the three other tonal target parameters (h_t, l_f0, h_f0)
* No singularity issues.
* Extreme outlier in higher residuals.


### LME model analysis using reduced model

```{r}
# set current equation
cur_equation <-
  "l_t ~ mode + gender + fin_phon + (1 + mode | speaker) + (1 | prompt) "

```

```{r fig.height=3, fig.width=9, message = FALSE, tidy = TRUE}
temp_mdl = lmer(
  formula(cur_equation),
  data = m_corpus,
  control = lmerControl(
    optimizer = "optimx",
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

```


```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
print(summary(temp_mdl))

drawResiduals(temp_mdl)

anova(temp_mdl) %>% print()

cat("Check_singularity(temp_mdl, tolerance = 1e-05 -->", check_singularity(temp_mdl, tolerance = 1e-05, "\n"))

```

* Outlier remains.

### LME model analysis using reduced model and removing outliers (|residual|>3.5)

Running new model again but trimming rows with |residual| > 3.5 (as per Baayen 2008: 279-282).

NB: Value of 3.5 set as it removes the minimal number of outliers (n=1)

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
m_corpus_trimmed <- m_corpus %>% filter(abs(scale(resid(temp_mdl))) <= 3.5)

temp_mdl.trimmed = lmer(
  formula(cur_equation),
  data = m_corpus_trimmed,
  control = lmerControl(
    optimizer = "optimx",
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)
```


```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
print(summary(temp_mdl.trimmed))

drawResiduals(temp_mdl.trimmed)

anova(temp_mdl.trimmed) %>% print()

cat("Check_singularity(temp_mdl.trimmed, tolerance = 1e-05 -->", check_singularity(temp_mdl.trimmed, tolerance = 1e-05, "\n"))
cat("\n")


removed <- m_corpus %>% filter(abs(scale(resid(temp_mdl))) > 3.5) 
removed %>% 
  select(c(speaker, gender, mode, prompt, phr_phon, l_t)) %>% 
  mutate(phr_phon = str_replace_all(phr_phon, "\\^", "\\\\^"),
         phr_phon = str_replace_all(phr_phon, "\\*", "\\\\*")) %>% 
  kable(caption = "Data trimmed from model")

```

* Much better distribution of residuals.

### Final model: ``r cur_equation``
#### Analysis of model

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
tidy(temp_mdl.trimmed) %>%
  filter(term != "sd__(Intercept)") %>%
  bonferroniAdjust("", 6) %>%
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = round(p.value, 4),
    p.adjusted = round(p.adjusted, 4)
  ) %>%
  filter(effect %notin% "ran_pars") %>%
  select(-c(group, effect)) %>%
  rename(t.value = statistic) %>%
  formattable(caption = "Intercept and lh_slopes of fixed effects",
              list(p.adjusted = formatter("span", style = p_color))) %>%
  mutate(
    p.value = if_else(
      p.value < 0.001,
      as.character(scientific(p.value)),
      as.character(round(p.value, 4))
    ),
    p.adjusted = if_else(
      p.adjusted < 0.001,
      as.character(scientific(p.adjusted)),
      as.character(round(p.adjusted, 4))
    )
  )

  kable(
    r2_nakagawa(temp_mdl),
    caption = "Conditional and marginal R^2^ of model",
    digits = 2,
    align = "l"
  )

do.call(rbind, r2_nakagawa(temp_mdl))[,1] %>% 
  write.csv("../output/LME_l_t_r2.csv")

print(
  plot_model(
    temp_mdl,
    title = cur_equation,
    show.intercept = FALSE,
    show.values = TRUE,
    vline.color = "red",
    colors = "Black"
  )
)
```


```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
# Get number of levels in target treatment variable.
cur_levels <- levels(m_corpus_trimmed$mode)
orig_levels <- cur_levels
num_levels <- length(cur_levels)
all_models_tidy = tibble()

for (i in 1:(num_levels)) {
  # Test GLMM model on data.
  cur_model <- lmer(
    formula(cur_equation),
    data = m_corpus_trimmed,
    control = lmerControl(
      optimizer = "optimx",
      calc.derivs = FALSE,
      optCtrl = list(
        method = "nlminb",
        starttests = FALSE,
        kkt = FALSE
      )
    )
  )
  # Convert the model output into a more readable format.
  cur_model_tidy <- tidy(cur_model) %>%
    filter(effect != "ran_pars") %>%
    select(-group) %>%
    bonferroniAdjust("", 6) %>%
    mutate(
      estimate = round(estimate, 3),
      std.error = round(std.error, 3),
      statistic = round(statistic, 3),
      p.value = round(p.value, 8),
      p.adjusted = round(p.adjusted, 8)
    )
  
  # Prepare current model for pasting to all models output.
  # Make 'pairwise' column = intercept.
  cur_model_tidy <- cur_model_tidy %>%
    mutate(
      pairwise =
        if_else(
          term == "(Intercept)",
          "intercept",
          if_else(
            term %notin% c("modeMDC",
                           "modeMWH",
                           "modeMYN",
                           "modeMDQ"),
            "N/A",
            paste("mode", cur_levels[1], sep = "") #???
          )
        ),
      # change 'term' so "intercept" states the target condition name.
      term =
        if_else(
          term == "(Intercept)",
          paste("mode", cur_levels[1], sep = ""),
          term
        )
    ) %>%
    # remove random effects detail
    filter(term != "sd__(Intercept)") %>%
    select(-effect) %>%
    # rename statistic column to "t.value"
    rename(t.value = statistic)
  
  
  # make list of pairwise comparisons to keeps
  keep = NULL
  for (j in i:4) {
    keep <- c(keep, paste("mode", orig_levels[j], sep = ""))
  }
  
  # remove pairwise comparisons which have already been done
  cur_model_tidy <- filter(cur_model_tidy, term %in% keep)
  
  # add remaining pairwise comparisons to main tibble.
  all_models_tidy <- bind_rows(all_models_tidy, cur_model_tidy)
  
  # restructure the order of levels for next LME cur_model.
  cur_levels <- c(cur_levels[2:num_levels], cur_levels[1])
  m_corpus_trimmed <- m_corpus_trimmed %>%
    mutate(mode = factor(mode,
                         levels = cur_levels))
}
```

#### Summary of intercepts 

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
# arrange tibble according to pairwise move it to first column position
all_models_tidy  <- arrange(all_models_tidy, pairwise) %>%
  relocate(pairwise) %>%
  sigCodesTidy()

# print b0 table
filter(all_models_tidy, pairwise == "intercept") %>%
  select(-pairwise) %>%
  rename(intercept = term) %>%
  sigCodesTidy() %>% mutate(
    p.value = if_else(
      p.value < 0.001,
      as.character(scientific(p.value)),
      as.character(round(p.value, 4))
    ),
    p.adjusted = if_else(
      p.adjusted < 0.001,
      as.character(scientific(p.adjusted)),
      as.character(round(p.adjusted, 4))
    )
  ) %>% 
  write_csv("../output/LME_l_t_b0.csv")  %>%
  formattable(list(p.adjusted = formatter("span", style = p_color)),
              caption =paste("b0 for", cur_equation, sep = " "))
```

#### Summary of pairwise comparisons

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
# print pairwise comparisons of (B1 table)
filter(all_models_tidy, pairwise %notin% c("intercept", "N/A")) %>%
  select(pairwise,
         term,
         estimate,
         std.error,
         t.value,
         p.value,
         p.adjusted) %>%
  rename(intercept = pairwise, lh_slope = term) %>%
  sigCodesTidy() %>% mutate(
    p.value = if_else(
      p.value < 0.001,
      as.character(scientific(p.value)),
      as.character(round(p.value, 4))
    ),
    p.adjusted = if_else(
      p.adjusted < 0.001,
      as.character(scientific(p.adjusted)),
      as.character(round(p.adjusted, 4))
    )
  )  %>%
  write_csv("../output/LME_l_t_b1.csv") %>%
  formattable(list(p.adjusted = formatter("span", style = p_color)),
              caption = paste("b1 for", cur_equation, sep = " "))
```





