---
title: "Chapter 7 - LME analysis of h_t ~ Mode and Nuclear pitch accents: "
author: "Antoin Rodgers"
date: "5.6.2022"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE)

# Load Functions.
source("../../functions/myFunctions.R") 

# Load packages.
installMissingPackages(
  c(
    # Include statistical packages.
    "performance",
    "lmerTest",
    "lme4",
    "optimx",
    "MuMIn",
    # Include packages for tidy output.
    "tidyverse",
    "broomExtra",
    "sjPlot",
    "formattable",
    "knitr",
    "RColorBrewer",
    # Include packages for %in% %notin% syntactic notation
    "mefa4"
  )
)

# Set themes and colour schemes.
theme_set(theme_minimal(base_size = 14))

mode_colours <- c("WHQ" = "#a6dba0",
                  "MDC" = "#7b32942",
                  "MYN" = "#008837",
                  "MDQ" = "#c2a5cf")

pitch_accent_colours <- c("H*"     = brewer.pal(6, "Spectral")[4],
                          "L*H"    = brewer.pal(6, "Spectral")[6],
                          ">H*"    = brewer.pal(6, "Spectral")[5],
                          "^[L*]H" = brewer.pal(6, "Spectral")[3],
                          "L*^[H]" = brewer.pal(6, "Spectral")[2],
                          "^[L*H]" = brewer.pal(6, "Spectral")[1])

p_color <- all_models_tidy ~ style(color =
                                     if_else(as.double(all_models_tidy) < 0.05,
                                             "green",
                                             "red")
                                   )

# Load Corpus
m_corpus <- get_m_corpus("../data/m_corpus.csv")  %>%
  # Consolidate nuc_contour based on observation and types of register shift.
  mutate(
    nuc_contour =
      if_else(
        nuc_contour == "^[L*H %]" | nuc_contour == "^[L*H L%]",
        "^[L*H] %",
        if_else(nuc_contour == "L*^[H L%]", "L*^[H] %", nuc_contour)
      ),
    acc_phon =
      factor(
        str_replace(nuc_contour,
                    "\\s%$|\\sL%$",
                    ""),
        levels = c("H*", "L*H", ">H*", "^[L*]H", "L*^[H]", "^[L*H]")
      ),
    fin_phon = str_replace_all(fin_phon, "L%]|%]", "%")
  ) %>%
  # Select only columns for analysis
  select(
    speaker,
    gender,
    mode,
    prompt,
    acc_phon,
    fin_phon,
    h_t
  )

```

The aims of this set of analyses is to see if, once we account for assumed speaker-motivated use of pitch accent, the apparent "paralinguistic" differences in scaling (and possibly time) of phonetic parametics disappear disappear.


```{r set equation,  fig.height=3, fig.width=9, message = FALSE, tidy = TRUE}
cur_equation <-
  "h_t ~ mode + acc_phon + fin_phon + gender + (1 | speaker) + (1 | prompt)"
```

### LME model analysis using model: `r cur_equation`

```{r check model,  fig.height=3, fig.width=9, message = FALSE, tidy = TRUE}
temp_mdl <- my_lme(cur_equation, m_corpus, show_results=TRUE, run_step=TRUE)


```

-   For timing, step() suggests Mode is not a significant factor. However, I can assume it will be a sig. factor for scaling, so it will be retained for the final analysis. 

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
printTidyModel(
  temp_mdl,
  bf_adj = 7,
  exclude_terms = "",
  write_r2 = "../output/LME_PHonetic_PA_h_t_r2.csv"
)

```

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
# Get number of levels in target treatment variable.
cur_levels <- levels(m_corpus$mode)
orig_levels <- cur_levels
num_levels <- length(cur_levels)
all_models_tidy = tibble()

for (i in 1:(num_levels)) {
  cur_model <- lmer(
    formula(cur_equation),
    data = m_corpus,
    control = lmerControl(
      optimizer = "optimx",
      calc.derivs = FALSE,
      optCtrl = list(
        method = "nlminb",
        starttests = FALSE,
        kkt = FALSE
      )
    )
  )
  
  # Convert the model output into a more readable format.
  cur_model_tidy <- tidy(cur_model) %>%
    filter(term != "sd__(Intercept)") %>%
  bonferroniAdjust(7, exclude_terms = c("GenderM", "fin_phonL%")) %>%
  rename(`p.adj. (bf=7)` = p.adj) %>% 
    mutate(
      estimate = round(estimate, 3),
      std.error = round(std.error, 3),
      statistic = round(statistic, 3),
      df = round(df, 2)
      
      
    ) %>%
    filter(effect %notin% "ran_pars") %>%
    select(-c(group, effect)) %>%
    rename(t.value = statistic)

  # Get CIs of model using Wald method (fast, fixed effects only)
  ci_Wald <- confint(cur_model, method = "Wald") %>%
    as_tibble() %>%
    filter(`2.5 %` != "NA") %>%
    mutate(`2.5 %` = round(`2.5 %`, 3),
           `97.5 %` = round(`97.5 %`, 3)) %>%
    rename("2.5% CI" = "2.5 %",
           "97.5% CI" = "97.5 %")
  
  # Bind tidy model to...
  cur_model_tidy <- cbind(cur_model_tidy, ci_Wald) %>%
    # re-order columns
    select(
      "term",
      "estimate",
      "std.error",
      "2.5% CI",
      "97.5% CI",
      "t.value",
      "df",
      "p.value",
      `p.adj. (bf=7)`
    ) %>%
  # Prepare current model for pasting to all models output.
  # Make 'pairwise' column = intercept.
    mutate(
      pairwise =
        if_else(
          term == "(Intercept)",
          "intercept",
          if_else(
            term %notin% c("modeMDC",
                           "modeMWH",
                           "modeMYN",
                           "modeMDQ"),
            "N/A",
            paste("mode", cur_levels[1], sep = "") #???
          )
        ),
      # change 'term' so "intercept" states the target condition name.
      term =
        if_else(
          term == "(Intercept)",
          paste("mode", cur_levels[1], sep = ""),
          term
        )
    ) %>%
    # remove random effects detail
    filter(term != "sd__(Intercept)")
  
  
  # make list of pairwise comparisons to keeps
  keep = NULL
  for (j in i:4) {
    keep <- c(keep, paste("mode", orig_levels[j], sep = ""))
  }
  
  # remove pairwise comparisons which have already been done
  cur_model_tidy <- filter(cur_model_tidy, term %in% keep)
  
  # add remaining pairwise comparisons to main tibble.
  all_models_tidy <- bind_rows(all_models_tidy, cur_model_tidy)
  
  # restructure the order of levels for next LME cur_model.
  cur_levels <- c(cur_levels[2:num_levels], cur_levels[1])
  m_corpus <- m_corpus %>%
    mutate(mode = factor(mode,
                         levels = cur_levels))
}
```

#### Summary of intercepts

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
# arrange tibble according to pairwise move it to first column position
all_models_tidy  <- arrange(all_models_tidy, pairwise) %>%
  relocate(pairwise) %>%
  sigCodesTidy(`p.adj. (bf=7)`)

# print b0 table
filter(all_models_tidy, pairwise == "intercept") %>%
  select(-pairwise) %>%
  rename(intercept = term) %>%
  sigCodesTidy(`p.adj. (bf=7)`) %>% mutate(
    p.value = if_else(
      p.value < 0.001,
      as.character(scientific(p.value), digits=2),
      as.character(round(p.value, 4), digits=2)
    ),
    `p.adj. (bf=7)` = if_else(
      `p.adj. (bf=7)` < 0.001,
      as.character(scientific(`p.adj. (bf=7)`, digits=2)),
      as.character(round(`p.adj. (bf=7)`, 4))
    )
  ) %>% 
  write_csv("../output/LME_PHonetic_PA_h_t_b0.csv")  %>%
  formattable(list(`p.adj. (bf=7)` = formatter("span", style = p_color)),
              caption =paste("b0 for", cur_equation, sep = " "))
```

#### Summary of pairwise comparisons

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
# print pairwise comparisons of (B1 table)
filter(all_models_tidy, pairwise %notin% c("intercept", "N/A")) %>%
  select(
    pairwise,
    term,
    estimate,
    std.error,
    `2.5% CI`,
    `97.5% CI`,
    t.value,
    df,
    p.value,
    `p.adj. (bf=7)`
  ) %>%
  rename(intercept = pairwise, slope = term) %>%
  sigCodesTidy(`p.adj. (bf=7)`) %>% mutate(
    p.value = if_else(
      p.value < 0.001,
      as.character(scientific(p.value), digits=2),
      as.character(round(p.value, 4), digits=2)
    ),
    `p.adj. (bf=7)` = if_else(
      `p.adj. (bf=7)` < 0.001,
      as.character(scientific(`p.adj. (bf=7)`, digits=2)),
      as.character(round(`p.adj. (bf=7)`, 4))
    )
  )  %>%
  rename(`p.adj. (bf=7)` = `p.adj. (bf=7)`) %>%
  write_csv("../output/LME_PHonetic_PA_h_t_b1.csv") %>%
  formattable(list(`p.adj. (bf=7)` = formatter("span", style = p_color)),
              caption = paste("b1 for", cur_equation, sep = " "))
```
