---
title: "Chapter 7 - LME analysis of Nuclear Contours: l_f0"
author: "Antoin Rodgers"
date: "5.6.2022"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE)
```

```{r load packages and functions,  message=FALSE, warning=FALSE, include=FALSE}

# Load Functions.
source("../../functions/myFunctions.R") 

# Load packages.
installMissingPackages(
  c(
    # Include statistical packages.
    "performance",
    "lmerTest",
    "lme4",
    "optimx",
    "MuMIn",
    # Include packages for tidy output.
    "tidyverse",
    "broomExtra",
    "sjPlot",
    "formattable",
    "knitr",
    "RColorBrewer",
    # Include packages for %in% %notin% syntactic notation
    "mefa4"
  )
)

# Set themes and colour schemes.
theme_set(theme_minimal(base_size = 14))

p_color <- all_models_tidy ~ style(color =
                                     if_else(as.double(all_models_tidy) < 0.05,
                                             "green",
                                             "red")
                                   )


```

```{r load corpus}
# Load Corpus
l_star_h_types = c("L*H", "^[L*]H", "L*^[H]", "^[L*H]")

m_corpus <- get_m_corpus("../data/m_corpus.csv")  %>%
  # Consolidate nuc_contour based on observation and types of register shift.
  mutate(
    nuc_contour =
      if_else(
        nuc_contour == "^[L*H %]" | nuc_contour == "^[L*H L%]",
        "^[L*H] %",
        if_else(nuc_contour == "L*^[H L%]", "L*^[H] %", nuc_contour)
      ),
    acc_phon =
      factor(
        str_replace(nuc_contour,
                    "\\s%$|\\sL%$",
                    ""),
        levels = l_star_h_types
      ),
    fin_phon = str_replace_all(fin_phon, "L%]|%]", "%")
  ) %>%
  # Select only columns for analysis
  select(
    speaker,
    gender,
    mode,
    prompt,
    foot_syls,
    acc_phon,
    phr_phon,
    fin_phon,
    l_f0,
    l_f0_z
  ) %>% 
  filter(acc_phon %in% l_star_h_types)  %>% 
  mutate(acc_phon = factor(acc_phon,
                           levels = l_star_h_types)
         )

```

```{r set initial equation}
# set current equation
cur_equation <- "l_f0 ~ acc_phon + fin_phon + gender + (1 + acc_phon | speaker)"
```

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
temp_mdl = lmer(
  formula(cur_equation),
  data = m_corpus,
  control = lmerControl(
    optimizer = "optimx",
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)
```

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
print(summary(temp_mdl))

drawResiduals(temp_mdl)

anova(temp_mdl) %>% print()

cat("Check_singularity(temp_mdl, tolerance = 1e-05 -->", check_singularity(temp_mdl, tolerance = 1e-05, "\n"))
cat("\n")

step(temp_mdl)
```

-   Distribution of residuals is leptokurtic.

### LME model analysis using reduced model and removing outliers (\|residual\|\>3.1)

Running new model again but trimming rows with \|residual\| \> 3.1 (as per Baayen 2008: 279-282).

NB a value of 3.1 was chosen as this removes the minimum number of outliers without leading to singularity errors while at the same time providing a more reasonably distribution or residuals.

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
m_corpus.trimmed <- m_corpus %>% filter(abs(scale(resid(temp_mdl))) < 3.1)

temp_mdl.trimmed = lmer(
  formula(cur_equation),
  data = m_corpus.trimmed,
  control = lmerControl(
    optimizer = "optimx",
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)
```

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
print(summary(temp_mdl.trimmed))

drawResiduals(temp_mdl.trimmed)

anova(temp_mdl.trimmed) %>% print()

cat("Check_singularity(temp_mdl.trimmed, tolerance = 1e-05 -->", check_singularity(temp_mdl.trimmed, tolerance = 1e-05, "\n"))
cat("\n")


removed <- m_corpus %>% filter(abs(scale(resid(temp_mdl))) >= 3.1) 
removed %>% 
  select(c(speaker, gender, mode, prompt, acc_phon, fin_phon, l_f0, l_f0_z)) %>% 
  mutate(acc_phon = str_replace_all(acc_phon, "\\^", "\\\\^"),
         acc_phon = str_replace_all(acc_phon, "\\*", "\\\\*")) %>% 
  kable(caption = "Data trimmed from model")

```

-   Plots indicate better fit.

### Final model: `r cur_equation`

#### Analysis of model

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
tidy_model <- tidy(temp_mdl) %>%
  filter(term != "sd__(Intercept)") %>%
  bonferroniAdjust(8, exclude_terms = c("fin_phonL%", "genderM")) %>%
  rename(`p.val. adj. (bf=8)` = p.adj) %>% 
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    df = round(df, 2)
    
    
  ) %>%
  filter(effect %notin% "ran_pars") %>%
  select(-c(group, effect)) %>%
  rename(t.value = statistic) %>% 
    mutate(term = str_replace(term, "\\^\\[", "\\\\^\\\\["))

# Get CIs of model using Wald method (fast, fixed effects only)
ci_Wald <- confint(temp_mdl, method = "Wald") %>%
  as_tibble() %>%
  filter(`2.5 %` != "NA") %>%
  mutate(`2.5 %` = round(`2.5 %`, 3),
         `97.5 %` = round(`97.5 %`, 3)) %>%
  rename("2.5% CI" = "2.5 %",
         "97.5% CI" = "97.5 %")

# Bind tidy model to...
cbind(tidy_model, ci_Wald) %>%
  # re-order columns
  select(
    "term",
    "estimate",
    "std.error",
    "2.5% CI",
    "97.5% CI",
    "t.value",
    "df",
    "p.value",
    `p.val. adj. (bf=8)`
  ) %>%
  formattable(
    caption = paste("Intercept and slopes of fixed effects:", cur_equation),
    list(`p.val. adj. (bf=8)` = formatter("span", style = p_color))
  ) %>%
  mutate(
    p.value = if_else(
      p.value < 0.001,
      as.character(scientific(p.value, digits=2)),
      as.character(round(p.value, 4), digits=2)
    ),
    `p.val. adj. (bf=8)` = if_else(
      `p.val. adj. (bf=8)` < 0.001,
      as.character(scientific(`p.val. adj. (bf=8)`, digits=2)),
      as.character(round(`p.val. adj. (bf=8)`, 4))
    )
  )

```

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
kable(
    r2_nakagawa(temp_mdl),
    caption = "Conditional and marginal R^2^ of model",
    digits = 2,
    align = "l"
  )

do.call(rbind, r2_nakagawa(temp_mdl))[,1] %>% 
  write.csv("../output/LME_PA_l_f0_r2.csv")

print(
  plot_model(
    temp_mdl,
    title = cur_equation,
    show.intercept = FALSE,
    show.values = TRUE,
    vline.color = "red",
    colors = "Black"
  )
)
```

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
# Get intercepts and pairwise comparisons for each PA

# Get number of levels in target treatment variable.
cur_levels <- levels(m_corpus.trimmed$acc_phon)
orig_levels <- cur_levels
num_levels <- length(cur_levels)
all_models_tidy = tibble()

for (i in 1:(num_levels)) {
  # Test GLMM model on data.
  cur_model <- lmer(
    formula(cur_equation),
    data = m_corpus.trimmed,
    control = lmerControl(
      optimizer = "optimx",
      calc.derivs = FALSE,
      optCtrl = list(
        method = "nlminb",
        starttests = FALSE,
        kkt = FALSE
      )
    )
  )
  
  # Convert the model output into a more readable format.
  cur_model_tidy <- tidy(cur_model) %>%
    filter(term != "sd__(Intercept)",
           effect %notin% "ran_pars") %>%
    select(-c(group, effect))  %>%
    rename(t.value = statistic) %>%
    bonferroniAdjust(8, exclude_terms = c("fin_phonL%", "genderM")) %>%
    rename(`p.val. adj. (bf=8)` = p.adj) %>% 
    mutate(
      estimate = round(estimate, 3),
      std.error = round(std.error, 3),
      t.value = round(t.value, 3),
      df = round(df, 2)
      
      
    )
  
  # Get CIs of model using Wald method (fast, fixed effects only)
  ci_Wald <- confint(cur_model, method = "Wald") %>%
    as_tibble() %>%
    filter(`2.5 %` != "NA") %>%
    mutate(`2.5 %` = round(`2.5 %`, 3),
           `97.5 %` = round(`97.5 %`, 3)) %>%
    rename("2.5% CI" = "2.5 %",
           "97.5% CI" = "97.5 %")
  
  # Bind tidy model to ci_wald
  cur_model_tidy <- cbind(cur_model_tidy, ci_Wald) %>%
    # re-order columns
    select(
      "term",
      "estimate",
      "std.error",
      "2.5% CI",
      "97.5% CI",
      "t.value",
      "df",
      "p.value",
      `p.val. adj. (bf=8)`
    ) %>%
    # Prepare current model for pasting to all models output.
    # Make 'pairwise' column = intercept.
    mutate(
      pairwise =
        if_else(
          term == "(Intercept)",
          "intercept",
          if_else(
            term %notin% paste("acc_phon", cur_levels, sep = ""),
            "N/A",
            paste("acc_phon", cur_levels[1], sep = "")
          )
        ),
      # change 'term' so "intercept" states the target condition name.
      term =
        if_else(
          term == "(Intercept)",
          paste("acc_phon", cur_levels[1], sep = ""),
          term
        )
    )
  
  # make list of pairwise comparisons to keeps
  keep = NULL
  for (j in i:num_levels) {
    keep <- c(keep, paste("acc_phon", orig_levels[j], sep = ""))
    
  }
  # remove pairwise comparisons which have already been done
  cur_model_tidy <- filter(cur_model_tidy, term %in% keep)
  
  # add remaining pairwise comparisons to main tibble.
  all_models_tidy <- bind_rows(all_models_tidy, cur_model_tidy)
  
  # restructure the order of levels for next LME cur_model.
  cur_levels <- c(cur_levels[2:num_levels], cur_levels[1])
  m_corpus.trimmed <- m_corpus.trimmed %>%
    mutate(acc_phon = factor(acc_phon,
                             levels = cur_levels))
}
```

#### Summary of intercepts

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
# print b0 table
filter(all_models_tidy, pairwise == "intercept") %>%
  select(-pairwise) %>%
  rename(intercept = term) %>%
  sigCodesTidy(`p.val. adj. (bf=8)`) %>% mutate(
    p.value = if_else(
      p.value < 0.001,
      as.character(scientific(p.value, digits=2)),
      as.character(round(p.value, 4), digits=2)
    ),
    `p.val. adj. (bf=8)` = if_else(
      `p.val. adj. (bf=8)` < 0.001,
      as.character(scientific(`p.val. adj. (bf=8)`, digits=2)),
      as.character(round(`p.val. adj. (bf=8)`, 4))
    )
  )  %>%
  mutate(intercept = str_replace(intercept, "acc_phon", "")) %>% 
  write_csv("../output/LME_PA_l_f0_b0.csv") %>%
  mutate(intercept = str_replace(intercept, "\\^\\[", "\\\\^\\\\[")) %>%
  formattable(list(`p.val. adj. (bf=8)` = formatter("span", style = p_color)),
              caption = paste("b0 for", cur_equation, sep = " "))
```

#### Summary of pairwise comparisons

```{r fig.height=3, fig.width=9, warning = FALSE, message = FALSE, tidy = TRUE}
# print pairwise comparisons of (B1 table)
filter(all_models_tidy, pairwise %notin% c("intercept", "N/A")) %>%
  select(
    pairwise,
    term,
    estimate,
    std.error,
    `2.5% CI`,
    `97.5% CI`,
    t.value,
    df,
    p.value,
    `p.val. adj. (bf=8)`
  ) %>%
  rename(intercept = pairwise) %>%
  sigCodesTidy(`p.val. adj. (bf=8)`) %>% mutate(
    p.value = if_else(
      p.value < 0.001,
      as.character(scientific(p.value, digits=2)),
      as.character(round(p.value, 4), digits=2)
    ),
    `p.val. adj. (bf=8)` = if_else(
      `p.val. adj. (bf=8)` < 0.001,
      as.character(scientific(`p.val. adj. (bf=8)`, digits=2)),
      as.character(round(`p.val. adj. (bf=8)`, 4))
    )
  ) %>%
  mutate(
    term = str_replace(term, "acc_phon", ""),
    intercept = str_replace(intercept, "acc_phon", "")) %>% 
  write_csv("../output/LME_PA_l_f0_b1.csv")  %>%
  mutate(
    term = str_replace(term, "\\^\\[", "\\\\^\\\\["),
    intercept = str_replace(intercept, "\\^\\[", "\\\\^\\\\[")
    ) %>%  
    rename(`est. (slope)` = estimate) %>% 
formattable(list(`p.val. adj. (bf=8)` = formatter("span", style = p_color)),
            caption = paste("b1 for", cur_equation, sep = " "))

```
