---
title: 'Chapter 07: Analysis of Function: Sentence Modes'
subtitle: '7.5.2 Register-tier analysis of intonational phonology'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

# 7.5.2 Register-tier Analysis

```{r setup B, include=F}
knitr::opts_chunk$set(warning = F, message = F, tidy = "styler")

# Load Functions and colour variables
source("../../2_R_Functions/myFunctions.R") 
source("../../2_R_Functions/myColours.R") 
# Load packages.
installMissingPackages(c("tidyverse",
                         "formattable",
                         "janitor",
                         "broomExtra",
                         "knitr",
                         "RColorBrewer",
                         "ggeffects",
                         "ggrepel",
                         "blme"))

# Get  Data.
m_corpus <- get_m_corpus("../../4_data/m_corpus.csv")  %>%
  mutate(nuc_H_reg = factor( str_detect(acc_phon, "\\^"),
                          levels = c(F, T))) %>% 
  # Select only columns for analysis
  select(speaker,
         gender,
         stim,
         prompt,
         mode,
         acc_phon,
         fin_phon,
         nuc_contour,
         nuc_H_reg)
```

## Raw Data {.tabset}

### Contour by sentence mode

```{r summarise nuc contour by speaker (raw data), warning=F}
m_corpus %>%
  group_by(mode, nuc_contour) %>%
  summarise(acc_count = n(), .groups = "keep") %>%
  pivot_wider(names_from = nuc_contour,
              values_from = acc_count,
              values_fill = 0) %>%
  adorn_totals(where = "row") %>%
  select(
    c(
      "mode",
      "H* L%",
      ">H* L%",
      "^[L*]H L%",
      "L*H %",
      "L*H L%",
      "L*^[H] %",
      "L*^[H L%]",
      "L*^[H] L%",
      "^[L*H] %",
      "^[L*H L%]",
      "^[L*H] L%"
      )
  ) %>%
  rename_all(list(~ str_replace_all(., "([\\*\\[\\^\\>])", "\\\\\\1"))) %>%
  formattable(align = c("l", rep("r",  ncol(.)-1)),
    caption = "Nuclear contours by mode in M-corpus (raw)")



```

### Pitch accents by speaker

```{r summarise PA by speaker (raw data), warning=F}
# Create summary table of Pitch Accent by Speaker.


m_corpus %>%
  group_by(speaker, acc_phon) %>%
  summarise(acc_count = n(), .groups = "keep") %>%
  pivot_wider(names_from = acc_phon,
              values_from = acc_count,
              values_fill = 0) %>%
  select("speaker", "H*", "^[L*]H", ">H*", "L*H", "L*^[H]", "^[L*H]")  %>% 
  rename_all(list(~str_replace_all(., "([\\*\\[\\^\\>])", "\\\\\\1"))) %>%  
  adorn_totals(where = "row") %>% 
  formattable(align = c("l", rep("r",  ncol(.)-1)),
    caption =
      "Distibution by speaker of nuclear pitch accents in M-corpus (raw)",
    list(area(col = 2:7, row= 1:nrow(.)-1) ~ proportion_bar(color = c(
      rep(brewer.pal(6, "Pastel2")[1], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[5], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[4], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[3], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[2], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[6], nrow(.)-1)
      )))
    )

```
### All pitch contours

```{r summarise nuc contour , warning=F, results="asis"}
# Create summary table of Pitch Accent by Speaker.
m_corpus %>%
  group_by(nuc_contour) %>%
  summarise(nuc_count = n(), .groups = "keep") %>% 
  mutate(nuc_contour = str_replace_all(
    nuc_contour, "([\\*\\[\\^\\>])", "\\\\\\1")) %>%
  rename(`nuclear contour` = nuc_contour,
         `count` = nuc_count) %>% 
  arrange(desc(`count`)) %>% 
  kable(caption = "Nuclear pitch contours (raw data)")
```

## Adjusted Data

The raw data is unbalanced since the number of utterances per target speaker varies slightly. For some speakers, there are no repetitions of some target utterances. Therefore, the data has been adjusted to project an estimated balanced number of tokens per speaker per target utterance.

Firstly a tibble is created summarising the proportion of pitch accents per mode per speaker. This is done in three initial stages:

 1. Count the number of phonological tokens per stimulus per speaker.
 2. Convert the counts to proportion per stimulus per speaker.
 3. Convert the proportion per stimulus per speaker to a proportion per stimulus per mode.

From this table summary tables are created:

1. pitch accents per speaker (adjusted)
   a. Calculate the sum all proportions per token per speaker.
   b. Convert proportions to an integer based on an ideal total number of tokens per speaker.
2. pitch accents per mode (adjusted):
   a. Calculate the sum all proportions per token per mode.
   b. Convert proportions to an integer based on an ideal total number of tokens per mode
3. pitch accents by gender and mode (adjusted):
   a. Calculate the sum all proportions per token by mode and gender.
  b. Convert proportions to an integer based on an ideal total number of tokens per mode as if there were an equal numnber of male and female participants.

```{r create tables for adjusted data, warning=F}

# Create tibble summarising  proportion of pitch accents per mode per speaker
# 
pa_adj <- m_corpus %>%
    # Count the number of acc_phon by speaker and mode
    group_by(speaker, stim, acc_phon) %>%
    summarise(nuc_count = n(), .groups = "keep") %>%
    pivot_wider(names_from=acc_phon, values_from=nuc_count, values_fill = 0) %>%
    # convert the totals to proportion per stim.
    adorn_totals("col") %>% 
    mutate(across(3:8,  ~  (. / Total))) %>%
    select(-Total) %>% 
    # Convert stim column to mode.
    rename(mode = stim) %>% 
    mutate(mode = substr(mode, 1, 3)) %>%
    # Convert the values to proportion per mode
    pivot_longer(3:last_col(), "acc_phon") %>% 
    group_by(speaker, mode, acc_phon) %>%
    pivot_wider(names_from=acc_phon,
                values_from=value,
                values_fill = 0,
                values_fn = ~sum(.)) %>%
    adorn_totals("col") %>% 
    mutate(across(3:8, ~ (. / Total)),
           mode = factor(mode, levels = c("MDC", "MWH", "MYN", "MDQ"))
           ) %>%
    select(-Total) %>%
    as_tibble()


# Create tibble of pitch accents per speaker (adjusted).
# 
pa_by_speaker_adj <- pa_adj %>%
  # Get the sum of all token proportions per speaker.
  pivot_longer(3:last_col(), "acc_phon") %>%
  select(-mode) %>% 
  pivot_wider(names_from = acc_phon,
              values_from = value,
              values_fill = 0,
              values_fn = ~ sum(.)
              ) %>%
  # Convert proportions to an integer based on an ideal total number of tokens
  # per speaker.
  mutate(across(2:7, ~ round(. * 15, 0))) %>% 
  select("speaker", "H*", ">H*", "^[L*]H", "L*H", "L*^[H]", "^[L*H]")


# Create tibble of pitch accents per mode (adjusted).
# 
pa_by_mode_adj <- pa_adj %>%
  # Get the sum of all token proportions per speaker.
  pivot_longer(3:last_col(), "acc_phon") %>%
  select(-speaker) %>% 
  pivot_wider(names_from = acc_phon,
              values_from = value,
              values_fill = 0,
              values_fn = ~ sum(.)
              ) %>% 
  # Convert proportions to an integer based on an ideal total number of tokens.
  mutate(across(2:7, ~ round(. * 15, 0))) %>% 
  select("mode", "H*", ">H*", "^[L*]H", "L*H", "L*^[H]", "^[L*H]") %>% 
  arrange(mode)


# Create tibble of pitch accents by gender and mode (adjusted).
# 
pa_by_gender_adj <- pa_adj %>%
  # Convert gender to speaker
  mutate(speaker = substr(speaker, 1, 1),
         speaker = factor(speaker, levels=unique(speaker))) %>% 
  rename(gender = speaker) %>% 
  # Get the sum of all token proportions per speaker.
  pivot_longer(3:last_col(), "acc_phon") %>%
  pivot_wider(names_from = acc_phon,
              values_from = value,
              values_fill = 0,
              values_fn = ~ sum(.)
              ) %>%
  # Convert proportions to an integer based on an ideal total number of tokens
  # per speaker.
  mutate(across(3:8, ~ if_else(gender=="F",
                               round(. * 15, 0),
                               # project equal number of Male & Female speakers.
                               round(. * 18, 0)
                               )
                )
         ) %>% 
  select("mode", "gender", "H*", ">H*", "^[L*]H", "L*H", "L*^[H]", "^[L*H]") %>% 
  arrange(mode) 


# Create tibble of pitch accents by gender and mode (adjusted).
# 
pa_by_mode_speaker_adj <- pa_adj %>%
  # Get the sum of all token proportions per speaker.
  pivot_longer(3:last_col(), "acc_phon") %>%
  pivot_wider(names_from = acc_phon,
              values_from = value,
              values_fill = 0,
              values_fn = ~ sum(.)
              ) %>%
  # Convert proportions to integer based on ideal number of by-speaker tokens.
  mutate(across(3:last_col(), ~ round(.* 16)))%>% 
  # re-order the columns
  select("mode", "speaker", "H*", ">H*", "^[L*]H", "L*H", "L*^[H]", "^[L*H]") %>% 
  arrange(speaker)

```

### Nuclear PA Tables {.tabset}
#### by mode (common)

```{r Print table of pitch accent by mode adjusted excluding outliers}
pa_by_mode_adj %>%
  select("mode", "L*H", "L*^[H]", "^[L*H]") %>%
  rename_all(list(~str_replace_all(., "([\\*\\[\\^\\>])", "\\\\\\1"))) %>%
  adorn_totals("row") %>% 
  formattable(align = c("l", rep("r",  ncol(.)-1)),
    caption = paste("Distibution by of pitch accents by mode in M-corpus",
                    "(adjusted, outliers excluded)"),
    list(area(col=2:4, row=1:nrow(.)-1) ~ proportion_bar(color = c(
      rep(brewer.pal(6, "Pastel2")[3], 4),
      rep(brewer.pal(6, "Pastel2")[2], 4),
      rep(brewer.pal(6, "Pastel2")[6], 4)
      )))
    )

```

#### by mode (all)

```{r print pitch accent by mode adjusted, warning=F}
# Calculate a projected balanced number of tokens for each PA as a function of # sentence mode.
#
# This takes into consideration the number of utterances per speaker per stimulus and the number of speakers per stimulus.

 pa_by_mode_adj %>%
  rename_all(list(~str_replace_all(., "([\\*\\[\\^\\>])", "\\\\\\1"))) %>%  
  adorn_totals("row") %>% 
  formattable(align = c("l", rep("r",  ncol(.)-1)),
    caption=
      "Distibution by mode of nuclear pitch accents in M-corpus (adjusted)",
    list(area(col=2:7, row=1:4) ~ proportion_bar(color = c(
      rep(brewer.pal(6, "Pastel2")[1], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[4], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[5], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[3], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[2], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[6], nrow(.)-1)
      )))
    )
```

#### by speaker

```{r print pa by speaker adjusted, warning=F}
pa_by_speaker_adj %>%
  rename_all(list(~str_replace_all(., "([\\*\\[\\^\\>])", "\\\\\\1"))) %>%  
  adorn_totals(where = "row") %>% 
  formattable(align = c("l", rep("r",  ncol(.)-1)),
    caption="Distibution by speaker of nuclear pitch accents in M-corpus (adjusted)",
    list(area(col = 2:7, row=1:nrow(.)-1) ~ proportion_bar(color = c(
      rep(brewer.pal(6, "Pastel2")[1], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[4], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[5], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[3], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[2], nrow(.)-1),
      rep(brewer.pal(6, "Pastel2")[6], nrow(.)-1)
      )))
    )
```

### Nuclear PA Figures {.tabset}

#### by speaker (chart)

```{r plot pitch accent by speaker adjusted, fig.width = 8, fig.height = 4}

# Percentage Stacked bar
ggplot(data = pa_by_speaker_adj %>%
         # uncount data for plotting
         gather("acc_phon", "count",-c(speaker)) %>%
         uncount(count),
       aes(fill = acc_phon, x = speaker)) +
    geom_bar(
        stat = "count",
        position = "fill",
        width = 0.50,
        colour = "black"
    ) +
    scale_y_continuous(labels = scales::label_percent(scale = 100))+
    scale_fill_manual(values = pitch_accent_colours[1:6]) +
    
    labs(y = "Proportion of Pitch Accents",
         x = "Speaker") +
    theme(panel.border = element_rect(fill = NA))
```

#### by mode (common)

```{r Plot common pitch accents by mode adjusted, fig.height=2.2, fig.width=5}

ggplot(pa_by_mode_adj %>%
         # uncount table for bar chart
         pivot_longer(2:last_col(), "acc_phon") %>%
         uncount(value) %>%
         # exclude outliers.
         filter(acc_phon %in% c("L*H", "L*^[H]", "^[L*H]")) %>%
         mutate(acc_phon = factor(acc_phon,
                                  levels = c("L*H", "L*^[H]", "^[L*H]")))
       ) +
  geom_bar(mapping = aes(x = mode,
                         y = (..count..) / sum(..count..),
                         fill = acc_phon),
           show.legend = F,
           colour = "black",
           position = position_dodge2(preserve = "single")) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.32)) +
  facet_grid(cols = vars(acc_phon)) +
  theme(panel.border = element_rect(fill = NA)) +
  ggtitle("Nuclear pitch accents by sentence mode (adjusted)") +
  scale_fill_manual(values = pitch_accent_colours, name = "Pitch Accent") +
  labs(x = "Mode", y = "tokens (%)") +
  theme(panel.border = element_rect(fill = NA)) +
  geom_text(aes(x = mode,
                y = (..count..) / sum(..count..),
                label = percent((..count..) / sum(..count..), 1)),
  stat = "count",
  vjust = -0.7,
  size = 3)

```


#### by mode (all)
```{r Plot common pitch accents by mode and M8 adjusted, fig.height=2.5, fig.width=8.5}

ggplot(
  pa_by_mode_speaker_adj %>%
    filter(speaker == "M8") %>% 
    # Uncount tables bar chart.
    pivot_longer(3:last_col(), "acc_phon") %>%
    uncount(value) %>% 
    mutate(acc_phon = factor(
      acc_phon,
      levels = c("H*", ">H*", "^[L*]H", "L*H", "L*^[H]", "^[L*H]")))
    ) +
  geom_bar(
    mapping = aes(
      x = mode,
      y = (..count..) / sum(..count..),
      fill = acc_phon
    ),
    show.legend = F,
    colour = "black",
    position = position_dodge2(preserve = "single")
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.28)) +
  facet_grid(cols = vars(acc_phon)) +
  ggtitle("Nuclear pitch accents by sentence mode (adjusted), M8 only") +
  theme(
    panel.border = element_rect(fill = NA)
  ) +
  scale_fill_manual(values = pitch_accent_colours, name = "Pitch Accent") +
  labs(x = "Mode", y = "tokens (%)") +
  theme(panel.border = element_rect(fill = NA)) +
  
  geom_text(aes(
    x = mode,
    y = (..count..)/ sum(..count..),
    label = percent((..count..) / sum(..count..),  1)
  ),
  stat = "count",
  vjust = -0.7,
  size = 3)

```

#### by mode and gender (common)

```{r Plot common pitch accents by mode and gender adjusted, fig.height=3.5, fig.width=5.15}
ggplot(pa_by_gender_adj %>%
    # Uncount tables bar chart.
    pivot_longer(3:last_col(), "acc_phon") %>%
    uncount(value) %>% 
    # Exclude outliers.
    filter(acc_phon %in% c("L*H", "L*^[H]", "^[L*H]")) %>% 
    mutate(acc_phon = factor(acc_phon, levels = c("L*H", "L*^[H]", "^[L*H]")))) +
  
  geom_bar(
    mapping = aes(
      x = mode,
      y = (..count..) / sum(..count..),
      fill = acc_phon
    ),
    show.legend = F,
    colour = "black",
    position = position_dodge2(preserve = "single")
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.16)) +
  facet_grid(rows = vars(gender), cols = vars(acc_phon)) +
  ggtitle("Nuclear pitch accents by sentence mode and gender (adjusted)") +
  theme(
    panel.border = element_rect(fill = NA)
  ) +
  scale_fill_manual(values = pitch_accent_colours, name = "Pitch Accent") +
  labs(x = "Mode", y = "tokens (%)") +
  theme(panel.border = element_rect(fill = NA)) +
  
  geom_text(aes(
    x = mode,
    y = (..count..) / sum(..count..),
    label = percent((..count..) / sum(..count..), 1)
  ),
  stat = "count",
  vjust = -0.7,
  size = 3)

```

#### by mode and speaker (all)

```{r Plot common pitch accents by mode and speaker adjusted, fig.height=12, fig.width=8.5}

ggplot(
  pa_by_mode_speaker_adj %>%
    # Uncount tables bar chart.
    pivot_longer(3:last_col(), "acc_phon") %>%
    uncount(value) %>% 
    mutate(acc_phon = factor(
      acc_phon,
      levels = c("H*", ">H*", "^[L*]H", "L*H", "L*^[H]", "^[L*H]")))
    ) +
  
  geom_bar(
    mapping = aes(
      x = mode,
      y = 11 * (..count..) / sum(..count..),
      fill = acc_phon
    ),
    show.legend = F,
    colour = "black",
    position = position_dodge2(preserve = "single")
  ) +

  geom_text(aes(x = mode,
                y = 11 * (..count..)/ sum(..count..),
                label = 11 * percent((..count..) / sum(..count..),  0)),
            stat = "count",
            vjust = -0.35,
            size = 2.5) +
  
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.30)) +
  
  facet_grid(rows = vars(speaker), cols = vars(acc_phon)) +
  
  ggtitle("Nuclear pitch accents by mode and speaker (adjusted)") +
  
  theme(panel.border = element_rect(fill = NA),
        axis.text.x = element_text(angle = 90, size = 7.5),
        axis.text.y = element_text(size = 7.5),
        #axis.text.x = element_text(, vjust = 0.5, hjust=1)
        ) +
  
  scale_fill_manual(values = pitch_accent_colours, name = "Pitch Accent") +
  
  labs(x = "Mode", y = "tokens (%)") +
  
  theme(panel.border = element_rect(fill = NA))

ggsave(
  filename =
    "output/nuc_PA_by_mode_and_speaker.png",
  width = 15.45,
  height = 24,
  units = "cm",
  dpi = 300
)

```



## Boundary Tone Analysis {.tabset}

```{r Wrangle reg data for boundary analysis}
fin_phon_adj <- m_corpus %>%
  mutate(fin_phon = str_replace_all(fin_phon, "\\[|\\]|\\^", "")) %>%
  mutate(fin_phon = factor(fin_phon, levels = unique(fin_phon))) %>% 
  # Count the number of fin_phon by speaker and mode
  group_by(speaker, stim, fin_phon) %>%
  summarise(nuc_count = n(), .groups = "keep") %>%
  pivot_wider(names_from = fin_phon,
              values_from = nuc_count,
              values_fill = 0) %>%
  # convert the totals to proportion per stim.
  adorn_totals("col") %>%
  mutate(across(3:(last_col()-1),  ~  (. / Total))) %>%
  select(-Total) %>%
  # Convert stim column to mode.
  rename(mode = stim) %>%
  mutate(mode = substr(mode, 1, 3)) %>%
  # Convert the values to proportion per mode
  pivot_longer(3:last_col(), "fin_phon") %>%
  group_by(speaker, mode, fin_phon) %>%
  pivot_wider(
    names_from = fin_phon,
    values_from = value,
    values_fill = 0,
    values_fn = ~ sum(.)
  ) %>%
  adorn_totals("col") %>%
  mutate(across(3:(last_col()-1), ~ (. / Total)),
         mode = factor(mode, levels = c("MDC", "MWH", "MYN", "MDQ"))) %>%
  select(-Total) %>%
  as_tibble()

# Create tibble of final boundaries per mode (adjusted).
fin_phon_by_mode_adj <- fin_phon_adj %>%
  # Get the sum of all token proportions per speaker.
  pivot_longer(3:last_col(), "fin_phon") %>%
  select(-c(speaker)) %>% 
  pivot_wider(names_from = fin_phon,
              values_from = value,
              values_fill = 0,
              values_fn = ~ sum(.)
              ) %>% 
  # Convert proportions to an integer based on an ideal total number of tokens.
  mutate(across(2:last_col(), ~ round(. * 15, 0))) %>% 
  arrange(mode)

# Create tibble of final boundaries  by gender and mode (adjusted).
fin_phon_by_gender_adj <- fin_phon_adj %>% 
  # Convert speaker to gender
  mutate(speaker = substr(speaker, 1, 1),
         speaker = factor(speaker, levels=unique(speaker))) %>% 
  rename(gender = speaker) %>% 
  # Get the sum of all token proportions per speaker.
  pivot_longer(3:last_col(), "fin_phon") %>%

  pivot_wider(names_from = fin_phon,
              values_from = value,
              values_fill = 0,
              values_fn = ~ sum(.)
              ) %>%
  # Convert proportions to an integer based on an ideal total number of tokens
  # per speaker.
  mutate(across(3:last_col(), ~ if_else(gender=="F",
                               round(. * 15, 0),
                               # project equal number of Male & Female speakers.
                               round(. * 18, 0)
                               )
                )
         ) %>% 
  arrange(mode)

# Make lists of projected boundary tokens.
fin_phon_by_mode_adj_list <- fin_phon_by_mode_adj %>%
  pivot_longer(2:last_col(), "fin_phon") %>%
    uncount(value) %>%
    mutate(fin_phon = factor(fin_phon, levels = unique(fin_phon)))


# Get list of projected boundary tokens by mode and gender.
fin_phon_by_gender_adj_list <- fin_phon_by_gender_adj %>%
    pivot_longer(3:last_col(), "fin_phon") %>%
    uncount(value) %>% 
    mutate(fin_phon = factor(fin_phon, levels = unique(fin_phon)))

```

### by sentence mode

```{r plot fin_phon by mode only mode reg, fig.width = 5, fig.height = 2.5}

ggplot(fin_phon_by_mode_adj_list) +
  geom_bar(
    mapping = aes(
      x = mode,
      y = (..count..) / sum(..count..),
      fill = fin_phon
    ),
    show.legend = F,
    colour = "black",
    position = position_dodge2(preserve = "single")
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.28)) +
  facet_grid(cols = vars(fin_phon)) +
  theme(
    panel.border = element_rect(fill = NA)
  ) +
  ggtitle("Final boundary by mode (adjusted, ^[L%]=L%)") +
  scale_fill_manual(values = fin_phon_colours, name = "final boundary") +
  labs(x = "Mode", y = "tokens (%)") +
  theme(panel.border = element_rect(fill = NA)) +
  geom_text(aes(
    x = mode,
    y = (..count..) / sum(..count..),
    label = percent((..count..) / sum(..count..), 1)
  ),
  stat = "count",
  vjust = -0.75,
  size = 3)


```

### by sentence mode and gender

```{r plot fin_phon by mode and gender mode reg, fig.width = 5, fig.height = 4}
ggplot(fin_phon_by_gender_adj_list) +
  geom_bar(
    mapping = aes(
      x = mode,
      y = (..count..) / sum(..count..),
      fill = fin_phon
    ),
    show.legend = F,
    colour = "black",
    position = position_dodge2(preserve = "single")
  ) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 0.13)) +
  facet_grid(rows = vars(gender), cols = vars(fin_phon)) +

  theme(
    panel.border = element_rect(fill = NA)
  ) +
  ggtitle("Final boundary by mode and gender") +
  scale_fill_manual(values = fin_phon_colours, name = "nuclear contour") +
  labs(x = "Mode", y = "tokens (%)") +
  theme(panel.border = element_rect(fill = NA),
        plot.title = element_blank()) +
  geom_text(aes(
    x = mode,
    y = (..count..) / sum(..count..),
    label = percent((..count..) / sum(..count..), 1)
  ),
  stat = "count",
  vjust = -0.75,
  size = 2.75)

ggsave(
  filename =
    "output/fin_phon_reg_tier.png",
  width = 8.5,
  height = 8,
  units = "cm",
  dpi = 300
)

```


## BGLMM Analysis

### H-Register and Mode {.tabset}

```{r run model m_corpus_h_reg, results = "hide"}

m_corpus_h_reg <- m_corpus %>%
  select(prompt, speaker, gender, nuc_H_reg, mode)
  

h_reg_model <- bglmer(
  nuc_H_reg ~ mode + gender + (1 | speaker)+ (1 | prompt),
  data = m_corpus_h_reg,
  family = binomial(link = "logit"),
  fixef.prior = normal()
  )


h_reg_model <- optimizeModel(h_reg_model, reject_nm = F)
```

#### LRT

```{r drop1 m_corpus_h_reg, fig.width = 3.05, fig.height = 2.5,}
write = "output/Mode_PA_GLMM"

model_formula <- getModelFormula(h_reg_model)

write(paste(str_replace_all(model_formula, "\\`", ""), sep = ""),
      paste(write, "_formula.txt", sep = ""))


# get, save, and print ANOVA of model
drop1_analysis <- drop1(h_reg_model, test = "Chisq") %>%  tidy() %>% 
  filter(!is.na(npar)) %>%
  mutate(`p.adj (BH)` = NA, signif. = NA) %>%
  rename(factor = term) %>% 
  rename(`Pr(Chi)` = Pr.Chi.) %>% 
  write_csv(paste0(write, "_anova.csv"))


adjustP_posthoc("output/",
                `Pr(Chi)`,
                suffix_id = "_anova")

read_csv(paste(write, "_anova.csv", sep = "")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("Drop 1 ChiSq test of model:", model_formula))

```

#### *R*^2^ and prediction plots

```{r Tidy Summary m_corpus_h_reg, fig.width = 3.05, fig.height = 2.5, results="asis"}
tidy_summary <- analyseModel(
  h_reg_model,
  write = write,
  is_GLM = T,
  type = "pred",
  caption_suffix = " (reg tier)",
  plot_rounding = 2,
  short_caption = T,
  caption_prefix = "A. Probability of"
)
```


```{r Tidy Summary m_corpus_h_reg gender, fig.width = 8.5/2.54, fig.height = 6.25/2.54, results="asis"}
ggpredict(h_reg_model,
            terms = c("mode", "gender"),
            ci.lvl = 0.95) %>%
  plot() +
    labs(caption = "A. Predicted prob. of nuc_H_ref by mode and gender.") +
  geom_text_repel(
    aes(label = round(predicted, 2) %>% niceRatio()),
    size = 3,
    label.padding = units(3, "mm"),
    point.padding = units(2, "mm"),
  ) +
   scale_color_manual(labels = c("Female", "Male"),
                      values = c("#000000" ,"#707070")) +
  scale_y_continuous(limits = c(0, 1)) +
  ylab("Probability") +
  theme(
    panel.spacing.x = unit(1, "mm"),
    plot.title = element_blank(),
    axis.title.x = element_blank(),
    plot.caption = element_text(vjust = 1, hjust = 0, size = 10),
    plot.caption.position = "plot",
    legend.position = c(0.175, 0.975),
    legend.justification = "top",
    legend.margin = margin(0,1,0,1, unit = "mm"),
    legend.spacing.y = unit(0, "mm"),
    legend.title = element_blank(),
    legend.background = element_rect(colour = "darkGray")
  )

ggsave(
  filename =
    paste0(write, "_multi_pred.png"),
  # width = 8.14,
  width = 8.5,
  height = 6.26,
  units = "cm",
  dpi = 300
)

```


#### prediction by factor
```{r  prediction h_reg, fig.width = 3.05, fig.height = 2.5, results="asis"}
printTidyPredictions(h_reg_model,
                     caption_suffix = " (reg tier)",
                     digits = 3)
```

#### prediction by mode and gender
```{r  prediction h_reg together, fig.width = 3.05, fig.height = 2.5, results="asis"}
x <- printTidyPredictions(h_reg_model,
                     caption_suffix = " (reg tier)",
                     digits = 3,
                     factor_matrix = T) 
```

#### Pairwise comparisons

```{r GLMM of H re mode Figures and Charts reg, fig.width = 3.05, fig.height = 2.5, results="asis"}
charts <- getModelFixedFX(
  h_reg_model,
  write = write,
  exponentiate = T
)
#charts$intercepts %>% tidyStatNumbers()
charts$slopes %>% tidyStatNumbers()
```

#### Model printout
```{r h reg model summary}
summary(h_reg_model)
```

### Final Boundary and Mode {.tabset}

* Random intercepts only since models with random slopes generate singularity issues.

```{r run GLMM fin phon model, results = "hide"}
m_corpus_boundaries <- m_corpus %>% 
  mutate(`L%` = factor(str_replace_all(fin_phon, "\\[|\\]|\\^", ""),
                           levels=c("%", "L%")))
                           
fin_phon_model <- bglmer(
  `L%` ~ mode + gender + (1 | speaker) + (1 | prompt),
  data = m_corpus_boundaries,
  family = binomial(link = "logit"),
  control = glmerControl(
    optimizer = "optimx",
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

fin_phon_model <- optimizeModel(fin_phon_model, reject_nm = F)

```

#### LRT

```{r drop1 fin phon, fig.width = 3.05, fig.height = 2.5,}
write = "output/Mode_fin_phon_GLMM"

model_formula <- getModelFormula(fin_phon_model)

write(paste(str_replace_all(model_formula, "\\`", ""), sep = ""),
      paste(write, "_formula.txt", sep = ""))


# get, save, and print ANOVA of model
drop1_analysis <- drop1(fin_phon_model, test = "Chisq")

tibble(factor = row.names(drop1_analysis))  %>%
  cbind(tibble(drop1_analysis)) %>%
  filter(!is.na(npar)) %>%
  mutate(`p.adj (BH)` = NA, signif. = NA) %>%
  write_csv(paste0(write, "_anova.csv"))


adjustP_posthoc("output/",
                `Pr(Chi)`,
                suffix_id = "_anova")

read_csv(paste(write, "_anova.csv", sep = "")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("Drop 1 ChiSq test of model:", model_formula))

```
#### *R*^2^ and prediction plots

```{r Tidy Summary fin phon, fig.width = 3.4, fig.height = 2.5, results="asis"}
tidy_summary <- analyseModel(
  fin_phon_model,
  write = write,
  is_GLM = T,
  type = "pred",
  caption_suffix = " (reg-tier analysis)",
  plot_rounding = 2,
  short_caption = T,
  caption_prefix = "A. Probability of"
)
```


```{r Tidy Summary fin phon gender, fig.width = 6.1, fig.height = 2.5, results="asis"}
ggpredict(fin_phon_model,
            terms = c("mode", "gender"),
            ci.lvl = 0.95) %>%
  plot(grid = T, colors = "bw") +
  geom_label(
    aes(label = round(predicted, 2)),
    label.padding = unit(0.5, "mm"),
    label.r = unit(0.0, "mm"),
    size = 3
  ) +
  scale_y_continuous(limits = c(0, 1)) +
  ylab("Probability") +
  theme(
    plot.title = element_blank(),
    axis.title.x = element_blank(),
    plot.caption = element_text(hjust = 0, size = 10),
    plot.caption.position = "plot",
    legend.position = c(0.12, 0.8),
    legend.background = element_rect(colour = "darkGray")
  )

ggsave(
  filename =
    paste0(write, "_multi_pred.png"),
  # width = 8.14,
  width = 14,
  height = 6.26,
  units = "cm",
  dpi = 300
)
```

#### prediction by factor
```{r  prediction fin_phon, fig.width = 3.05, fig.height = 2.5, results="asis"}
printTidyPredictions(fin_phon_model,
                     caption_suffix = " (reg tier)",
                     digits = 3)
```

#### prediction by mode and gender
```{r  prediction fin_phon together, fig.width = 3.05, fig.height = 2.5, results="asis"}
x <- printTidyPredictions(fin_phon_model,
                     caption_suffix = " (reg tier)",
                     digits = 3,
                     factor_matrix = T) 
```


#### pairwise comparisons

```{r GLMM fin_phon Figures and Charts, fig.width = 3.05, fig.height = 2.5, results="asis"}
charts <- getModelFixedFX(
  fin_phon_model,
  write = write,
  exponentiate = T
)
#charts$intercepts
charts$slopes %>% tidyStatNumbers()
```

#### Model printout

```{r summary of GLMM fin phon model}

summary(fin_phon_model)

```
