---
title: 'Chapter 7. Analysis of Function: Sentence modes'
author: "Antoin Eoin Rodgers"
date: '2022-07-05'
output:
  html_document: default
subtitle: 'Appendix: Posthoc p value adjustment'
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, warning = F, message = F)
```

```{r load functions, include=F}
source("../functions/myFunctions.R") 

installMissingPackages("tidyverse")
```

Adjust all p values using F discovery rate (FDR). $\beta_0$, $\beta_1$, and model ANOVAs are processed separately.

c.f. Benjamini, Y., and Hochberg, Y. (1995). Controlling the F discovery rate: a practical and powerful approach to multiple testing. Journal of the *Royal Statistical Society Series B, **57***, 289--300. <doi:10.1111/j.2517-6161.1995.tb02031.x>. <https://www.jstor.org/stable/2346101>.

## 1. Mode and Phonology

#### a. $\beta_0$

```{r Phonology and Mode b0, results="asis"}
adjustP_posthoc(
    "1.Phonology and Mode/", p.value, print=T, marginal=F, suffix_id="_b0")
```

#### b. $\beta_1$

```{r Phonology and Mode b1, results="asis"}
adjustP_posthoc(
    "1.Phonology and Mode/", p.value, print=T, marginal=F, suffix_id="_b1")
```

### c. ANOVAs for phonological analysis models

```{r PA and PHonological Parameters ANOVAs, results="asis"}
adjustP_posthoc("1.Phonology and Mode/", `Pr(>Chisq)`, suffix_id="_anova",
                print=TRUE)
```


## 2. Mode, acc_phon, and phonetic parameters

#### a. $\beta_0$

```{r Mode Nuc PA and Phonetic Parameters b0, results="asis"}
adjustP_posthoc("output/", p.value, print=T, marginal=F, suffix_id="_b0")
```

#### b. $\beta_1$

```{r Mode Nuc PA and Phonetic Parameters b1, results="asis"}
adjustP_posthoc("output/", p.value, print=T, marginal=F, suffix_id="_b1")
```

### c. ANOVAs of models for mode, acc_phon, and phonetic parameters

```{r PA and Phonetic Parameters ANOVAs, results="asis"}
adjustP_posthoc("output_model_anovas/", p.value, suffix_id="_anova",
                print=TRUE)
```
