---
title: "Visual Analysis of PN peaks in A-Corpus PNs (H\\*, >H\\*, L\\*H)"
author: "Antoin Rodgers"
date: "18.05.2022"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F)

## Get personal functions and install / load necessary packages.
source("../../2_R_Functions/myFunctions.R")

installMissingPackages(c("tidyverse", "formattable","janitor","mefa4", "knitr",
                         "gridExtra", "RColorBrewer"))

## Set themes and colour schemes.
theme_set(theme_minimal(base_size = 10))
```

```{r get data, include=F}
## Get  data for analysis (from anacrusis set)
ah_dir <- "../../4_data/"
source("../0_get_AH.R")

# Get PN data
h_star_type <-  c("H*", ">H*",  "L*H")

pn_wrd_stims <- c("A0423", "A1422","A2422", "A3422")

pn_wrd_sent <- c("Valerie's is valid.",
                 "The valley's by the river",
                 "There's a valley with a river",
                 "There was a valley with a river")

pn_phr <- c("Valerie's is",
            "valley's by the river",
            "valley with a",
            "valley with a")

pn <- pn %>%
  select(
    sent,
    speaker,
    stim,
    acc_phon,
    wrd_end_syl,
    h_t,
    v_grand_mean_t,
    h_grand_mean_t,
    h_syl,
    h_syl_ratio,
    ana_syls
    ) %>%
  # Get time normalised peaks
    mutate(
      ana_syls = as.character(ana_syls) %>% as.numeric(),
      h_syl_ft = as.integer(h_syl) - ana_syls,
      h_syl_norm_t = h_syl_ft + h_syl_ratio - 1,
      h_grand_mean_t_re_v = h_grand_mean_t - v_grand_mean_t
      ) %>% 
  # remove parameters no longer needed
  select(-c(v_grand_mean_t, h_grand_mean_t))

pn_wrd_data <- pn %>%
  filter(acc_phon %in% h_star_type,
         stim %in% pn_wrd_stims) %>%
  mutate(acc_phon = factor(acc_phon, levels = h_star_type))


pn_wrd_adj <- pn %>% 
  filter(ana_syls %in% c(1,2,3)) %>% 
  balancedData(ana_syls, acc_phon,"", 11, 5) %>%
  pivot_longer(values_to = "acc_phon", cols = 2:last_col()) %>% uncount(acc_phon) %>%
  rename(acc_phon = name)

# Get NUC data
nuc <- nuc %>% 
  select(speaker, stim, code, fin_phon, e_f0_z, e_f0)

rm(nuc_foot, pn_lex, pn_ana, pn_foot, corpus, nuc_pre, stress)


```


# PN Pitch Accents

## Check Disribution of tokens per speaker {.tabset}

### All tokens

```{r check data 1}
pn %>%
  filter(acc_phon %in% h_star_type) %>%
  mutate(acc_phon = str_replace_all(acc_phon, "([\\*\\>])", "\\\\\\1")) %>% 
  group_by(acc_phon, ana_syls) %>%
  summarise(count = n()) %>% pivot_wider(values_from = count,
                                         names_from = acc_phon,
                                         values_fill = 0) %>%
  arrange(ana_syls)  %>%
  adorn_totals(where = "col") %>%
  adorn_totals(where = "row") %>% 
  formattable(caption = "Distribution of prenuclear PAs with H targets in H-corpus (raw data)")

pn_wrd_data %>%
  filter(acc_phon %in% h_star_type) %>%
  group_by(speaker, ana_syls) %>%
  summarise(count = n()) %>% pivot_wider(values_from = count,
                                         names_from = speaker,
                                         values_fill = 0)  %>% 
  adorn_totals(where = "col") %>%
  adorn_totals("row", fill = "") %>%
  formattable(caption = "L\\*H, \\>H\\*, and H\\*")

```
It appears that ana_syls_0 is not evenly represented across speakers, so it will be excluded

### All tokens except ana_syls_0
```{r check data excluding ana_0 condition}
pn_wrd_data <- pn_wrd_data %>%
  filter(acc_phon %in% h_star_type,
         ana_syls != 0) %>% 
  mutate(ana_syls = factor(ana_syls, levels = c(1,2,3)),
    pn_wrds = ana_syls,
    pn_wrds = str_replace(pn_wrds, "1",  "The valley's by the"),
    pn_wrds = str_replace(pn_wrds, "2",  "There's a valley with a"),
    pn_wrds = str_replace(pn_wrds, "3",  "There was a valley with a"))

pn_wrd_data %>%
  filter(acc_phon %in% h_star_type,
         ana_syls != 0) %>% 
  group_by(speaker, ana_syls) %>%
  summarise(count = n()) %>%
  pivot_wider(values_from = count,
              names_from = speaker,
              values_fill = 0)  %>%
  adorn_totals("col") %>%
  adorn_totals("row", fill = "") %>%
  formattable(caption = "ana_syls 1 to 3 only")
```

```{r Get means etc }
pn_wrd_data_sum <- pn_wrd_data %>%
  group_by(speaker, acc_phon, ana_syls) %>%
  summarise(mean_gm_h_t = mean(h_grand_mean_t_re_v)) %>% 
  arrange(ana_syls)
```
### Graph of PA distribution by ana_syls condition

```{r adjusted PA distribution, fig.width = 6.1, fig.height = 3}
ggplot(data = pn_wrd_adj %>%
                    filter(acc_phon %in% h_star_type) %>% 
                    mutate(acc_phon = factor(acc_phon, levels = h_star_type),
                           pn_wrds = factor(ana_syls, levels = c(1,2,3))),
  mapping = aes(x = ana_syls, fill = acc_phon)
) +
  geom_bar(stat = "count",
           position = "dodge",
           colour = "black") +
  
  geom_text(aes(
    label = (..count..),
    vjust = -0.5
  ),
  stat = "count",
  position = position_dodge(.9),
  size = 3) +
  
  #facet_wrap(vars(pairing)) +
  scale_fill_manual(
    values = brewer.pal(4, "Dark2"),
    name = NULL
  ) +
  theme(
    legend.position = "bottom",
    legend.key.size = unit(0.4, "cm"),
    panel.grid.major = element_blank()
  ) +
  ylim(0, 60) +
  xlab("prenuclear pitch accent") +
  ylab("count (adjusted)")

```

## Cumulative Density Distibution{.tabset}

### >H\*, H\* Peak alignment
```{r plot H types only  cumulative, fig.width = 5, fig.height = 3}
ggplot(
  data = pn_wrd_data %>% filter(acc_phon != "L*H"),
  mapping = aes(x = h_grand_mean_t_re_v
                , fill = acc_phon
                )
) +
  geom_dotplot(stackgroups = TRUE, binpositions = "all", stackratio = 1.2) +
  geom_density(stat = "density", alpha = 0.4, fill = "grey") +
  #facet_wrap(vars(ana_syls)) +
  scale_fill_manual(
    values = brewer.pal(3, "Dark2"),
    name = NULL
  ) +
  theme(
    legend.position = c(.15, .7),
    legend.margin = margin(-6, 6, -6, 6),
    legend.background = element_rect(),
    legend.key.size = unit(1, "cm"),
    panel.grid.major.x = element_line(color = "grey",
                                      size = 0.5),
    panel.grid.minor.x = element_blank() +
    theme_minimal(base_size = 12)
    )+
  ylim(0, 0.01) +
  xlab("grand syllable-mean normalised time (ms)") +
  ylab("density") +
  labs(title = "PN Peak Distribution (A1422, A2422, and A3422, H* and >H*)")
```

### Peak alignment (all PAs)
```{r plot acc_phon cumulative, fig.width = 5, fig.height = 3}
ggplot(
  data = pn_wrd_data,
  mapping = aes(x = h_grand_mean_t_re_v
                , fill = acc_phon
                )
) +
  geom_dotplot(dotsize = 0.925, stackgroups = TRUE, binpositions = "all", stackratio = 0.8) +
  geom_density(stat = "density", alpha = 0.4, fill = "grey") +
  #facet_wrap(vars(ana_syls)) +
  scale_fill_manual(
    values = brewer.pal(3, "Dark2"),
    name = NULL
  ) +
  theme(
    legend.position = c(.85, .7),
    legend.margin = margin(-6, 6, -6, 6),
    legend.background = element_rect(),
    legend.key.size = unit(1, "cm"),
    panel.grid.major.x = element_line(color = "grey",
                                      size = 0.5),
    panel.grid.minor.x = element_blank() +
    theme_minimal(base_size = 12)
    )+
  ylim(0, 0.01) +
  xlab("grand syllable-mean normalised time (ms)") +
  ylab("density") +
  labs(title = "PN Peak Distribution in A1422, A2422, and A3422")

```

# Boundary Tones

## Distribution of tokens {.tabset}

### Boundary Tone by Speaker
```{r check boundary tone by speaker}
nuc %>%
  group_by(speaker, fin_phon) %>%
  summarise(count = n()) %>% pivot_wider(values_from = count,
                                         names_from = speaker,
                                         values_fill = 0)  %>% 
  adorn_totals(where = "col") %>%
  adorn_totals("row", fill = "") %>%
  formattable(caption = "A and H corpora fin_phon per speaker")
```

It is clear that only F12 uses L% with any regularity.

### Representation of Stimuli
```{r check data 2}
nuc %>%
  group_by(speaker, stim) %>%
  summarise(count = n()) %>% pivot_wider(values_from = count,
                                         names_from = speaker,
                                         values_fill = 0) %>% 
  adorn_totals(where = "col") %>%
  adorn_totals("row", fill = "") %>%
  formattable(caption = "A and H corpora repetitions per stimulus per speaker")

nuc %>%
  group_by(stim, fin_phon) %>%
  summarise(count = n()) %>% pivot_wider(values_from = count,
                                         names_from = fin_phon,
                                         values_fill = 0)  %>% 
  adorn_totals(where = "col") %>%
  adorn_totals("row", fill = "") %>%
  formattable(caption = "A and H corpora repetitions per stimulus")
```
A0423, A1211, and A1231 are underrepresented. These will be removed from the dataset so that repetitions per speaker per target are balanced across utterances.

### Well-represented Stimuli Only
```{r check data 3}
unbalanced = c("A0423", "A1211", "A1231")
nuc_balanced <- nuc %>% filter(stim %notin% unbalanced)

nuc_balanced %>%
  group_by(speaker, fin_phon) %>%
  summarise(count = n()) %>% pivot_wider(values_from = count,
                                         names_from = speaker,
                                         values_fill = 0)  %>% 
  adorn_totals(where = "col") %>%
  adorn_totals("row", fill = "") %>%
  formattable(caption = "A and H corpora fin_phon per speaker (excl. A0423, A1211, A1231)")

nuc_balanced %>%
  mutate(fin_phon = str_replace_all(fin_phon, "([\\*\\>])", "\\\\\\1")) %>% 
  group_by(fin_phon, stim) %>%
  summarise(count = n()) %>% pivot_wider(values_from = count,
                                         names_from = fin_phon,
                                         values_fill = 0) %>%
  arrange(stim)  %>%
  adorn_totals(where = "col") %>%
  adorn_totals(where = "row") %>% 
  formattable(caption = "Distribution of nuclear boundaries in A- and H-corpora (excl. A0423, A1211, A1231)")


```


## Means and Standard Deviations
```{r Get means and SD}

nuc_balanced_sum <- nuc_balanced %>%
  group_by(fin_phon) %>%
  summarise(e_f0_z = mean(e_f0_z)) %>% 
  arrange(fin_phon)

nuc_balanced_sum %>% formattable(caption = "Average peak to fin_phon excursion size (ST)")

nuc_balanced %>%
  group_by(fin_phon) %>%
  summarise(e_f0_z = sd(e_f0_z)) %>% 
  arrange(fin_phon) %>% 
  formattable(caption = "SD peak to fin_phon excursion size (ST)")
```
## Graphs {.tabset}
### Boundary scaling All Speakers
```{r Density and dotplot for all, fig.width = 5, fig.height = 3}
my_palette <- brewer.pal(3, "Dark2")[2:3]


ggplot(
  data = nuc_balanced,
  mapping = aes(x = e_f0_z, fill = fin_phon)
  ) +
  geom_dotplot(stackgroups = TRUE, binpositions = "all", stackratio = 0.24) +
  geom_density(stat = "density", alpha = 0.4, fill = "grey") +
  scale_fill_manual(
    values = my_palette,
    name = NULL
    ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = c(.9, .8),
    legend.margin = margin(-6, 6, -6, 6),
        legend.background = element_rect(),
    legend.key.size = unit(1, "cm"),
    panel.grid.major.x = element_line(color = "grey",
                                      size = 0.5),
    panel.grid.minor.x = element_blank()) +
  xlab("f0 (z-score)") +
  ylab("density") +
  xlim(-4, 4) +
 ylim(0, 0.4) +
  labs(title = "Density Distribution of Boundary f0 in A-and H-Corpora")
```

### Boundary Scaling, F12 only 
```{r plot F12 only, fig.width = 5, fig.height = 3}

ggplot(
  data = nuc %>% 
    filter(speaker == "F12"),
  mapping = aes(x = e_f0, fill = fin_phon)
  ) +
  geom_dotplot(binpositions = "all") +
  geom_density(stat = "density", alpha = 0.4, fill = "grey") +
  scale_fill_manual(
    values = my_palette,
    name = NULL
  ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = c(.9, .8),
    legend.margin = margin(-6, 6, -6, 6),
        legend.background = element_rect(),
    legend.key.size = unit(1, "cm"),
    panel.grid.major.x = element_line(color = "grey",
                                      size = 0.5),
    panel.grid.minor.x = element_blank()) +
  xlab("f0 (ST re speaker median)") +
  ylab("density") +
 # xlim(-4, 4) +
  ylim(0, 0.2) +
  labs(title = "Density Distribution of Boundary f0 in A- and H- corpora, F12")

```
