---
title: 'Chapter 7. Analysis of Function: Sentence modes'
author: "Antoin Eoin Rodgers"
date: '2022-07-05'
output:
  html_document: default
subtitle: 'Appendix: Posthoc p value adjustment'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
```

```{r load functions, include=FALSE}
source("../functions/myFunctions.R") 

installMissingPackages("tidyverse")
```

Adjust all p values using false discovery rate (FDR).
$\beta_0$ and $\beta_1$ are processed separately.


c.f. Benjamini, Y., and Hochberg, Y. (1995). Controlling the false discovery rate: a practical and powerful approach to multiple testing. Journal of the *Royal Statistical Society Series B, **57***, 289--300. <doi:10.1111/j.2517-6161.1995.tb02031.x>. <https://www.jstor.org/stable/2346101>.

## 1. Mode and Phonology
#### a. $\beta_0$

```{r Phonology and Mode b0, results="asis"}
adjustP_posthoc("1.Phonology and Mode/", p.value, print=T, marginal=F, suffix_id="_b0")
```
#### b. $\beta_1$
```{r Phonology and Mode b1, results="asis"}
adjustP_posthoc("1.Phonology and Mode/", p.value, print=T, marginal=F, suffix_id="_b1")
```

## 2. Utterance-wide f0 parameters
#### a. $\beta_0$
```{r Utterance Phonetic Parameters and Mode b0, results="asis"}
adjustP_posthoc("2.Utterance and phonetic params/",
                p.value,
                print=TRUE,
                suffix_id="_b0")
```
#### b. $\beta_1$
```{r Utterance Phonetic Parameters and Mode b1, results="asis"}
adjustP_posthoc("2.Utterance and phonetic params/",
                p.value,
                print=TRUE,
                suffix_id="_b1")
```

## 3. Mode, acc_phon, and phonetic parameters
#### a. $\beta_0$

```{r Mode Nuc PA and Phonetic Parameters b0, results="asis"}
adjustP_posthoc("output/", p.value, print=T, marginal=F, suffix_id="_b0")
```

#### b. $\beta_1$

```{r Mode Nuc PA and Phonetic Parameters b1, results="asis"}
adjustP_posthoc("output/", p.value, print=T, marginal=F, suffix_id="_b1")
```

### c. ANOVAs of models for mode, acc_phon, and phonetic parameters

```{r PA and Phonetic Parameters ANOVAs, results="asis"}
adjustP_posthoc("output_model_anovas/", p.value, suffix_id="_anova",
                print=TRUE)
```
