---
title: "Chapter 7 - LME analysis of f0_exc ~ Mode and Nuclear pitch accents: "
author: "Antoin Rodgers"
date: "5.6.2022"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      warning=FALSE,
                      message=FALSE)

# Load Functions.
source("../../functions/myFunctions.R") 

# Load packages.
installMissingPackages(
  c(
    # Include statistical packages.
    "performance",
    "lmerTest",
    "lme4",
    "optimx",
    "MuMIn",
    # Include packages for tidy output.
    "tidyverse",
    "broomExtra",
    "sjPlot",
    "formattable",
    "knitr",
    "RColorBrewer",
    # Include packages for %in% %notin% syntactic notation
    "mefa4"
  )
)

# Set themes and colour schemes.
theme_set(theme_minimal(base_size = 14))

```

```{r load corpus, include=FALSE}
l_star_h_types = c("L*H", "^[L*]H", "L*^[H]", "^[L*H]")

# Load Corpus
m_corpus <- get_m_corpus("../data/m_corpus.csv")  %>%
  # Consolidate nuc_contour based on observation and types of register shift.
  mutate(
    nuc_contour =
      if_else(
        nuc_contour == "^[L*H %]" | nuc_contour == "^[L*H L%]",
        "^[L*H] %",
        if_else(nuc_contour == "L*^[H L%]", "L*^[H] %", nuc_contour)
      ),
    acc_phon =
      factor(
        str_replace(nuc_contour,
                    "\\s%$|\\sL%$",
                    ""),
        levels = c("H*", "L*H", ">H*", "^[L*]H", "L*^[H]", "^[L*H]")
      ),
    fin_phon = str_replace_all(fin_phon, "L%]|%]", "%"),
    fin_phon = factor(fin_phon, levels=unique(fin_phon))
  ) %>%
  # Select only columns for analysis
  select(
    speaker,
    gender,
    mode,
    prompt,
    acc_phon,
    fin_phon,
    f0_exc) %>% 
  filter(acc_phon %in% l_star_h_types)  %>% 
  mutate(acc_phon = factor(acc_phon,
                           levels = l_star_h_types)
         )
```

```{r run initial model,  fig.height=3, fig.width=9, eval=FALSE, tidy = TRUE}

f0_exc_model <-
  lmer(
    f0_exc ~ mode * acc_phon + fin_phon + 
      (1 + acc_phon + mode | speaker) + (1 | gender) + (1 | prompt),

    data = m_corpus,
    control = lmerControl(
      optimizer = "optimx",
      calc.derivs = FALSE,
      optCtrl = list(
        method = "nlminb",
        starttests = FALSE,
        kkt = FALSE
      )
    )
  )


summarise_lme(f0_exc_model, run_step=TRUE) 

```

-  Convergence and singularity issues in initial model.
-  step() suggests dropping (1 | gender) + (1 | prompt).

```{r run step() model,  fig.height=3, fig.width=9, eval=FALSE, tidy = TRUE}

f0_exc_model <-
  lmer(
    f0_exc ~ mode * acc_phon + fin_phon + 
      (1 + acc_phon + mode | speaker),

    data = m_corpus,
    control = lmerControl(
      optimizer = "optimx",
      calc.derivs = FALSE,
      optCtrl = list(
        method = "nlminb",
        starttests = FALSE,
        kkt = FALSE
      )
    )
  )

summarise_lme(f0_exc_model, run_step=TRUE) 

```

-  Convergence and singularity issues remain.
-  reintroducing (1 | gender) + (1 | prompt) but using intercepts only model for speaker.

```{r intercepts-only model,  fig.height=3, fig.width=9, eval=FALSE, tidy = TRUE}

f0_exc_model <-
  lmer(
    f0_exc ~ mode * acc_phon + fin_phon + 
      (1 | speaker) + (1 | gender) + (1 | prompt),
    data = m_corpus,
    control = lmerControl(
      optimizer = "optimx",
      calc.derivs = FALSE,
      optCtrl = list(
        method = "nlminb",
        starttests = FALSE,
        kkt = FALSE
      )
    )
  )


summarise_lme(f0_exc_model, run_step=TRUE)
```
- singularity remains in random intercepts only model.
- Running model with only speaker and gender as random effect.
- Model works.
- Trimming model: removing observation with model residual of > 3 standard deviations

#### `f0_exc ~ mode * acc_phon + fin_phon + (1 | speaker) + (1 | gender)`
```{r reduced intercepts-only model,  fig.height=3, fig.width=9, tidy = TRUE}

f0_exc_model <-
  lmer(f0_exc ~ mode * acc_phon + fin_phon + (1 | speaker) + (1 | gender),
    data = m_corpus,
    control = lmerControl(
      optimizer = "optimx",
      calc.derivs = FALSE,
      optCtrl = list(
        method = "nlminb",
        starttests = FALSE,
        kkt = FALSE
      )
    )
  )

m_corpus.trimmed <- m_corpus %>% filter(abs(scale(resid(f0_exc_model))) <= 3)

f0_exc_model.trimmed <-
  lmer(f0_exc ~ mode * acc_phon + fin_phon + (1 | speaker) + (1 | gender),
    data = m_corpus.trimmed,
    control = lmerControl(
      optimizer = "optimx",
      calc.derivs = FALSE,
      optCtrl = list(
        method = "nlminb",
        starttests = FALSE,
        kkt = FALSE
      )
    )
  )


summarise_lme(f0_exc_model.trimmed, run_step=TRUE,
              write="../output_model_anovas/Mode_PA_f0_exc_anova.csv")
```

```{r make tidy model, fig.height=3, fig.width=9, ,  warning = FALSE, message = FALSE, tidy = TRUE}

f0_exc_model.tidy <- printTidyModel(f0_exc_model.trimmed,
                                write_r2 = "../output/Mode_PA_f0_exc"
                                )
f0_exc_model.tidy$r2

```

```{r do pairwise analysis, fig.height=3, fig.width=9, ,  warning = FALSE, message = FALSE, tidy = TRUE}

charts <- getModelFixedFX(
  my_equation =
    f0_exc ~ mode * acc_phon + fin_phon + (1 | speaker) + (1 | gender),
  my_data = m_corpus.trimmed,
  write="../output/Mode_PA_f0_exc"
)
charts$intercepts
charts$pairwise


```
