---
title: 'Chapter 07: Analysis of Function: Sentence Modes'
subtitle: '7.5.4 - 7.5.5 Utterance-wide Phonology, High Register, and Mode'
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup, include=F}
knitr::opts_chunk$set(warning = F, message = F, tidy = "styler")


# Load Functions and colour variables.
source("../../2_R_Functions/myFunctions.R")
source("../../2_R_Functions/myColours.R")

# Load packages.
installMissingPackages(
  c(
    # Load tidy packages.
    "formattable",
    "janitor",
    "kableExtra",
    "RColorBrewer",
    "tidyverse",
    "ggeffects",
    # Load stats package.
    "blme"
  )
)

# Set themes and colour schemes.
theme_set(theme_minimal(base_size = 10))

# Change this as required
options("speakr.praat.path" = "C:/Program Files/Praat/Praat.exe")

# Get data
m_corpus <- get_m_corpus("../../4_data/m_corpus.csv") %>%
  # Consolidate nuc_contour based on observation and types of register shift.
  mutate(
    # remove non-accentuation from phr_phon
    phr_phon = str_replace_all(phr_phon, "\\(\\*\\)", ""),
    `nuclear PA only` = factor(str_count(phr_phon, "\\*") == 1,
      levels = c(F, T)
    ),
    `high register in nucleus` = factor(str_detect(acc_phon, "\\^"),
      levels = c(F, T)
    ),
    `high reg. or nuc. PA only` = factor(
      `high register in nucleus` == T | `nuclear PA only` == T,
      levels = c(F, T)
    )
  ) %>%
  # Select only columns for analysis
  select(
    speaker,
    gender,
    mode,
    prompt,
    phr_phon,
    `nuclear PA only`,
    `high register in nucleus`,
    `high reg. or nuc. PA only`
  )


```

## Raw Data

```{r set thresholds}
#set lower limit for list: more than 8 tokens in the corpus
tot_rows = nrow(m_corpus)
threshold = round(tot_rows / 4 / 20,0)

keep_list <- m_corpus %>%
  mutate(phr_phon = factor(phr_phon, levels = unique(phr_phon))) %>%
  group_by(phr_phon) %>%
  summarise(phon_count = n()) %>% 
  filter(phon_count >= threshold)
```

### Utterance-wide Intonation {.tabset}

#### Mode

```{r Show distribution of utterance level phonology by mode, warning=F}
# List utterances phonology by mode with 8+ tokens in the corpus
m_corpus %>%
  mutate(phr_phon = factor(phr_phon, levels = unique(phr_phon))) %>%
  filter(phr_phon %in% keep_list$phr_phon) %>%
  group_by(mode, phr_phon) %>%
  summarise(phon_count = n(), .groups="rowwise") %>%
  arrange(desc(phon_count)) %>%
  mutate(phr_phon = str_replace_all(phr_phon,
                                    "([\\*\\[\\^\\>])", "\\\\\\1")) %>%
  pivot_wider(names_from = mode,
              values_from = phon_count,
              values_fill = 0) %>%
  select(phr_phon, MDC, MWH, MYN, MDQ) %>%
  rowwise(phr_phon) %>%
  mutate(total = sum(c_across())) %>%
  tibble::rowid_to_column("#") %>%
  rename("Phonology" = phr_phon) %>%
  formattable(
    align = "r",
    caption = paste(
      "Utterance level intonational phonology by mode ",
      "for tokens accounting for at least ",
      percent(threshold / tot_rows),
      " of all tokens.",
      sep = ""
    ),
    list(
      area(col = 3:6) ~ proportion_bar(color = c(
        rep(brewer.pal(4, "Pastel2")[3], nrow(.)),
        rep(brewer.pal(4, "Pastel2")[2], nrow(.)),
        rep(brewer.pal(4, "Pastel2")[1], nrow(.)),
        rep(brewer.pal(4, "Pastel2")[4], nrow(.))
        )),
      total = proportion_bar(color = "#aaaaaa")
    )
  )
```
```{r Prep. for utterance-level phonology, warning=F}

# get tibble of total counts per cell 
total_counts <- m_corpus %>%
  group_by(mode) %>%
  summarise(phon_count = n(), .groups="rowwise") %>%
  arrange(desc(phon_count)) %>%
  pivot_wider(names_from = mode,
              values_from = phon_count,
              values_fill = 0) %>% 
  mutate(total = sum(c_across())) %>%
  select(MDC, MWH, MYN, MDQ, total) 
```

#### All

```{r All utterance-level phonology, warning=F}
# List all utterance phonology
m_corpus %>%
  mutate(phr_phon = factor(phr_phon, levels = unique(phr_phon))) %>%
  group_by(phr_phon) %>%
  summarise(phon_count = n()) %>%
  arrange(desc(phon_count)) %>%
  mutate(
    phr_phon = str_replace_all(phr_phon, "([\\*\\[\\^\\>])", "\\\\\\1")
    ) %>%
  formattable(
    caption = "Utterance level intonational phonology in M-corpus ")



```

### Possible Q-Signalling Events {.tabset}

#### Nuclear PA-only
```{r Show nuc only utterance level phonology summary, warning=F}
nuc_only <- m_corpus %>%
  filter(`nuclear PA only` == T) %>%
  group_by(mode) %>%
  summarise(phon_count = n(), .groups="rowwise") %>%
  arrange(desc(phon_count)) %>%
  pivot_wider(names_from = mode,
              values_from = phon_count,
              values_fill = 0) %>% 
  mutate(total = sum(c_across())) %>%
  select(MDC, MWH, MYN, MDQ, total)

nuc_only %>% 
  rbind(round((nuc_only / total_counts*100),1) %>% paste("%", sep="")) %>% 
  mutate("info" =c("count", "percentage"), .before=1) %>%
  kable(
  caption = "Utterance level intonational phonology by mode for tokens with only one PA")%>% 
  kable_styling(full_width = F, position="left")
```

#### High register in nuclear PA
```{r Show `high register in nucleus` utterance level phonology summary, warning=F}
uses_h_reg <- m_corpus %>%
  filter(`high register in nucleus` == T) %>%
  group_by(mode) %>%
  summarise(phon_count = n(), .groups="rowwise") %>%
  complete(mode, fill=list(phon_count=0)) %>% 
  arrange(desc(phon_count)) %>%
  pivot_wider(names_from = mode,
              values_from = phon_count,
              values_fill = 0) %>% 
  mutate(total = sum(c_across())) %>%
  select(MDC, MWH, MYN, MDQ, total)

uses_h_reg %>% 
  rbind(round((uses_h_reg / total_counts*100),1) %>% paste("%", sep="")) %>% 
  mutate("info" =c("count", "percentage"), .before=1) %>%
  kable(
  caption = "High register in nuclear PA")%>% 
  kable_styling(full_width = F, position="left")
```

#### Either
```{r Show `high reg. or nuc. PA only` utterance level phonology summary, warning=F}
uses_h_reg_or_nuc_PA_only <- m_corpus %>%
  filter(`high register in nucleus` == T | `nuclear PA only` == T) %>%
  group_by(mode) %>%
  summarise(phon_count = n(), .groups="rowwise") %>%
  arrange(desc(phon_count)) %>%
  pivot_wider(names_from = mode,
              values_from = phon_count,
              values_fill = 0) %>% 
  mutate(total = sum(c_across())) %>%
  select(MDC, MWH, MYN, MDQ, total)

uses_h_reg_or_nuc_PA_only %>% 
  rbind(round((uses_h_reg_or_nuc_PA_only / total_counts*100),1) %>% paste("%", sep="")) %>% 
  mutate("info" = c("count", "percentage"), .before=1) %>%
  kable(
  caption = paste("Utterance level intonational phonology by mode with",
                  "high register or nuclear PA only"))%>% 
  kable_styling(full_width = F, position="left")
```


## BGLMM analysis

### Nuclear PA only  {.tabset}

```{r run nuc_pa_only_model, results = "hide"}

nuc_pa_only_model <- bglmer(
  `nuclear PA only` ~ mode + gender + (1 | speaker) + (1 | prompt),
  data = m_corpus,
  fixef.prior = normal(),
  family = binomial(link = "logit"),
  # change optimizer to avoid convergence errors
  control = glmerControl(
    optimizer = "optimx",
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)
h_reg_model <- optimizeModel(nuc_pa_only_model, reject_nm = F)

```

#### ANOVA (drop1())

```{r drop1 nuc_pa_only_model, fig.width = 3.05, fig.height = 2.5,}
write = "output/Utt_nuc_pa_only_GLMM"

model_formula <- getModelFormula(nuc_pa_only_model)

write(paste(str_replace_all(model_formula, "\\`", ""), sep = ""),
      paste(write, "_formula.txt", sep = ""))


# get, save, and print ANOVA of model
drop1_analysis <- drop1(nuc_pa_only_model, test = "Chisq")

tibble(factor = row.names(drop1_analysis))  %>%
  cbind(tibble(drop1_analysis)) %>%
  filter(!is.na(npar)) %>%
  mutate(`p.adj (BH)` = NA, signif. = NA) %>%
  write_csv(paste0(write, "_anova.csv"))


adjustP_posthoc("output/",
                `Pr(Chi)`,
                suffix_id = "_anova")

read_csv(paste(write, "_anova.csv", sep = "")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("Drop 1 ChiSq test of model:", model_formula))

```

#### *R*^2^ and prediction plots

```{r Tidy Summary nuc_pa_only_model, fig.width = 3.05, fig.height = 2.5, results="asis"}
tidy_summary <- analyseModel(nuc_pa_only_model,
                             write = write,
                             is_GLM = T,
                             type = "pred",
                             plot_rounding = 1,
                             short_caption = T)
```

#### prediction tables
```{r prediction nuc_pa_only_model, fig.width = 3.05, fig.height = 2.5, results="asis"}
printTidyPredictions(nuc_pa_only_model,
                     caption_suffix = "(reg tier)")
```

#### pairwise comparisons

```{r nuc_pa_only_model pairwise, fig.width = 3.05, fig.height = 2.5, results="asis"}
charts <- getModelFixedFX(
  nuc_pa_only_model,
  write = write,
  exponentiate = F
)
#charts$intercepts
charts$slopes %>% tidyStatNumbers()
```

#### Model printout

```{r nuc_pa_only_model printout}
summary(nuc_pa_only_model)
```

### Either Nuc-PA or High Reg {.tabset}

#### model summary

```{r h_reg_or_nuc_pa_only model}
h_reg_or_nuc_pa_only_model <- bglmer(
  `high reg. or nuc. PA only` ~ mode + gender +(1 | speaker) + (1 | prompt),
  data = m_corpus,
  family = binomial(link = "logit"),
  # change optimizer to avoid convergence errors
  control = glmerControl(
    optimizer = "optimx",
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

h_reg_model <- optimizeModel(h_reg_or_nuc_pa_only_model, reject_nm = F)

summary(h_reg_or_nuc_pa_only_model)
```

#### ANOVA (`drop1()`)

```{r drop1 h_reg_or_nuc_pa_only, fig.width = 3.05, fig.height = 2.5,}


write = "output/Utt_h_reg_or_nuc_pa_only_GLMM"

model_formula <- getModelFormula(h_reg_or_nuc_pa_only_model)

write(paste(str_replace_all(model_formula, "\\`", ""), sep = ""),
      paste(write, "_formula.txt", sep = ""))


# get, save, and print ANOVA of model
drop1_analysis <- drop1(h_reg_or_nuc_pa_only_model, test = "Chisq")

tibble(factor = row.names(drop1_analysis))  %>%
  cbind(tibble(drop1_analysis)) %>%
  filter(!is.na(npar)) %>%
  mutate(`p.adj (BH)` = NA, signif. = NA) %>%
  write_csv(paste0(write, "_anova.csv"))


adjustP_posthoc("output/",
                `Pr(Chi)`,
                suffix_id = "_anova")

read_csv(paste(write, "_anova.csv", sep = "")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("Drop 1 ChiSq test of model:", model_formula))

```

#### *R*^2^ and prediction plots

```{r Tidy Summary h_reg_or_nuc_pa_only, fig.width = 3.05, fig.height = 2.5, results="asis"}
tidy_summary <- analyseModel(h_reg_or_nuc_pa_only_model,
                             write = write,
                             is_GLM = T,
                             type = "pred",
                             plot_rounding = 1,
                             short_caption = T)
```

#### Prediction tables
```{r prediction h_reg_or_nuc_pa_onl, fig.width = 3.05, fig.height = 2.5, results="asis"}
printTidyPredictions(h_reg_or_nuc_pa_only_model,
                     caption_suffix = "(reg tier)")
```

#### Pairwise comparisons

```{r h_reg_or_nuc_pa_only_model pairwise, fig.width = 3.05, fig.height = 2.5, results="asis"}
charts <- getModelFixedFX(
  h_reg_or_nuc_pa_only_model,
  write = write,
  exponentiate = F
)
#charts$intercepts
charts$slopes %>% tidyStatNumbers()
```

#### Model printout
```{r h_reg_or_nuc_pa_only_model printout}
summary(h_reg_or_nuc_pa_only_model)
```


