---
title: "LME Analysis of NUCs"
author: "Antoin Rodgers"
date: "18.05.2022"
output:
  html_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F, 
                      message = F,
                      warning = F)

## Get personal functions and install / load necessary packages.
source("../2_R_Functions/myFunctions.R")

installMissingPackages(
  c(
    ## Include statistical packages.
    "performance",
    "lmerTest",
    "lme4",
    "ggeffects",
    "optimx",
    "MuMIn",
    "dfoptim",

    ## Include packages for tidy output.
    "tidyverse",
    "broomExtra",
    "sjPlot",
    "formattable",
    "knitr",
    "janitor",
    "gridExtra",
    
    ## Include packages for %in% %notin% syntactic notation
    "mefa4"
  )
)

## Set themes and colour schemes.
theme_set(theme_minimal(base_size = 12))

## Get  data for analysis

source("0_get_AH.R")

nuc_data <- nuc %>%
  ## there must be a more efficient way of doing this but...
  ## Create column of stressed syllables in foot 2.
    select(
    stim,
    speaker,
    gender,
    pre_syls,
    foot_syls,
    nuc_pre_text,
    nuc_new_word,
    nuc_str_syl,
    fin_phon,
    stim,
    v_onset_t,
    l_t,
    l_f0,
    h_t,
    h_f0,
    e_t,
    e_f0,
    f0_exc,
    lh_dur,
    lh_slope,
    foot_dur,
    foot_start_t,
    foot_end_t,
    h_t_foot_ratio,
    speech_rate
  )


optimizer = "optimx"

rm(nuc, pn, pn_ana, pn_foot, nuc_pre, nuc_foot, pn_lex, corpus)

```

```{r check levels in factors, results="asis", eval=F}

star_found = F
cat("\n\nFactors for nuc_data\n")
for (i in colnames(nuc_data)) {
  if(length(unique(nuc_data[[i]])) == 1){
    star = " *"
  star_found = T
  }
  else{
    star = ""
    }

  if (is.factor(nuc_data[[eval(i)]])) {
    cat("\n", i, star, "  \tunique = ", length(unique(nuc_data[[i]])),
        "  \tlevels = ", length(levels(nuc_data[[i]])), sep = "")
  }
    else {
    cat("\n", i, sep = "")
    }
}
cat("\n")
if(star_found){
  cat("\n* Factor only has one level in the actual data.\n")
}

nuc_data %>%
  group_by(pre_syls) %>%
  summarise(count = n()) %>% 
  adorn_totals("row") %>% 
  formattable(caption = "pre_syls stim variety")

nuc_data %>%
  group_by(nuc_new_word) %>%
  summarise(count = n()) %>%
  adorn_totals("row") %>% 
  formattable(caption = "nuc_new_word stim variety")
```

### L targets

#### Analysis of L timing

```{r run l_t models, fig.height=3, fig.width=9}
l_t_equation = formula(
  l_t ~ foot_syls
  + pre_syls
  + fin_phon
  + nuc_new_word
  + gender
  + (1 | speaker)
  + (1 | nuc_str_syl)
)


nuc_l_t_mdl = lmer(
  l_t_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

## model needs trimming
nuc_l_t_data.trimmed <- nuc_data %>% filter(abs(scale(resid(nuc_l_t_mdl))) <= 3)
nuc_l_t_mdl.trimmed = update(nuc_l_t_mdl, data = nuc_l_t_data.trimmed)
 mdl_summary <- summariseLME(nuc_l_t_mdl.trimmed,
             run_step = F,
             write = "5_nuc_output/nuc_l_t_anova.csv")
 
 adjustP_posthoc("5_nuc_output/",
                 p.value,
                 suffix_id = "_anova")
 
 read_csv("5_nuc_output/nuc_l_t_anova.csv") %>%
   tidyStatNumbers() %>%
   formattable(caption = paste("ANOVA of model:",
                               getModelFormula(nuc_l_t_mdl.trimmed)))
```

```{r l_t model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

re_plot <- plot_model(nuc_l_t_mdl, type = "re")[[2]]

re_plot <- re_plot +
  geom_hline(yintercept = 0, colour = "red") +
  labs(caption = "Random effect of stressed syllable on L timing") +
  ylab("time (ms)") +
  theme(
    plot.title = element_blank(),
    plot.caption = element_text(hjust = 0, size = 9),
    plot.caption.position = "plot"
  )
## re_plot
nuc_l_t_mdl.tidy <- analyseModel(nuc_l_t_mdl,
                                 write = "5_nuc_output/nuc_l_t",
                                 type = "pred",
                                 y_lab = "time (ms)",
                                 y_lim = c(-0, 650),
                                 breaks = seq(-0, 650, 100),
                                 plot_rounding = 0,
                                 caption_prefix = "a. "
                                 )

```

```{r do l_t pairwise analysis, fig.height=2.5, fig.width = 3.01, results = "asis"}
printTidyPredictions(nuc_l_t_mdl)

charts <- getModelFixedFX(nuc_l_t_mdl, write="5_nuc_output/nuc_l_t")

charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.),0)))
```

#### Analysis of L F0

```{r set l_f0 equation}
l_f0_equation = formula(
  l_f0 ~
    foot_syls
  + pre_syls
  + fin_phon
  + nuc_new_word
  + gender
  + (1 | speaker)
  + (1 | nuc_str_syl)
)
optimizer = "optimx"
```

```{r run l_f0 models, fig.height=3, fig.width=9}
nuc_l_f0_mdl = lmer(
  l_f0_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

## model needs trimming
nuc_l_f0.trimmed <-
  nuc_data %>% filter(abs(scale(resid(nuc_l_f0_mdl))) <= 2.5)

nuc_l_f0_mdl.trimmed = update(nuc_l_f0_mdl, data = nuc_l_f0.trimmed)

mdl_summary <- summariseLME(nuc_l_f0_mdl.trimmed,
                      run_step = F,
                      write = "5_nuc_output/nuc_l_f0_anova.csv")

adjustP_posthoc("5_nuc_output/",
                p.value,
                suffix_id = "_anova")

read_csv("5_nuc_output/nuc_l_f0_anova.csv") %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                              getModelFormula(nuc_l_f0_mdl.trimmed)))
```

```{r l_f0 model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

nuc_l_f0_mdl.tidy <- analyseModel(nuc_l_f0_mdl.trimmed,
                                  write = "5_nuc_output/nuc_l_f0",
                                  type = "pred",
                                  y_lab = "f0 (ST re median)",
                                  y_lim = c(-7, 4.5),
                                  breaks = seq(-7, 4.5, 2),
                                  plot_rounding = 1,
                                  caption_prefix = "b. "
                                  )
```

```{r do l_f0 pairwise analysis, fig.height=2.5, fig.width = 3.01, results = "asis"}
printTidyPredictions(nuc_l_f0_mdl.trimmed)

charts <- getModelFixedFX(nuc_l_f0_mdl.trimmed,
                write = "5_nuc_output/nuc_l_f0")

charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.), 1)))


```

```{r l binomial plots, fig.height=5.5/2.54, fig.width=15/2.54}
## set terms to plot
terms = c("fin_phon [L%]",
          "nuc_new_word [T]",
          "gender [M]")

## create _t plot
l_t_binom_plot <- plot_model(nuc_l_t_mdl,
           show.values = T,
           show.intercept = T,
           terms = terms) +
  geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylab("estimate (ms)") +
    labs(
         caption = paste0 ("a. Nuclear l_t model binomial factor effects.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
      axis.title = element_text(size = 10)
    )

## create _f0 plot
l_f0_binom_plot <- plot_model(nuc_l_f0_mdl.trimmed,
           show.values = T,
           terms = terms) +
  geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylab("estimate (ST)") +
    labs(title = "",
         caption = paste0 ("b. Nuclear l_f0 model binomial factor effects.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
      axis.title = element_text(size = 10)
    )

## join the plots
grid.arrange(l_t_binom_plot, l_f0_binom_plot, ncol = 2)

## save the plots
ggsave("5_nuc_output/l_target_binomial.png",
        device = "png",
        arrangeGrob(l_t_binom_plot, l_f0_binom_plot, ncol = 2),
        width = 15.5,
        height = 5.5,
        units = "cm")
```
### H targets

#### Analysis of H timing

```{r set h_t equation}
h_t_equation = formula(
  h_t ~
    foot_syls
  + pre_syls
  + fin_phon
  + nuc_new_word
  + gender
  + (1 | speaker)
  + (1 | nuc_str_syl)
)

optimizer = "optimx"
```

```{r run h_t models, fig.height=3, fig.width=9}
nuc_h_t_mdl = lmer(
  h_t_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

## model needs trimming
nuc_h_t_mdl = update(nuc_h_t_mdl,
                     data = nuc_data %>%
                       filter(abs(scale(resid(nuc_h_t_mdl))) <= 3))

mdl_summary <- summariseLME(nuc_h_t_mdl,
                      run_step = F,
                      write = "5_nuc_output/nuc_h_t_anova.csv")

adjustP_posthoc("5_nuc_output/",
                p.value,
                suffix_id = "_anova")

read_csv("5_nuc_output/nuc_h_t_anova.csv") %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                              getModelFormula(nuc_h_t_mdl)))
```


```{r h_t model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

nuc_h_t_mdl.tidy <- analyseModel(nuc_h_t_mdl,
                                write="5_nuc_output/nuc_h_t",
                                type = "pred",
                                y_lab = "time (ms)",
                                y_lim = c(0, 650),
                                breaks = seq(0, 650, 100),
                                plot_rounding = 0,
                                caption_prefix = "a. "
                                )


```

```{r do h_t pairwise analysis, fig.height=2.5, fig.width = 3.01, results = "asis"}
printTidyPredictions(nuc_h_t_mdl)

charts <- getModelFixedFX(nuc_h_t_mdl,
                write = "5_nuc_output/nuc_h_t")

charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.),0)))
```

#### Analysis of H F0

```{r set h_f0 equation}
h_f0_equation = formula(
  h_f0 ~ 
    foot_syls
  + pre_syls
  + fin_phon
  + nuc_new_word
  + gender
  + (1 | speaker)
  + (1 | nuc_str_syl)
)

optimizer = "optimx"
```

```{r run h_f0 models, fig.height=3, fig.width=9}
nuc_h_f0_mdl = lmer(
  h_f0_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

## model needs trimming
nuc_h_f0_mdl.trimmed = update(nuc_h_f0_mdl,
                              data = nuc_data %>%
                                filter(abs(scale(resid(nuc_h_f0_mdl))) <= 4))

mdl_summary <- summariseLME(nuc_h_f0_mdl.trimmed,
                      run_step = F,
                      write = "5_nuc_output/nuc_h_f0_anova.csv")

adjustP_posthoc("5_nuc_output/",
                p.value,
                suffix_id = "_anova")

read_csv("5_nuc_output/nuc_h_f0_anova.csv") %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                              getModelFormula(nuc_h_f0_mdl.trimmed)))
```

```{r h_f0 model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

nuc_h_f0_mdl.tidy <- analyseModel(nuc_h_f0_mdl.trimmed,
                                write="5_nuc_output/nuc_h_f0",
                                type = "pred",
                                y_lab = "f0 (ST re median)",
                                y_lim = c(-7, 4.5),
                                breaks = seq(-7, 4.5, 2),
                                plot_rounding = 1,
                                caption_prefix = "b. "
                                )

```

```{r do h_f0 pairwise analysis, fig.height=2.5, fig.width = 3.01, results = "asis"}
printTidyPredictions(nuc_h_f0_mdl.trimmed)
charts <- getModelFixedFX(nuc_h_f0_mdl.trimmed,
                          write = "5_nuc_output/nuc_h_f0")

charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.), 1)))
```

```{r h binomial plots, fig.height=5.5/2.54, fig.width=15/2.54}
## set terms to plot
terms = c("fin_phon [L%]",
          "nuc_new_word [T]",
          "gender [M]")

## create _t plot
h_t_binom_plot <- plot_model(nuc_h_t_mdl,
           type = "est",
           transform = NULL,
           show.values = T,
           vline = "Red",
           terms = terms) +
  geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylab("estimate (ms)") +
    labs(
         caption = paste0 ("a. Nuclear h_t model binomial factor effects.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
      axis.title = element_text(size = 10)
    )

## create _f0 plot
h_f0_binom_plot <- plot_model(nuc_h_f0_mdl.trimmed,
           type = "est",
           transform = NULL,
           show.values = T,
           vline = "Red",
           terms = terms) +
  geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylab("estimate (ST)") +
    labs(title = "",
         caption = paste0 ("b. Nuclear h_f0 model binomial factor effects.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
      axis.title = element_text(size = 10)
    )

## join the plots
grid.arrange(h_t_binom_plot, h_f0_binom_plot, ncol = 2)

## save the plots
ggsave("5_nuc_output/h_target_binomial.png",
        device = "png",
        arrangeGrob(h_t_binom_plot, h_f0_binom_plot, ncol = 2),
        width = 15.5,
        height = 5.5,
        units = "cm")
```

### E targets
e_t is not important for evaluation. However, the timing has been modelled just for visualisation purposes.
```{r set e_t equation}
#### Analysis of E timing
## E timing is not of much interest as it is just the last f0 time point in the IP... except in terms of seeing how h_t and e_t differ.

e_t_equation = formula(
  e_t ~ 
    foot_syls
  + pre_syls
  + fin_phon
  + nuc_new_word
  + gender
  + (1 | speaker)
  + (1 | nuc_str_syl)
)

optimizer = "optimx"
```

```{r run e_t models, fig.height=3, fig.width=9}
nuc_e_t_mdl = lmer(
  e_t_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

## model needs trimming
nuc_e_t_mdl.trimmed = update(nuc_e_t_mdl,
                             data = nuc_data %>%
                               filter(abs(scale(resid(
                                 nuc_e_t_mdl
                               ))) <= 4))

mdl_summary <- summariseLME(nuc_e_t_mdl.trimmed,
                            run_step = F,
                            write = "5_nuc_output/nuc_e_t_anova.csv")

adjustP_posthoc("5_nuc_output/",
                p.value,
                suffix_id = "_anova")

read_csv("5_nuc_output/nuc_e_t_anova.csv") %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                              getModelFormula(nuc_e_t_mdl.trimmed)))
```

```{r e_t model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

nuc_e_t_mdl.tidy <- analyseModel(nuc_e_t_mdl.trimmed,
                                write="5_nuc_output/nuc_e_t",
                                type = "pred",
                                y_lab = "time (ms)",
                                y_lim = c(0, 650),
                                breaks = seq(0, 650, 100),
                                plot_rounding = 0,
                                caption_prefix = "a. "
                                )

```

```{r do e_t pairwise analysis, fig.height=2.5, fig.width = 3.01, results="asis"}
printTidyPredictions(nuc_e_t_mdl.trimmed)

charts <- getModelFixedFX(
  nuc_e_t_mdl.trimmed,
  write="5_nuc_output/nuc_e_t"
)

charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.),0)))
```

#### Analysis of E F0

```{r set e_f0 equation}
e_f0_equation = formula(
  e_f0 ~ 
    foot_syls
  + pre_syls
  + fin_phon
  + nuc_new_word
  + gender
  + (1 | speaker)
  + (1 | nuc_str_syl)
)

optimizer = "optimx"
```

- model needs trimming (|sd| residuals > 3.5)
```{r run e_f0 models, fig.height=3, fig.width=9}
nuc_e_f0_mdl = lmer(
  e_f0_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

## model needs trimming
nuc_e_f0_mdl.trimmed = update(nuc_e_f0_mdl,
                              data = nuc_data %>% filter(abs(scale(
                                resid(nuc_e_f0_mdl)
                              )) <= 3.5))

mdl_summary <- summariseLME(nuc_e_f0_mdl.trimmed,
                            run_step = F,
                            write = "5_nuc_output/nuc_e_f0_anova.csv")

adjustP_posthoc("5_nuc_output/",
                p.value,
                suffix_id = "_anova")

read_csv("5_nuc_output/nuc_e_f0_anova.csv") %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                              getModelFormula(nuc_e_f0_mdl.trimmed)))
```


```{r e_f0 model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

nuc_e_f0_mdl.tidy <- analyseModel(nuc_e_f0_mdl.trimmed,
                                write="5_nuc_output/nuc_e_f0",
                                type = "pred",
                                y_lab = "f0 (ST re median)",
                                y_lim = c(-7, 4.5),
                                breaks = seq(-7, 4.5, 2),
                                plot_rounding = 1,
                                caption_prefix = "letters"
                                )

```

```{r do e_f0 pairwise analysis, fig.height=2.5, fig.width = 3.01, results="asis"}
printTidyPredictions(nuc_e_f0_mdl.trimmed)

charts <- getModelFixedFX(nuc_e_f0_mdl.trimmed,
                write="5_nuc_output/nuc_e_f0")

charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.), 1))) 
```

```{r e binomial plots, fig.height=5.5/2.54, fig.width=15/2.54}
## set terms to plot
terms = c("fin_phon [L%]",
          "nuc_new_word [T]",
          "gender [M]")

## create _t plot
e_t_binom_plot <- plot_model(nuc_e_t_mdl,
           type = "est",
           transform = NULL,
           show.values = T,
           vline = "Red",
           terms = terms) +
  geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylab("estimate (ms)") +
    labs(
         caption = paste0 ("a. Nuclear e_t model binomial factor effects.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
      axis.title = element_text(size = 10)
    )

## create _f0 plot
e_f0_binom_plot <- plot_model(nuc_e_f0_mdl.trimmed,
           type = "est",
           transform = NULL,
           show.values = T,
           vline = "Red",
           terms = terms) +
  geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylab("estimate (ST)") +
    labs(title = "",
         caption = paste0 ("b. Nuclear e_f0 model binomial factor effects.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
      axis.title = element_text(size = 10)
    )

## join the plots
grid.arrange(e_t_binom_plot, e_f0_binom_plot, ncol = 2)

## save the plots
ggsave("5_nuc_output/e_target_binomial.png",
        device = "png",
        arrangeGrob(e_t_binom_plot, e_f0_binom_plot, ncol = 2),
        width = 15.5,
        height = 5.5,
        units = "cm")
```
### Truncation and Compression Effects

#### LH Excursion

```{r set f0_exc equation}

f0_exc_equation = formula(
 f0_exc ~ 
   foot_syls + fin_phon
 + (1 + foot_syls | speaker)
 + (1 | nuc_str_syl)
 + (1 | pre_syls)
)

optimizer = "optimx"
```

- model needs trimming (|sd| residuals > 2.5)
```{r run f0_exc models, fig.height=3, fig.width=9}
nuc_f0_exc_mdl = lmer(
  f0_exc_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

#nuc_f0_exc_mdl <- optimizeModel(nuc_f0_exc_mdl)
mdl_summary <- summariseLME(nuc_f0_exc_mdl,
             run_step = F, print_summary = F)


nuc_f0_exc_mdl.trimmed = lmer(
  f0_exc_equation,
  data = nuc_data %>% filter(abs(scale(
    resid(nuc_f0_exc_mdl)
  )) <= 2.5),
  control = lmerControl(
    optimizer = "nloptwrap",
    optCtrl = list(
      algorithm = "NLOPT_LN_COBYLA",
      maxfun = 1e9,
      maxeval = 1e7,
      xtol_abs = 1e-9,
      ftol_abs = 1e-9
    )
  )
)
 nuc_f0_exc_mdl.trimmed <- optimizeModel(nuc_f0_exc_mdl.trimmed)                               

mdl_summary <- summariseLME(nuc_f0_exc_mdl.trimmed,
                            run_step = F,
                            write = "5_nuc_output/nuc_f0_exc_anova.csv")

adjustP_posthoc("5_nuc_output/",
                p.value,
                suffix_id = "_anova")

read_csv("5_nuc_output/nuc_f0_exc_anova.csv") %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                              getModelFormula(nuc_f0_exc_mdl.trimmed)))
```
```{r plot f0_exc estimated effects, fig.height=2.5, fig.width=4.3}
a <- plot_model(nuc_f0_exc_mdl.trimmed,
                type = "est",
                transform = NULL,
                show.values = T) +
  
  labs(caption = "b. Estimated effects on f0_exc.") +
  
  theme(title = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot") +
  
  geom_hline(yintercept = 0, size = 0.5, color = "red")
```

```{r f0_exc model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

nuc_f0_exc_mdl.tidy <- analyseModel(nuc_f0_exc_mdl.trimmed,
                                write="5_nuc_output/nuc_f0_exc",
                                type = "pred",
                                y_lab = "f0 (ST)",
                                caption_prefix = "a. ",
                                short_caption = T)

```

```{r do f0_exc pairwise analysis, fig.height=2.5, fig.width = 3.01, results="asis"}
printTidyPredictions(nuc_f0_exc_mdl.trimmed)

charts <- getModelFixedFX(nuc_f0_exc_mdl.trimmed,
                          write = "5_nuc_output/nuc_f0_exc")

charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(any_of(c("estimate", "conf.low", "conf.high")),
                ~ round(as.numeric(.), 1)))


```


#### LH Slope

Based on results from tonal target analysis, it looksl like compression and truncation may be in effect.
As compression leads to decrease in duration of rise (but not necessarily much descrease in excursion size), and as slope increases exponentially as duration decreases, slope has been logged here.

- model works once random intercepts of speech_rate and foot_dur removed
```{r set lh_slope equation}
nuc_data <- nuc_data %>% 
    mutate(log_lh_slope = log(lh_slope))

lh_slope_equation = formula(
  log_lh_slope ~
    foot_syls + fin_phon
  + (1 + foot_syls | speaker)
  + (1 | nuc_str_syl)
  + (1 | pre_syls)
  )
optimizer = "optimx"
```

- model needs trimming (|sd| residuals > 3.5)
```{r run lh_slope models, fig.height=3, fig.width=9}
nuc_lh_slope_mdl = lmer(lh_slope_equation,
                        nuc_data,
                        control = lmerControl(optimizer = optimizer,
                                              calc.derivs = F,
                                              optCtrl = list(method = "nlminb",
                                                             starttests = F,
                                                             kkt = F)
                                              )
                        )

nuc_lh_slope_mdl <- optimizeModel(nuc_lh_slope_mdl, verbose = F, reject_nm = T)

## model needs trimming
nuc_lh_slope_mdl.trimmed = update(
  nuc_lh_slope_mdl,
  data = nuc_data %>% filter(abs(scale(resid(nuc_lh_slope_mdl))) <= 3.5))



nuc_lh_slope_mdl.trimmed <-
  optimizeModel(nuc_lh_slope_mdl.trimmed,
                verbose = F,
                reject_nm = T)

mdl_summary <- summariseLME(nuc_lh_slope_mdl.trimmed,
                            run_step = F,
                            write = "5_nuc_output/nuc_lh_slope_anova.csv")


adjustP_posthoc("5_nuc_output/",
                p.value,
                suffix_id = "_anova")

read_csv("5_nuc_output/nuc_lh_slope_anova.csv", show_col_types = FALSE) %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                              getModelFormula(nuc_lh_slope_mdl.trimmed)))
```

```{r plot lh slope estimated effects, fig.height=2.5, fig.width=4.3}
b <- plot_model(nuc_lh_slope_mdl.trimmed,
                type = "est",
                transform = NULL,
                show.values = T) +
  
  labs(caption = "b. Estimated effects on log_lh_slope.") +

    theme(title=element_blank(),
        plot.caption=element_text(hjust = 0, size = 10),
        plot.caption.position= "plot") +
  
  geom_hline(yintercept = 0, size = 0.5, color = "red")
```

```{r lh_slope model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

nuc_lh_slope_mdl.tidy <- analyseModel(nuc_lh_slope_mdl.trimmed,
                                write="5_nuc_output/nuc_lh_slope",
                                type = "pred",
                                caption_prefix = "b. ",
                                y_lab = "log(ST/sec)",
                                plot_rounding = 2,
                                short_caption = T)
```

```{r do lh_slope pairwise analysis, results="asis"}
printTidyPredictions(nuc_lh_slope_mdl.trimmed)

charts <- getModelFixedFX(nuc_lh_slope_mdl.trimmed,
                          write = "5_nuc_output/nuc_lh_slope")

charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.), 2)))


```

```{r compare effects exc and log_slope effects, fig.height=2.5, fig.width=6}
grid.arrange(a, b, ncol = 2) 
ggsave("5_nuc_output/nuc_truncation_params.png",
       device = "png",
       arrangeGrob(a, b, ncol = 2),
       width = 15.7,
       height = 6.5,
       units = "cm")
```
```{r lh_dur model, fig.height=3, fig.width=9}
lh_dur_mdl <-
  lmer(lh_dur ~ foot_syls + fin_phon + (1 + foot_syls |
                               speaker), data = nuc_data) %>%
  optimizeModel()

lh_dur_mdl <- update(lh_dur_mdl, data = nuc_data %>% 
                       filter(abs(scale(resid(lh_dur_mdl))) <= 3))
mdl_summary <- summariseLME(lh_dur_mdl,
                            run_step = F,
                            write = "5_nuc_output/lh_dur_mdl_anova.csv")

adjustP_posthoc("5_nuc_output/",
                p.value,
                suffix_id = "_anova")

read_csv("5_nuc_output/lh_dur_mdl_anova.csv") %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                              getModelFormula(lh_dur_mdl)))
```


```{r lh_dur model analysis, fig.height=2.5, fig.width=2.823, results="asis"}

mdl_summary$omega2
printTidyPredictions(lh_dur_mdl)
```


### Additional model

#### H time proportional to duration of foot


```{r  h_t_re_ft_dur, fig.height=3, fig.width=9}
#### Analysis of E timing
## E timing is not of much interest as it is just the last f0 time point in the IP

h_t_re_ft_dur_equation = formula(
  `h_t as ratio of foot` ~ 
    foot_syls
  + (1 | speaker)
  + (1 | gender)
  + (1 | pre_syls)
  + (1 | nuc_new_word)
  + (1 | nuc_str_syl)
  + (1 | fin_phon)
)

optimizer = "optimx"

nuc_data <- nuc_data %>%
  mutate(`h_t as ratio of foot` = h_t_foot_ratio)

nuc_h_t_re_ft_dur_mdl = lmer(
  h_t_re_ft_dur_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

nuc_h_t_re_ft_dur_mdl <- nuc_h_t_re_ft_dur_mdl %>% optimizeModel()


mdl_summary <- summariseLME(nuc_h_t_re_ft_dur_mdl,
                            run_step = F,
                            write = "5_nuc_output/nuc_h_t_re_ft_dur_anova.csv")

adjustP_posthoc("5_nuc_output/",
                p.value,
                suffix_id = "_anova")

read_csv("5_nuc_output/nuc_h_t_re_ft_dur_anova.csv") %>%
  tidyStatNumbers() %>%
  formattable(caption = paste(
    "ANOVA of model:",
    getModelFormula(nuc_h_t_re_ft_dur_mdl)
  ))
```

```{r h_t_re_ft_dur model analysis, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

nuc_h_t_re_ft_dur_mdl.tidy <- analyseModel(nuc_h_t_re_ft_dur_mdl,
                                write="5_nuc_output/nuc_h_t_re_ft_dur",
                                type = "pred",
                                y_lab = "time (proportion)",
                                y_lim = c(0, 1.2),
                                breaks = seq(0, 1.2, 0.2),
                                plot_rounding = 2,
                                caption_prefix = "a. ",
                                short_caption = T
                                )

printTidyPredictions(nuc_h_t_re_ft_dur_mdl)

charts <- getModelFixedFX(
  nuc_h_t_re_ft_dur_mdl,
  write="5_nuc_output/nuc_h_t_re_ft_dur_mdl"
)

charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.),2)))
```

#### H time proportional to duration of voicing

```{r  nuc_h_t_re_v_dur, fig.height=3, fig.width=9}
#### Analysis of E timing
## E timing is not of much interest as it is just the last f0 time point in the IP

nuc_h_t_re_v_dur_equation = formula(
  `h_t as ratio of voicing` ~ 
    foot_syls
  + (1 | speaker)
  + (1 | gender)
  + (1 | pre_syls)
  + (1 | nuc_new_word)
  + (1 | nuc_str_syl)
  + (1 | fin_phon)
)

optimizer = "optimx"

nuc_data <- nuc_data %>%
  mutate(`h_t as ratio of voicing` = h_t  / e_t)

nuc_h_t_re_v_dur_mdl = lmer(
  nuc_h_t_re_v_dur_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)
nuc_h_t_re_v_dur_mdl <- nuc_h_t_re_v_dur_mdl %>% optimizeModel()
 mdl_summary <- summariseLME(nuc_h_t_re_v_dur_mdl,
             run_step = T,
             print_summary = F
             )
 
nuc_nuc_h_t_re_v_dur.trimmed <- nuc_data %>%
    filter(abs(scale(resid(nuc_h_t_re_v_dur_mdl))) <= 3)

## model needs trimming
nuc_h_t_re_v_dur_mdl.trimmed = update(
  nuc_h_t_re_v_dur_mdl,
  data = nuc_nuc_h_t_re_v_dur.trimmed,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

nuc_h_t_re_v_dur_mdl.trimmed <-
  nuc_h_t_re_v_dur_mdl.trimmed %>%
  optimizeModel()

mdl_summary <- summariseLME(nuc_h_t_re_v_dur_mdl.trimmed,
                            run_step = F,
                            write = "5_nuc_output/nuc_nuc_h_t_re_v_dur_anova.csv")

adjustP_posthoc("5_nuc_output/",
                p.value,
                suffix_id = "_anova")

read_csv("5_nuc_output/nuc_nuc_h_t_re_v_dur_anova.csv") %>%
  tidyStatNumbers() %>%
  formattable(caption = paste(
    "ANOVA of model:",
    getModelFormula(nuc_h_t_re_v_dur_mdl.trimmed)
  ))
```

```{r nuc_h_t_re_v_dur model analysis, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

nuc_h_t_re_v_dur_mdl.tidy <- analyseModel(nuc_h_t_re_v_dur_mdl.trimmed,
                                write="5_nuc_output/nuc_nuc_h_t_re_v_dur",
                                type = "pred",
                                y_lab = "time (proportion)",
                                y_lim = c(0, 1.2),
                                breaks = seq(0, 1.2, 0.2),
                                plot_rounding = 2,
                                caption_prefix = "b. ",
                                short_caption = T
                                )

printTidyPredictions(nuc_h_t_re_v_dur_mdl.trimmed)


charts <- getModelFixedFX(nuc_h_t_re_v_dur_mdl.trimmed,
                          write = "5_nuc_output/nuc_h_t_re_v_dur_mdl")

charts$slopes %>% tidyStatNumbers() %>%
  mutate(across(any_of(c("estimate", "conf.low", "conf.high")),
  ~ round(as.numeric(.), 2)))
```

