---
title: 'Chapter 07: Analysis of Function: Sentence Modes'
subtitle: '7.X f0 Density Distributions'
author: "Antoin Rodgers"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup mode only}
#| results="hide",
#| message=F,
#| warning=F

knitr::opts_chunk$set(warning = F, message = F, tidy = "styler")

## Load Functions.
source("../../2_R_Functions/myFunctions.R") 

## Load packages.
installMissingPackages(
  c(
    ## Include statistical packages.
    "performance",
    "lmerTest",
    "lme4",
    "optimx",
    "MuMIn",
    ## Include packages for tidy output.
    "tidyverse",
    "broomExtra",
    "sjPlot",
    "formattable",
    "knitr",
    "RColorBrewer",
    "janitor",
    ## Include packages for %in% %notin% syntactic notation
    "mefa4"
  )
)

y_lim_t = c(0, 360)
y_lim_f0 = c(-4.1, 10.5)

```


```{r load corpus}

## Load Corpus
m_corpus <- get_m_corpus("../../4_data/m_corpus.csv")  %>%
  ## Select only columns for analysis
  select(
    speaker,
    mode,
    stim,
    h_f0,
    h_f0_z,
    utt_mean_f0,
    utt_mean_f0_z
    )
```

## Check balance across repetitions

```{r Check data for balance}
m_corpus %>%
  group_by(speaker, stim) %>%
  summarise(count = n(), .groups = "keep") %>% pivot_wider(values_from = count,
                                         names_from = speaker,
                                         values_fill = 1)  %>% 
  adorn_totals(where = "col", fill = "grey") %>%
  adorn_totals(where = "row", fill = "grey") %>%
  formattable(caption = "Tokens per stimulus per speaker in M-Corpus")

```

While there are some missing tokens per speaker and target phrse, most noticeably for M9, I don't think that using the raw data will lead to a misrepresentation of the data in the dotplots and density distribution graphs.

## Means and standard deviations (z-scores) {.tabset}

### Nuclear peak f0 scaling
```{r Get means and SD by ST}

m_corpus %>%
  group_by(mode) %>%
  summarise(h_f0_z = mean(h_f0_z)) %>% 
  arrange(mode) %>%
  mutate(h_f0_z =  round(h_f0_z, 2)) %>% 
  formattable(caption = "Average nuclear peak by mode (z-score)")

m_corpus %>%
  group_by(mode) %>%
  summarise(h_f0_z = sd(h_f0_z)) %>% 
  arrange(mode) %>%
  mutate(h_f0_z =  round(h_f0_z, 2)) %>% 
  formattable(caption = "Standard deviation of nuclear peak by mode (ST)")
```

### Utterance mean f0
```{r Get means by s-score}

m_corpus %>%
  group_by(mode) %>%
  summarise(utt_mean_f0_z = mean(utt_mean_f0_z)) %>% 
  arrange(mode) %>%
  mutate(utt_mean_f0_z =  round(utt_mean_f0_z, 2)) %>% 
  formattable(caption = "Utterance mean f0 by mode (z-score)")

m_corpus %>%
  group_by(mode) %>%
  summarise(utt_mean_f0_z = sd(utt_mean_f0_z)) %>% 
  arrange(mode) %>%
  mutate(utt_mean_f0_z =  round(utt_mean_f0_z, 2)) %>% 
  formattable(caption = "Utterance mean f0 by mode (z-score)")
```
## Graphs {.tabset}

### Nuclear Peak f0 Density Distribution

```{r Nuclear Peak f0 Distribution, fig.width = 6, fig.height = 4}
#| results="hide"

my_palette <- brewer.pal(4, "Dark2")

ggplot(
  data = m_corpus,
  mapping = aes(x = h_f0_z, fill = mode)
  ) +
  geom_dotplot(dotsize = 1, stackgroups = TRUE, binpositions = "all", stackratio = 0.185) +
  geom_density(stat = "density", alpha = 0.4, fill = "grey") +
  scale_fill_manual(
    values = my_palette,
    name = NULL
    ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = c(.9, .7),
    legend.margin = margin(-6, 6, -6, 6),
        legend.background = element_rect(),
    legend.key.size = unit(1, "cm"),
    panel.grid.major.x = element_line(color = "grey",
                                      size = 0.5),
    panel.grid.minor.x = element_blank()) +
  xlab("f0 (z-score)") +
  ylab("density") +
  ylim(0, 0.5) +
  labs(title = "Density Distribution of M-Corups nuclear peak f0 Scaling")


```


### Utterance Mean f0 Density Distribution

```{r Utterance Mean f0 Density Distribution, fig.width = 6, fig.height = 4}
#| results="hide"

my_palette <- brewer.pal(4, "Dark2")

ggplot(
  data = m_corpus,
  mapping = aes(x = utt_mean_f0_z, fill = mode)
  ) +
  geom_dotplot(dotsize = 1, stackgroups = TRUE, binpositions = "all", stackratio = 0.22) +
  geom_density(stat = "density", alpha = 0.4, fill = "grey") +
  scale_fill_manual(
    values = my_palette,
    name = NULL
    ) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = c(.9, .7),
    legend.margin = margin(-6, 6, -6, 6),
    legend.background = element_rect(),
    legend.key.size = unit(1, "cm"),
    panel.grid.major.x = element_line(color = "grey",
                                      size = 0.5),
    panel.grid.minor.x = element_blank()) +
  xlab("f0 (z-score)") +
  ylab("density") +
  ylim(0, 0.85) +
  labs(title = "Density Distribution of M-Corpus utterance mean f0")


```
