---
title: "LME Analysis of PNs"
author: "Antoin Rodgers"
date: "18.05.2022"
output:
  html_document: default
---

```{r setup, include=F}
knitr::opts_chunk$set(echo = F,
                      message = F,
                      warning = F)

# Get personal functions and install / load necessary packages.
source("../2_R_Functions/myFunctions.R")

installMissingPackages(
  c(
    # Include statistical packages.
    "performance",
    "lmerTest",
    "lme4",
    "ggeffects",
    "optimx",
    "MuMIn",
    "dfoptim",
    "effectsize",

    # Include packages for tidy output.
    "tidyverse",
    "broomExtra",
    "sjPlot",
    "formattable",
    "knitr",
    "gridExtra",

    # Include packages for %in% %notin% syntactic notation
    "mefa4"
  )
)

# Set themes and colour schemes.
theme_set(theme_minimal(base_size = 12))

# Get  data for analysis

source("0_get_AH.R")

h_star_type <-  c("L*H", ">H*", "H*")
l_star_type <-  c("L*H", "L*")

pn <- pn %>%
  select(speaker,
         gender,
         stim,
         ana_syls,
         foot_syls,
         wrd_end_syl,
         ana_text,
         pn_str_syl,
         pn_new_word,
         nuc_pre_text,
         acc_phon,
         speech_rate,
         l_t,
         l_f0,
         h_t,
         h_f0,
         lh_dur,
         lh_slope,
         f0_exc) %>%
  mutate(
    stim = factor(stim, levels = unique(stim)),
    gender = factor(gender, levels = c("F", "M")),
    foot_syls = factor(foot_syls, levels = 1:4)
  ) 

pn_l_data <- pn %>%
  filter(acc_phon %in% l_star_type) %>%
  mutate(acc_phon = factor(acc_phon, levels = l_star_type))%>%
  select(-c(h_t, h_f0))

pn_h_data <- pn %>%
  filter(acc_phon %in% h_star_type) %>%
  mutate(acc_phon = factor(acc_phon, levels = h_star_type)) %>%
  select(-c(l_t, l_f0))

#get data ready for T and C analysis
pn_tnc_data <- pn_l_data %>%
  mutate(log_lh_slope = log(lh_slope)) %>%
  filter(acc_phon == "L*H")

rm(nuc, nuc_foot, pn_lex, pn_ana, pn_foot, corpus, nuc_pre)

optimizer = "optimx"

```

```{r check levels in factors, results="asis", eval=F}

star_found = F
cat("Factors for pn_l_data\n")
for (i in colnames(pn_l_data)) {
  if(length(unique(pn_l_data[[i]])) == 1){
    star =" *"
  star_found = T
  }
  else{
    star = ""
    }
  if (is.factor(pn_l_data[[eval(i)]])) {
    cat("\n", i, star,"  \tunique = ", length(unique(pn_l_data[[i]])),
        "  \tlevels = ", length(levels(pn_l_data[[i]])), sep = "")
  }
  else {
    cat("\n", i, sep = "")
    }
}
cat("\n")
if(star_found){
  cat("\n* Factor only has one level in the actual data.\n")
}
cat("\nacc_phon counts\n")
summary(pn_l_data$acc_phon)
star_found = F
cat("\n\nFactors for pn_h_data\n")
for (i in colnames(pn_h_data)) {
  if(length(unique(pn_h_data[[i]])) == 1){
    star = " *"
  star_found = T
  }
  else{
    star = ""
    }

  if (is.factor(pn_h_data[[eval(i)]])) {
    cat("\n", i, star, "  \tunique = ", length(unique(pn_h_data[[i]])),
        "  \tlevels = ", length(levels(pn_h_data[[i]])), sep = "")
  }
    else {
    cat("\n", i, sep = "")
    }
}
cat("\n")
if(star_found){
  cat("\n* Factor only has one level in the actual data.\n")
}
cat("\nacc_phon counts\n")
summary(pn_h_data$acc_phon)

pn_l_data <- pn %>%
  filter(acc_phon %in% l_star_type) %>%
  mutate(acc_phon = factor(acc_phon, levels = l_star_type))%>%
  select(-c(h_t, h_f0))


```

## L targets

### Analysis of L timing

```{r run l_t models, fig.height=3, fig.width=9}
l_t_equation = formula(
  l_t ~
    acc_phon
  + ana_syls
  + foot_syls
  + wrd_end_syl
  + pn_new_word
  + gender
  + (1 | speaker)
  + (1 | pn_str_syl)
)


pn_l_t_mdl = lmer(
  l_t_equation,
  pn_l_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

#pn_l_t_mdl <- optimizeModel(pn_l_t_mdl, reject_nm = T, verbose = F)
mdl_summary <- summariseLME(pn_l_t_mdl, run_step = F, print_summary = F)


sd_threshold =2.5
cat("\nTrimming data with residual sd > ", sd_threshold, ".\n", sep ="")
pn_l_t_data.trimmed <-
  pn_l_data %>% filter(abs(scale(resid(pn_l_t_mdl))) <= sd_threshold)
pn_l_t_mdl.trimmed <- lmer(
  l_t_equation,
  pn_l_t_data.trimmed,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

pn_l_t_mdl.trimmed <- optimizeModel(pn_l_t_mdl.trimmed, reject_nm = T, verbose = F)


tibble(`Summary of model` = getModelFormula(pn_l_t_mdl.trimmed)) %>%
          formattable(align = "l")
```


```{r summarise l_t model, fig.height=3, fig.width=9}
mdl_summary <- summariseLME(
  pn_l_t_mdl.trimmed,
  run_step = F
  , write = "4_pn_output/pn_l_t_anova.csv"
  )

adjustP_posthoc("4_pn_output/",
                p.value,
                suffix_id = "_anova")


read_csv("4_pn_output/pn_l_t_anova.csv") %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(pn_l_t_mdl.trimmed)))

```
```{r l_t model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

pn_l_t_mdl.tidy <- analyseModel(
  pn_l_t_mdl.trimmed,
  write="4_pn_output/pn_l_t",
  type = "pred",
  y_lab = "time (ms)",
  y_lim = c(-80, 180),
  breaks = seq(-80, 180, 40),
  plot_rounding = 0,
  panel_prefix = "a. "
  )



```
```{r do l_t pairwise analysis, fig.height=2.5, fig.width=2.823, results = "asis"}
printTidyPredictions(pn_l_t_mdl.trimmed)

charts <- getModelFixedFX(pn_l_t_mdl.trimmed, write="4_pn_output/pn_l_t")
charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.),0)))
```


### Analysis of L F0

```{r run l_f0 models, fig.height=3, fig.width=9}
l_f0_equation = formula(
  l_f0 ~
    acc_phon
  + ana_syls
  + foot_syls
  + wrd_end_syl
  + pn_new_word
  + gender
  + (1 + foot_syls | speaker)
  )

pn_l_f0_mdl = lmer(
  l_f0_equation,
  pn_l_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

pn_l_f0_mdl <- optimizeModel(pn_l_f0_mdl, reject_nm = T, verbose = F)
mdl_summary <- summariseLME(pn_l_f0_mdl, print_summary = F, plot_resids = F, run_step = F)

l_f0_equation = formula(
  l_f0 ~
    acc_phon
  + ana_syls
  + foot_syls
  + wrd_end_syl
  + pn_new_word
  + gender
  + (1 | speaker)
 )

pn_l_f0_mdl <- update(pn_l_f0_mdl, formula = l_f0_equation)
pn_l_f0_mdl <- optimizeModel(pn_l_f0_mdl, reject_nm = T, verbose = F)
mdl_summary <- summariseLME(pn_l_f0_mdl, print_summary = F, run_step = F)

# model needs trimming
pn_l_f0.trimmed <- pn_l_data %>%
  filter(abs(scale(resid(pn_l_f0_mdl))) <= 3.5)

pn_l_f0_mdl <- update(pn_l_f0_mdl, data = pn_l_f0.trimmed)
pn_l_f0_mdl <- optimizeModel(pn_l_f0_mdl, reject_nm = T, verbose = F)

tibble(`Summary of model` = getModelFormula(pn_l_f0_mdl)) %>%
          formattable(align = "l")

mdl_summary <- summariseLME(pn_l_f0_mdl,
                  run_step = F,
                  write = "4_pn_output/pn_l_f0_anova.csv")

adjustP_posthoc("4_pn_output/",
                p.value,
                suffix_id = "_anova")

read_csv("4_pn_output/pn_l_f0_anova.csv") %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                            getModelFormula(pn_l_f0_mdl)))

```

```{r l_f0 model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

pn_l_f0_mdl.tidy <- analyseModel(
  pn_l_f0_mdl,
  write = "4_pn_output/pn_l_f0",
  type = "pred",
  y_lab = "f0 (ST re median)",
  y_lim = c(-3, 1.5),
  breaks = seq(-3, 1.5, 1),
  plot_rounding = 1,
  panel_prefix = "b. "
  )
```
```{r do l_f0 pairwise analysis, fig.height=2.5, fig.width=2.823, results = "asis"}
printTidyPredictions(pn_l_f0_mdl)

charts <- getModelFixedFX(pn_l_f0_mdl, write = "4_pn_output/pn_l_f0")
charts$slopes %>% tidyStatNumbers()%>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.),1)))
```


```{r l binomial plots, fig.height=5.5/2.54, fig.width=15/2.54}
# set terms to plot
terms = c("acc_phon [L*]",
          "pn_new_word [T]",
          "gender [M]")

# create _t plot
l_t_binom_plot <- plot_model(pn_l_t_mdl.trimmed,
           type = "est",
           transform = NULL,
           show.values = T,
           vline = "Red",
           terms = terms) +
  geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylab("estimate (ms)") +
    labs(
         caption = paste0 ("a. PN l_t model binomial factor effects.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
      axis.title = element_text(size = 10)
    )

# create _f0 plot
l_f0_binom_plot <- plot_model(pn_l_f0_mdl,
           type = "est",
           transform = NULL,
           show.values = T,
           vline = "Red",
           terms = terms) +
  geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylab("estimate (ST)") +
    labs(title = "",
         caption = paste0 ("b. PN l_f0 model binomial factor effects.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
      axis.title = element_text(size = 10)
    )

# join the plots
grid.arrange(l_t_binom_plot, l_f0_binom_plot, ncol = 2)

# save the plots
ggsave("4_pn_output/l_target_binomial.png",
        device = "png",
        arrangeGrob(l_t_binom_plot, l_f0_binom_plot, ncol = 2),
        width = 15.5,
        height = 5.5,
        units = "cm")
```

## H targets

### Analysis of H timing

```{r run h_t models, fig.height=3, fig.width=9}
h_t_equation = formula(
  h_t ~
    acc_phon
  + ana_syls
  + foot_syls
  + wrd_end_syl
  + pn_new_word
  + gender
  + (1 + foot_syls | speaker)
  + (1 | pn_str_syl)
)


pn_h_t_mdl = lmer(
  h_t_equation,
  pn_h_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

pn_h_t_mdl <- optimizeModel(pn_h_t_mdl, reject_nm = T)
mdl_summary <- summariseLME(pn_h_t_mdl,
                  run_step = F, print_summary = F
                  )

# model needs trimming
pn_h_t.trimmed <- pn_h_data %>% filter(abs(scale(resid(pn_h_t_mdl))) <= 2.5)
pn_h_t_mdl.trimmed = update(pn_h_t_mdl, data = pn_h_t.trimmed)
#pn_h_t_mdl.trimmed <- optimizeModel(pn_h_t_mdl.trimmed, reject_nm = T, verbose = F)
pn_h_t_mdl.trimmed = update(pn_h_t_mdl.trimmed,
                            control = lmerControl(
                              optimizer = "nloptwrap",
                              optCtrl = list(
                                algorithm = "NLOPT_LN_PRAXIS",
                                maxfun = 1e9,
                                maxeval = 1e7,
                                xtol_abs = 1e-9,
                                ftol_abs = 1e-9
                              )
                            ))

tibble(`Summary of model` = getModelFormula(pn_h_t_mdl.trimmed)) %>%
          formattable(align = "l")

mdl_summary <- summariseLME(pn_h_t_mdl.trimmed,
                  run_step = F,
                  write = "4_pn_output/pn_h_t_anova.csv"
  )

adjustP_posthoc("4_pn_output/",
                p.value,
                suffix_id = "_anova")

read_csv("4_pn_output/pn_h_t_anova.csv") %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(pn_h_t_mdl.trimmed)))
```

```{r plot h_t estimated effects, fig.height=1.5, fig.width=4, eval=F}

my_terms = c("acc_phon [>H*, H*]",
             "ana_syls [1, 2, 3]",
             "foot_syls [2, 3, 4]",
             "wrd_end_syl [2, 3]",
             "gender [M]")


for(cur_term in my_terms) {
  cur_factor <-cur_term %>% str_split("\\s", simplify = T)
  cur_factor <- cur_factor[1]
  my_plot <-
    plot_model(pn_h_t_mdl.trimmed,
               show.values = T,
               terms = cur_term) +
    geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylim(-100, 100) +
    ylab("time (ms)") +
    labs(title = "",
         caption = paste0 ("a. Estimated effects of ",
                           cur_factor,
                           " on h_t in PN pitch accent.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
    )
  png(
    filename =
      paste0("4_pn_output/pn_h_t_re_", cur_factor, "_est.png"),
    width = 4,
    height = 0.5 +
      (length(unique(pn_h_t_mdl.trimmed@frame[[cur_factor]])) - 0.5) * 0.5,
    units = "in",
    res = 300
  )
  print(my_plot)
  dev.off()
  print(my_plot)
}

```

```{r h_t model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

pn_h_t_mdl.tidy <- analyseModel(
  pn_h_t_mdl.trimmed,
  write="4_pn_output/pn_h_t",
  type = "pred",
  y_lab = "time (ms)",
  y_lim = c(40, 300),
  breaks = seq(50, 300, 100),
  plot_rounding = 0,
  hjust = 1,
  panel_prefix = "a. ")

```

```{r do h_t pairwise analysis, fig.height=2.5, fig.width=2.823, results = "asis"}
printTidyPredictions(pn_h_t_mdl.trimmed)

charts <-
  getModelFixedFX(pn_h_t_mdl.trimmed, write="4_pn_output/pn_h_t")
charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.),0)))
```

### Analysis of H F0

```{r run h_f0 models, fig.height=3, fig.width=9}
h_f0_equation = formula(
  h_f0 ~
    acc_phon
  + ana_syls
  + foot_syls
  + wrd_end_syl
  + pn_new_word
  + gender
  + (1 + foot_syls | speaker)
  + (1 | pn_str_syl)
)


pn_h_f0_mdl = lmer(
  h_f0_equation,
  pn_h_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

pn_h_f0_mdl <- optimizeModel(pn_h_f0_mdl, reject_nm = T, verbose = F)

mdl_summary <- summariseLME(run_step = F,
  pn_h_f0_mdl, write = "4_pn_output/pn_h_f0_anova.csv"
  )

adjustP_posthoc("4_pn_output/",
                p.value,
                suffix_id = "_anova")

read_csv("4_pn_output/pn_h_f0_anova.csv") %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(pn_h_f0_mdl)))

```

```{r plot h_f0 estimated effects, fig.height=3.5, fig.width=4}
# plot_model(
#   pn_h_f0_mdl,
#   type = "est",
#   transform = NULL,
#   show.values = T,
#   vline = "Red"
#  )
```

```{r h_f0 model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

pn_h_f0_mdl.tidy <- analyseModel(
  pn_h_f0_mdl,
  write = "4_pn_output/pn_h_f0",
  type = "pred",
  y_lab = "f0 (ST re median)",
  y_lim = c(-1, 4),
  plot_rounding =  1,
  breaks = seq(-1, 4, 1),
  panel_prefix = "b. ")
```

```{r do h_f0 pairwise analysis, fig.height=2.5, fig.width=2.823, results = "asis"}
printTidyPredictions(pn_h_f0_mdl)

charts <- getModelFixedFX(pn_h_f0_mdl, write="4_pn_output/pn_h_f0")
charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.),1)))
```


```{r h binomial plots, fig.height=4.5/2.54, fig.width=15/2.54}
# set terms to plot
terms = c("pn_new_word [T]",
          "gender [M]")

# create _t plot
h_t_binom_plot <- plot_model(pn_h_t_mdl.trimmed,
           type = "est",
           transform = NULL,
           show.values = T,
           vline = "Red",
           terms = terms) +
  geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylab("estimate (ms)") +
    labs(
         caption = paste0 ("a. PN h_t model binomial factor effects.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
      axis.title = element_text(size = 10)
    )

# create _f0 plot
h_f0_binom_plot <- plot_model(pn_h_f0_mdl,
           type = "est",
           transform = NULL,
           show.values = T,
           vline = "Red",
           terms = terms) +
  geom_hline(yintercept = 0, size = 0.5, color = "red") +
    ylab("estimate (ST)") +
    labs(title = "",
         caption = paste0 ("b. PN h_f0 model binomial factor effects.")) +
    theme(
      plot.title = element_blank(),
      plot.caption = element_text(hjust = 0, size = 10),
      plot.caption.position = "plot",
      axis.title = element_text(size = 10)
    )

# join the plots
grid.arrange(h_t_binom_plot, h_f0_binom_plot, ncol = 2)

# save the plots
ggsave("4_pn_output/h_target_binomial.png",
        device = "png",
        arrangeGrob(h_t_binom_plot, h_f0_binom_plot, ncol = 2),
        width = 15.5,
        height = 4.5,
        units = "cm")
```

## Truncation and Compression

Based on results from tonal target analysis, it looks there is an element of truncation in LH contours as a function of foot size, alongside a degree of rightward drift in L targets as foot size increases. However, superficially, it looks like the slope is stable.

As with nuclear data, I am retaining the overall model structure as much as possible BUT will discard all but foot_syls for analysis.

### LH Excursion

```{r run f0_exc models, fig.height=3, fig.width=9}

f0_exc_equation = formula(
  f0_exc ~ 
  + foot_syls
  + (1 | speaker)
  + (1 | ana_syls)
  + (1 | gender)
  + (1 | pn_str_syl)

)


pn_f0_exc_mdl = lmer(
  f0_exc_equation,
  pn_tnc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

pn_f0_exc_mdl <- optimizeModel(pn_f0_exc_mdl, reject_nm = T, verbose = F)

mdl_summary <- summariseLME(pn_f0_exc_mdl, run_step = F, print_summary = F)

# model needs trimming

pn_f0_exc.trimmed <-
  pn_tnc_data %>% filter(abs(scale(resid(pn_f0_exc_mdl))) <= 2.5)
pn_f0_exc_mdl.trimmed = update(pn_f0_exc_mdl, data = pn_f0_exc.trimmed)

# pn_f0_exc_mdl.trimmed <-
#   optimizeModel(pn_f0_exc_mdl.trimmed,
#              reject_nm = T,
#              verbose = F)

mdl_summary <- summariseLME(pn_f0_exc_mdl.trimmed,
             run_step = F,
             write = "4_pn_output/pn_f0_exc_anova.csv")

# pn_f0_exc_mdl.trimmed = optimizeModel(pn_f0_exc_mdl.trimmed, verbose = F)


adjustP_posthoc("4_pn_output/",
                p.value,
                suffix_id = "_anova")

read_csv("4_pn_output/pn_f0_exc_anova.csv",
         show_col_types = F) %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                              getModelFormula(pn_f0_exc_mdl.trimmed)))
```

```{r plot f0 exc estimated effects, fig.height=3.5, fig.width=4}
a <- plot_model(
  pn_f0_exc_mdl.trimmed,
  rm.terms = "gender [M]",
  type = "est",
  transform = NULL,
  show.values = T,
  vline = "Red",
 ) + labs(caption = "a. Estimated effects on f0_exc.") +
  theme(title=element_blank(),
                  plot.caption=element_text(hjust = 0, size = 10),
                  plot.caption.position= "plot") +
  geom_hline(yintercept = 0, size = 0.5, color = "red")
```

```{r f0_exc model summary, fig.height=2.5, fig.width=2.823, , results="asis"}
mdl_summary$omega2

pn_f0_exc_mdl.tidy <- analyseModel(
  pn_f0_exc_mdl.trimmed,
  write = "4_pn_output/pn_f0_exc",
  type = "pred",
  y_lab = "f0 (ST re median)",
  y_lim = c(-1.5, 5),
  plot_rounding =  1
)


```


```{r do f0_exc pairwise analysis, fig.height=2.5, fig.width=2.823, results = "asis"}
printTidyPredictions(pn_f0_exc_mdl.trimmed)

charts <- getModelFixedFX(pn_f0_exc_mdl.trimmed,
                write = "4_pn_output/pn_f0_exc")
charts$slopes %>% tidyStatNumbers()%>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.),1)))
```

### LH Slope

```{r run lh_slope models, fig.height=3, fig.width=9}
lh_slope_equation = formula(
  log_lh_slope ~
  + foot_syls
  + (1 | speaker)
  + (1 | gender)
  + (1 | ana_syls)
  + (1 | pn_str_syl)
  + (1 | wrd_end_syl)
)

pn_lh_slope_mdl = lmer(
  lh_slope_equation,
  pn_tnc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

 pn_lh_slope_mdl<-
   optimizeModel(pn_lh_slope_mdl,
              reject_nm = T,
              verbose = T)

mdl_summary <- summariseLME(pn_lh_slope_mdl,
                  print_summary = F,
                  run_step = T,
                  plot_resids = T)


```
- data needs trimming: residual SD > 4
```{r run lh_slope models trimmed, fig.height=3, fig.width=9}
# model needs trimming
pn_lh_slope.trimmed <-
  pn_tnc_data %>% filter(abs(scale(resid(pn_lh_slope_mdl))) <= 4)

pn_lh_slope_mdl.trimmed = lmer(
  lh_slope_equation,
  pn_lh_slope.trimmed,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

# pn_lh_slope_mdl.trimmed <-
#   optimizeModel(pn_lh_slope_mdl.trimmed,
#              reject_nm = T,
#              verbose = F)

mdl_summary <- summariseLME(pn_lh_slope_mdl.trimmed,
             run_step = F,
             write = "4_pn_output/pn_lh_slope_anova.csv")

adjustP_posthoc("4_pn_output/",
                p.value,
                suffix_id = "_anova")

read_csv("4_pn_output/pn_lh_slope_anova.csv",
         show_col_types = F) %>%
  tidyStatNumbers() %>%
  formattable(caption = paste("ANOVA of model:",
                              getModelFormula(pn_lh_slope_mdl.trimmed)))
```

```{r plot lh slope estimated effects, fig.height=3.5, fig.width=4}
b <- plot_model(
  pn_lh_slope_mdl.trimmed,
  type = "est",
#  rm.terms = "ana_syls [2,3,4]",
  transform = NULL,
  show.values = T
 ) + labs(caption = "b. Estimated effects on log_lh_slope.") +
  theme(title=element_blank(),
                  plot.caption=element_text(hjust = 0, size = 10),
                  plot.caption.position= "plot") +
  geom_hline(yintercept = 0, size = 0.5, color = "red")
```

```{r lh_slope model summary, fig.height=2.5, fig.width=2.823, results="asis"}
mdl_summary$omega2

pn_lh_slope_mdl.tidy <- analyseModel(
  pn_lh_slope_mdl.trimmed,
  write = "4_pn_output/pn_lh_slope",
  type = "pred",
  y_lab = "log(ST/sec)",
  y_lim = c(1.5, 4),
  plot_rounding =  2,
  hjust = 0.9
  )

```

```{r do lh_slope pairwise analysis, fig.height=2.5, fig.width=2.823, results = "asis"}
printTidyPredictions(pn_lh_slope_mdl.trimmed)

charts <- getModelFixedFX(pn_lh_slope_mdl.trimmed,
                write = "4_pn_output/pn_lh_slope")
charts$slopes %>% tidyStatNumbers() %>% 
  mutate(across(
          any_of(c("estimate", "conf.low", "conf.high")),
          ~ round(as.numeric(.),2)))
```


```{r compare effects exc and log_slope effects, fig.height=2.5, fig.width=6}
grid.arrange(a, b, ncol = 2) 
ggsave("4_pn_output/pn_trunc_comp.png",
       device = "png",
       arrangeGrob(a, b, ncol = 2),
       width = 15.7,
       height = 6.5,
       units = "cm")
```

