---
title: "LMM Analysis of PNs"
author: "Antoin Rodgers"
date: "1/19/2020"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
```


Get packages from library.
```{r message=FALSE, warning=FALSE, include=FALSE}
source('../functions/drawResiduals.R', echo=TRUE)
source("0_Initialise_libs.R")
source("0_Get_tbls.R")
```


Create data tbls
```{r message=FALSE, warning=FALSE, include=FALSE}

# Get PN ANAC data
pn_foot <- read_csv("../data/a_corpus.csv") %>%
    filter(
           code %notin% c("F15_A1422_1", "M10_A2422_1", "F5_H0422_2"),
           stim %in% c("A0131", "A0221", "A0321", "H0422"),
           cur_foot == 1,
           substr(stim, 2, 4) == metre_ID
           )%>%
    select(
           speaker, gender, stim,
           init_phon, foot_syls, tot_syls,
           acc_phon, phr_end_t, foot_end_t,
           v_onset_t, l_t, h_t,
           l_f0, h_f0, mean_st, slope_st
           ) %>%
    mutate(
           d_f0 = h_f0 - l_f0,
           d_t = h_t - l_t,
           l_t = l_t - v_onset_t,
           h_t = h_t - v_onset_t,
           SR = round(1000 * tot_syls / phr_end_t, 3),
           stim = factor(stim,levels = unique(stim)),
           acc_phon = factor(acc_phon,
                             levels = c("(*)", "L*", "H*", ">H*", "L*H")
                             ),
           stim = factor(stim, levels = unique(stim))
           ) %>%
    rename(foot_dur = foot_end_t) %>%
    select(-c("phr_end_t", "tot_syls", "v_onset_t")) %>%
    convert(fct(speaker:acc_phon))
```


h_t re acc_phon
```{r fig.height=4, fig.width=12}
temp_mdl = lmer(h_t ~ foot_syls*SR + acc_phon*gender + (1 + foot_syls | speaker),
              data = pn_foot %>% 
                     filter(acc_phon %in% c("H*", ">H*", "L*H")) %>% 
                     mutate(acc_phon = factor(acc_phon,
                                              levels = c("H*", ">H*", "L*H"))
                                              ),
              control = lmerControl(
                                    optimizer = "optimx",
                                    calc.derivs = FALSE,
                                    optCtrl = list(
                                                   method = "nlminb",
                                                   starttests = FALSE,
                                                   kkt = FALSE
                                                   )
                                    ),
              contrasts = list(foot_syls = code_helmert_forward(4))
              )




step(temp_mdl)
#drawResiduals(temp_mdl)
#plot_model(temp_mdl)
#tidy(anova(temp_mdl)) %>%
#  mutate(p.value = round(p.value, 6),
#         statistic = round(statistic, 3)) %>% 
#  rename("F value" = statistic)
summary(temp_mdl)
anova(temp_mdl)
  

```


```{r echo=FALSE, fig.height=4, fig.width=6, message=TRUE, warning=TRUE}
plot_model(ht_anac,
  title = "H time ~ anac + PN + gen + SR + (1 + PN + SR | speaker)",
  show.intercept = FALSE,
  show.values = TRUE,
  vline.color = "red",
  colors = "Black"
)

summary(ht_anac)
anova(ht_anac)
r.squaredGLMM(ht_anac)
```


lmertest::step() was run to see what the most efficient model might be.
This suggesest the same model but without the random slope.
```{r eval=FALSE, fig.height=4, fig.width=6, message=TRUE, warning=TRUE, include=FALSE}

step(ht_anac)
```
Re-running model without random slopes increased r^2 of fixed slightly but reduced the r^2 of the whole model (slightly).
This was not pursued further.
```{r eval=FALSE, fig.height=4, fig.width=6, message=TRUE, warning=TRUE, include=FALSE}
ht_anac <- lmer(h_t ~ ana_syls + gender + SR +
  (1 | speaker),
data = filter(pn_anac, acc_phon == "L*H"),
control = lmerControl(
  optimizer = "optimx",
  calc.derivs = FALSE,
  optCtrl = list(
    method = "nlminb",
    starttests = FALSE,
    kkt = FALSE
  )
)
)
```
```{r eval=FALSE, fig.height=4, fig.width=6, message=TRUE, warning=TRUE, include=FALSE}
plot_model(ht_anac,
  title = "H time ~ anac + PN + gen + SR + (1 + PN + SR | speaker)",
  show.intercept = FALSE,
  show.values = TRUE,
  vline.color = "red",
  colors = "Black"
)

summary(ht_anac)
anova(ht_anac)
r.squaredGLMM(ht_anac)
```


### Testing HT as a function of anacrusis, pitchaccent, and gender * speechrate
```{r fig.height=4, fig.width=6, message=TRUE, warning=TRUE}
ht_anac <- lmer(h_t ~ ana_syls + acc_phon + gender + SR +
  (1 + ana_syls + SR | speaker),
# data = filter(pn_anac, acc_phon == "L*H"),
data = pn_anac,
control = lmerControl(
  optimizer = "optimx",
  calc.derivs = FALSE,
  optCtrl = list(
    method = "nlminb",
    starttests = FALSE,
    kkt = FALSE
  )
)
)
```
```{r echo=FALSE, fig.height=4, fig.width=6, message=TRUE, warning=TRUE}

plot_model(ht_anac,
  title = "H time ~ anac + PN + gen + SR + (1 + SR | speaker)",
  show.intercept = FALSE,
  show.values = TRUE,
  vline.color = "red",
  colors = "Black"
)
summary(ht_anac)
anova(ht_anac)
r.squaredGLMM(ht_anac)
```
