---
title: "LME Analysis of NUCs"
author: "Antoin Rodgers"
date: "18.05.2022"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE)

# Get personal functions and install / load necessary packages.
source("../functions/myFunctions.R")

installMissingPackages(
  c(
    # Include statistical packages.
    "performance",
    "lmerTest",
    "lme4",
    "ggeffects",
    "optimx",
    "MuMIn",
    "dfoptim",

    # Include packages for tidy output.
    "tidyverse",
    "broomExtra",
    "sjPlot",
    "formattable",
    "knitr",
    "janitor",
    
    # Include packages for %in% %notin% syntactic notation
    "mefa4"
  )
)

# Set themes and colour schemes.
theme_set(theme_minimal(base_size = 14))

# Get  data for analysis

source("get_AH_corpora.R")

nuc_data <- nuc %>%
  select(
    speaker,
    gender,
    pre_syls,
    foot_syls,
    nuc_pre_text,
    nuc_new_word,
    fin_phon,
    stim,
    h_t,
    h_f0,
    l_t,
    l_f0,
    e_t,
    e_f0,
    f0_exc,
    lh_dur,
    lh_slope,
    foot_dur,
    speech_rate
  )
rm(pn, nuc, pn_ana, pn_foot, nuc_pre, nuc_foot, pn_lex, corpus)

```

```{r check levels in factors, results="asis", eval=FALSE}

cat("Tibble factors for nuc_data\n")
for (i in colnames(nuc_data))
{
  if (is.factor(nuc_data[[eval(i)]]))
  {
    cat("\n", i, "\tlevels = ", levels(nuc_data[[i]]))
  }
  else
  {
    cat("\n", i)
    }
  
}

nuc_data %>%
  group_by(stim, pre_syls) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = pre_syls, values_from = count, values_fill = 0) %>%
  adorn_totals("row") %>% 
  formattable(caption = "pre_syls stim variety")

nuc_data %>%
  group_by(stim, nuc_new_word) %>%
  summarise(count = n()) %>%
  pivot_wider(names_from = nuc_new_word, values_from = count, values_fill = 0) %>%
  adorn_totals("row") %>% 
  formattable(caption = "nuc_new_word stim variety")
```

## L targets

### Analysis of L timing

```{r set l_t equation}
l_t_equation = formula(
  l_t ~ foot_syls
  + pre_syls
  + fin_phon
  + nuc_new_word
  + gender
  + (1 | speaker)
  + (1 | speech_rate)
 #+ (1 | foot_dur)
  + (1 | nuc_pre_text)
  )

optimizer = "optimx"
```

- intercepts only model

```{r run l_t models, fig.height=3, fig.width=9}
nuc_l_t_mdl = lmer(
  l_t_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

summariseLME(nuc_l_t_mdl,
             run_step = TRUE,
             extra_text = "[nuclear contour]",
             write = "nuc_anovas/nuc_l_t_anova.csv")
```

```{r l_t model summary, fig.height=2.5, fig.width=5, results="asis"}
plot_model(nuc_l_t_mdl, type = "re",
           title = "Random effects of preceding unstressed content")[2]

##tidyPrediction(nuc_l_t_mdl, write = "predictions/nuc_l_t_mdl")

nuc_l_t_mdl.tidy <- analyseModel(nuc_l_t_mdl,
                                write="nuc_phonetic_models/nuc_l_t",
                                type = "pred",
                                factor_matrix = TRUE
                                )
nuc_l_t_mdl.tidy$table

```

```{r do l_t pairwise analysis, fig.height=2.5, fig.width=5}

charts <- getModelFixedFX(
  l_t_equation,
  nuc_data,
  optimizer,
  write="nuc_phonetic_models/nuc_l_t"
)
charts$intercepts
charts$slopes


```

### Analysis of L F0

- Only intercepts only model avoided convergence and singularity issues
- Speech rate caused massive estimation issues

```{r set l_f0 equation}

l_f0_equation = formula(
l_f0 ~ foot_syls
  + pre_syls
  + fin_phon
  + nuc_new_word
  + (1 | speaker)
 #+ (1 | speech_rate)
 #+ (1 | foot_dur)
  + (1 | nuc_pre_text)
  )
optimizer = "optimx"
```

- model needs trimming (|sd| residuals > 3)
```{r run l_f0 models, fig.height=3, fig.width=9}
nuc_l_f0_mdl = lmer(
  l_f0_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

# model needs trimming
nuc_l_f0.trimmed <-
  nuc_data %>% filter(abs(scale(resid(nuc_l_f0_mdl))) <= 3)

nuc_l_f0_mdl.trimmed = lmer(
  l_f0_equation,
  nuc_l_f0.trimmed,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

summariseLME(nuc_l_f0_mdl.trimmed,
             run_step = TRUE,
             extra_text = "[nuclear contour]",
             write = "nuc_anovas/nuc_l_f0_anova.csv")
```

```{r l_f0 model summary, fig.height=2.5, fig.width=5, results="asis"}
##tidyPrediction(nuc_l_f0_mdl.trimmed, write = "predictions/nuc_l_f0_mdl")

nuc_l_f0_mdl.tidy <- analyseModel(nuc_l_f0_mdl.trimmed,
                                write="nuc_phonetic_models/nuc_l_f0",
                                type = "pred"
                                )
nuc_l_f0_mdl.tidy$table

```

```{r do l_f0 pairwise analysis, fig.height=2.5, fig.width=5}

charts <- getModelFixedFX(
  l_f0_equation,
  nuc_l_f0.trimmed,
  optimizer,
  write="nuc_phonetic_models/nuc_l_f0"
)
charts$intercepts
charts$slopes


```

## H targets

### Analysis of H timing

- Only intercepts only model avoided convergence and singularity issues
- Singularity continues, so redundant `nuc_new_word` factor removed as per `step()` output 
```{r set h_t equation}
h_t_equation = formula(
  h_t ~ foot_syls
  + pre_syls
  + fin_phon
  + gender
 #+ nuc_new_word
  + (1 | speaker)
  + (1 | speech_rate)
  + (1 | foot_dur)
  + (1 | nuc_pre_text)
  )

optimizer = "optimx"
```

```{r run h_t models, fig.height=3, fig.width=9}
nuc_h_t_mdl = lmer(
  h_t_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

summariseLME(nuc_h_t_mdl,
             run_step = TRUE,
             extra_text = "[nuclear contour]",
             write = "nuc_anovas/nuc_h_t_anova.csv")
```


```{r h_t model summary, fig.height=2.5, fig.width=5, results="asis"}
##tidyPrediction(nuc_h_t_mdl, write = "predictions/nuc_h_t_mdl")

nuc_h_t_mdl.tidy <- analyseModel(nuc_h_t_mdl,
                                write="nuc_phonetic_models/nuc_h_t",
                                type = "pred",
                                factor_matrix = TRUE
                                )
nuc_h_t_mdl.tidy$table

```

```{r do h_t pairwise analysis, fig.height=2.5, fig.width=5}

charts <- getModelFixedFX(
  h_t_equation,
  nuc_data,
  optimizer,
  write="nuc_phonetic_models/nuc_h_t"
)
charts$intercepts
charts$slopes
```

### Analysis of H F0
```{r set h_f0 equation}
h_f0_equation = formula(
  h_f0 ~ foot_syls
  + pre_syls
  + fin_phon
 #+ nuc_new_word
  + (1 | speaker)
 #+ (1 | speech_rate)
  + (1 | foot_dur)
  + (1 | nuc_pre_text)
  )

optimizer = "optimx"
```

- model needs trimming (residuals +/- 4 s.d.)
```{r run h_f0 models, fig.height=3, fig.width=9}
nuc_h_f0_mdl = lmer(
  h_f0_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

# model needs trimming
nuc_h_f0.trimmed <-
  nuc_data %>% filter(abs(scale(resid(nuc_h_f0_mdl))) <= 4)

nuc_h_f0_mdl.trimmed = lmer(
  h_f0_equation,
  nuc_h_f0.trimmed,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

summariseLME(nuc_h_f0_mdl.trimmed,
             run_step = TRUE,
             write = "nuc_anovas/nuc_h_f0_anova.csv")
```

```{r h_f0 model summary, fig.height=2.5, fig.width=5, results="asis"}
##tidyPrediction(nuc_h_f0_mdl, write = "predictions/nuc_h_f0_mdl")

nuc_h_f0_mdl.tidy <- analyseModel(nuc_h_f0_mdl,
                                write="nuc_phonetic_models/nuc_h_f0",
                                type = "pred"
                                )
nuc_h_f0_mdl.tidy$table

```

```{r do h_f0 pairwise analysis, fig.height=2.5, fig.width=5}

charts <- getModelFixedFX(
  h_f0_equation,
  nuc_data,
  optimizer,
  write="nuc_phonetic_models/nuc_h_f0"
)
charts$intercepts
charts$slopes


```

## E targets

### Analysis of E timing

```{r set e_t equation}
e_t_equation = formula(
  e_t ~ foot_syls
  + pre_syls
  + fin_phon
  + gender
  + nuc_new_word
  + (1 | speaker)
  #+ (1 | speech_rate)
  + (1 | foot_dur)
  + (1 | nuc_pre_text)
  )

optimizer = "optimx"
```

- intercepts only model
- model needs trimming
```{r run e_t models, fig.height=3, fig.width=9}
nuc_e_t_mdl = lmer(
  e_t_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

# model needs trimming
nuc_e_t.trimmed <-
  nuc_data %>% filter(abs(scale(resid(nuc_e_t_mdl))) <= 4)

nuc_e_t_mdl.trimmed = lmer(
  e_t_equation,
  nuc_e_t.trimmed,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

summariseLME(nuc_e_t_mdl.trimmed,
             run_step = TRUE,
             extra_text = "[nuclear contour]",
             write = "nuc_anovas/nuc_e_t_anova.csv")
```

```{r e_t model summary, fig.height=2.5, fig.width=5, results="asis"}
##tidyPrediction(nuc_e_t_mdl.trimmed, write = "predictions/nuc_e_t_mdl")

nuc_e_t_mdl.tidy <- analyseModel(nuc_e_t_mdl.trimmed,
                                write="nuc_phonetic_models/nuc_e_t",
                                type = "pred",
                                factor_matrix = TRUE
                                )
nuc_e_t_mdl.tidy$table

```

```{r do e_t pairwise analysis, fig.height=2.5, fig.width=5}

charts <- getModelFixedFX(
  e_t_equation,
  nuc_e_t.trimmed,
  optimizer,
  write="nuc_phonetic_models/nuc_e_t"
)
charts$intercepts
charts$slopes


```

### Analysis of E F0

- Only intercepts only model avoided convergence and singularity issues
- Speech rate caused massive estimation issues

```{r set e_f0 equation}
e_f0_equation = formula(
  e_f0 ~ foot_syls
  + pre_syls
  + fin_phon
  + (1 | speaker)
  + (1 | nuc_pre_text)
  )
optimizer = "optimx"
```

- model needs trimming (|sd| residuals > 3.5)
```{r run e_f0 models, fig.height=3, fig.width=9}
nuc_e_f0_mdl = lmer(
  e_f0_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

# model needs trimming
nuc_e_f0.trimmed <-
  nuc_data %>% filter(abs(scale(resid(nuc_e_f0_mdl))) <= 3.5)

nuc_e_f0_mdl.trimmed = lmer(
  e_f0_equation,
  nuc_e_f0.trimmed,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

summariseLME(nuc_e_f0_mdl.trimmed,
             run_step = TRUE,
             extra_text = "[nuclear contour]",
             write = "nuc_anovas/nuc_e_f0_anova.csv")
```


```{r e_f0 model summary, fig.height=2.5, fig.width=5, results="asis"}
##tidyPrediction(nuc_e_f0_mdl, write = "predictions/nuc_e_f0_mdl")

nuc_e_f0_mdl.tidy <- analyseModel(nuc_e_f0_mdl,
                                write="nuc_phonetic_models/nuc_e_f0",
                                type = "pred"
                                )
nuc_e_f0_mdl.tidy$table

```

```{r do e_f0 pairwise analysis, fig.height=2.5, fig.width=5}

charts <- getModelFixedFX(
  e_f0_equation,
  nuc_data,
  optimizer,
  write="nuc_phonetic_models/nuc_e_f0"
)
charts$intercepts
charts$slopes
```


## Truncation and Compression Effects

NB: only foot_syls is of interest here, not pre-syls or fin-phon
However, to preserve previous models, they are included as fixed effects,
but they are discarded.

### LH Excursion

```{r set f0_exc equation}

f0_exc_equation = formula(
 f0_exc ~ foot_syls
 + pre_syls
 + fin_phon
 + (1 | speaker)
 + (1 | nuc_pre_text)
 )

optimizer = "optimx"
```

- model needs trimming (|sd| residuals > 3.5)
```{r run f0_exc models, fig.height=3, fig.width=9}
nuc_f0_exc_mdl = lmer(
  f0_exc_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

```

```{r run lh_exc models trimmed, fig.height=3, fig.width=9}
# model needs trimming
nuc_f0_exc.trimmed <-
  nuc_data %>% filter(abs(scale(resid(nuc_f0_exc_mdl))) <= 2.5)

nuc_f0_exc_mdl.trimmed = lmer(
  f0_exc_equation,
  nuc_f0_exc.trimmed,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)
summariseLME(nuc_f0_exc_mdl.trimmed,
             run_step = TRUE,
             extra_text = "[nuclear contour]",
             write = "nuc_anovas/nuc_f0_exc_anova.csv")
```

```{r f0_exc model summary, fig.height=2.5, fig.width=4, results="asis"}

nuc_f0_exc_mdl.tidy <- analyseModel(nuc_f0_exc_mdl.trimmed,
                                write="nuc_phonetic_models/nuc_f0_exc",
                                type = "pred"
                                )
nuc_f0_exc_mdl.tidy$table
##tidyPrediction(nuc_f0_exc_mdl, write = "predictions/nuc_f0_exc_mdl")

```

```{r do f0_exc pairwise analysis, fig.height=2.5, fig.width=5}

charts <- getModelFixedFX(
  f0_exc_equation,
  nuc_f0_exc.trimmed,
  optimizer,
  write="nuc_phonetic_models/nuc_f0_exc",
  ignore_list = c("pre_syls", "fin_phon")
)
charts$intercepts
charts$slopes


```


### LH Slope

Based on results from tonal target analysis, it looksl like compression and truncation may be in effect.
As compression leads to decrease in duration of rise (but not necessarily much descrease in excursion size),
and as slope increases exponentially as duration decreases, slope has been logged here.
```{r set lh_slope equation}
nuc_data <- nuc_data %>% 
    mutate(log_lh_slope = log(lh_slope))

lh_slope_equation = formula(log_lh_slope ~ foot_syls
                            + pre_syls
                            + fin_phon
                            + (1 | speaker)
                            + (1 | nuc_pre_text))
optimizer = "optimx"
```

- model needs trimming (|sd| residuals > 3.5)
```{r run lh_slope models, fig.height=3, fig.width=9}
nuc_lh_slope_mdl = lmer(
  lh_slope_equation,
  nuc_data,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)

```

```{r run lh_slope models trimmed, fig.height=3, fig.width=9}
# model needs trimming
nuc_lh_slope.trimmed <-
  nuc_data %>% filter(abs(scale(resid(nuc_lh_slope_mdl))) <= 3)

nuc_lh_slope_mdl.trimmed = lmer(
  lh_slope_equation,
  nuc_lh_slope.trimmed,
  control = lmerControl(
    optimizer = optimizer,
    calc.derivs = FALSE,
    optCtrl = list(
      method = "nlminb",
      starttests = FALSE,
      kkt = FALSE
    )
  )
)
summariseLME(nuc_lh_slope_mdl.trimmed,
             run_step = TRUE,
             extra_text = "[nuclear contour]",
             write = "nuc_anovas/nuc_lh_slope_anova.csv")
```

```{r lh_slope model summary, fig.height=2.5, fig.width=4, results="asis"}
##tidyPrediction(nuc_lh_slope_mdl, write = "predictions/nuc_lh_slope_mdl")

nuc_lh_slope_mdl.tidy <- analyseModel(nuc_lh_slope_mdl.trimmed,
                                write="nuc_phonetic_models/nuc_lh_slope",
                                type = "pred"
                                )
nuc_lh_slope_mdl.tidy$table

```

```{r do lh_slope pairwise analysis, fig.height=2.5, fig.width=5}

charts <- getModelFixedFX(
  lh_slope_equation,
  nuc_lh_slope.trimmed,
  optimizer,
  write="nuc_phonetic_models/nuc_lh_slope",
  ignore_list = c("pre_syls", "fin_phon")
)
charts$intercepts
charts$slopes


```
