---
title: 'Chapter 07: Analysis of Function: Sentence Modes'
subtitle: '7.6.1.2 mode-and-phonology Models'
author: "Antoin Rodgers"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

## 7.6.1.2 mode-and-phonology Models

```{r setup mode phon}
#| results="hide",
#| message=F,
#| warning=F

knitr::opts_chunk$set(warning = F, message = F, tidy = "styler")

## Load Functions.
source("../../2_R_Functions/myFunctions.R") 

## Load packages.
installMissingPackages(
  c(
    ## Include statistical packages.
    "performance",
    "lmerTest",
    "lme4",
    "optimx",
    "MuMIn",
    ## Include packages for tidy output.
    "tidyverse",
    "broomExtra",
    "sjPlot",
    "formattable",
    "knitr",
    "RColorBrewer",
    ## Include packages for %in% %notin% syntactic notation
    "mefa4"
  )
)

y_lim_t = c(0, 360)
y_lim_f0 = c(-4.1, 10.5)

```


```{r load corpus l_t_mode_phon}
l_star_h_types = c("L*H", "^[L*]H", "L*^[H]", "^[L*H]")

## Load Corpus
m_corpus <- get_m_corpus("../../4_data/m_corpus.csv")  %>%
  ## Select only columns for analysis
  select(
    speaker,
    gender,
    mode,
    prompt,
    acc_phon,
    fin_phon,
    l_t,
    l_f0,
    h_t,
    h_f0
    ) %>% 
  filter(acc_phon %in% l_star_h_types)  %>% 
  mutate(acc_phon = factor(acc_phon,
                           levels = l_star_h_types)
         )
```

### A. Low Targets

#### A.11 l_t model {.tabset}

- Following intercepts only model for parity with other time parameter models.
- model needs trimmed (residual SD > 3.25)

```{r run initial model l_t_mode_phon}
#| results="hide"

## run model
l_t_mode_phon_mdl <-
  lmer(l_t ~
         mode
       + acc_phon
       + gender
       + (1 | speaker)
       + (1 | prompt)
       + (1 | fin_phon),
       data = m_corpus,
       control = lmerControl(
         optimizer = "optimx",
         calc.derivs = F,
         optCtrl = list(
           method = "nlminb",
           starttests = F,
           kkt = F
      )
    )
  )

l_t_mode_phon_mdl <- optimizeModel(l_t_mode_phon_mdl)

## Model needs trimmed.
l_t_mode_phon_mdl <- update(l_t_mode_phon_mdl,
                    data = m_corpus %>% filter(abs(scale(resid(
                      l_t_mode_phon_mdl
                    ))) <= 3.25))

l_t_mode_phon_mdl <- optimizeModel(l_t_mode_phon_mdl)                    


```
 
```{r summarise l_t_mode_phon model}
write = "output/l_t_mode_phon"

l_t_mode_phon.summary <- summariseLME(l_t_mode_phon_mdl,
                            run_step = F,
                            write = paste0(write, "_anova.csv"),
                            print_summary = F,
                            plot_resids = F)

``` 

##### Predictions

```{r l_t_mode_phon predictions}
#| fig.height=2.5,
#| fig.width=3.01,
#| results="asis"

l_t_mode_phon.tidy <- analyseModel(
  l_t_mode_phon_mdl,
  write="output/l_t_mode_phon",
  print_r2 = F,
  short_caption = T,
  y_lim = y_lim_t,
  y_lab = "t (ms)",
  caption_suffix = " (mode-and-phonology model)",
  plot_rounding = 0
)

printTidyPredictions(l_t_mode_phon_mdl,
                     digits = 0,
                     caption_suffix = " (mode-and-phonology model)")
```

##### Effect Sizes

```{r l_t_mode_phon effect sizes}
l_t_mode_phon.tidy$r2
l_t_mode_phon.summary$omega2
```

##### Pairwise comparisons

```{r l_t_mode_phon model Pairwise Comparisons}
#| fig.align="left"

charts <- getModelFixedFX(
  l_t_mode_phon_mdl,
  write = write
)

charts$slopes %>% tidyStatNumbers(digits = 1)

```

##### ANOVA

```{r print l_t_mode_phon model summary text info}
#| fig.align="left"

adjustP_posthoc("output/",
                p.value,
                suffix_id = "_anova")


read_csv(paste0(write, "_anova.csv")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(l_t_mode_phon_mdl)))
```

##### Details

```{r print l_t_mode_phon model summary}
#| fig.height=3,
#| fig.width=9

l_t_mode_phon.summary <- summariseLME(l_t_mode_phon_mdl,
                            run_step = F)
```


#### A.2. l_f0 model {.tabset}

- Using intercepts only model for parity with mode-and-phonology analysis. 
- model needs trimmed (residual SD > 3)

```{r run initial model l_f0_mode_phon}
#| results="hide"

## model works with intercepts_phon model
## as per mode and acc_phon model

## run model
l_f0_mode_phon_mdl <-
  lmer(l_f0 ~
         mode
       + acc_phon
       + gender
       + (1 | speaker)
       + (1 | fin_phon)
       #+ (1 | prompt)
       ,
       data = m_corpus,
       control = lmerControl(
         optimizer = "optimx",
         calc.derivs = F,
         optCtrl = list(
           method = "nlminb",
           starttests = F,
           kkt = F
      )
    )
  )

l_f0_mode_phon_mdl <- optimizeModel(l_f0_mode_phon_mdl)

## Model needs trimmed.
l_f0_mode_phon_mdl <- update(l_f0_mode_phon_mdl,
                    data = m_corpus %>% filter(abs(scale(resid(
                      l_f0_mode_phon_mdl
                    ))) <= 3))

l_f0_mode_phon_mdl <- optimizeModel(l_f0_mode_phon_mdl)                    


```
 
```{r summarise l_f0_mode_phon model}
write = "output/l_f0_mode_phon"

l_f0_mode_phon.summary <- summariseLME(l_f0_mode_phon_mdl,
                            run_step = F,
                            write = paste0(write, "_anova.csv"),
                            print_summary = F,
                            plot_resids = F)

``` 

##### Predictions

```{r l_f0_mode_phon predictions}
#| fig.height=2.5,
#| fig.width=3.01,
#| results="asis"

l_f0_mode_phon.tidy <- analyseModel(
  l_f0_mode_phon_mdl,
  write = write,
  print_r2 = F,
  short_caption = T,
  y_lim = y_lim_f0,
  y_lab = "ST re speaker median",
  caption_suffix = " (mode-and-phonology model)"
)

printTidyPredictions(l_f0_mode_phon_mdl,
                     digits = 1,
                     caption_suffix = " (mode-and-phonology model)")
```

##### Effect Sizes

```{r l_f0_mode_phon effect sizes}
l_f0_mode_phon.tidy$r2
l_f0_mode_phon.summary$omega2
```

##### Pairwise comparisons

```{r l_f0_mode_phon model Pairwise Comparisons}
#| fig.align="left"

charts <- getModelFixedFX(
  l_f0_mode_phon_mdl,
  write = write
)

charts$slopes %>% tidyStatNumbers(digits = 1)

```

##### ANOVA

```{r print l_f0_mode_phon model summary text info}
#| fig.align="left"

adjustP_posthoc("output/",
                p.value,
                suffix_id = "_anova")


read_csv(paste0(write, "_anova.csv")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(l_f0_mode_phon_mdl)))
```

##### Details

```{r print l_f0_mode_phon model summary}
#| fig.height=3,
#| fig.width=9

l_f0_mode_phon.summary <- summariseLME(l_f0_mode_phon_mdl,
                            run_step = F)
```

### B. High Targets

#### B.2. h_t model {.tabset}

- Following intercepts only model as in other time parameter models.
- model needs trimmed (residual SD > 3.5)

```{r run initial model h_t_mode_phon}
#| results="hide"

## model works with intercepts_phon model
## as per mode and acc_phon model

## run model
h_t_mode_phon_mdl <-
  lmer(h_t ~
         mode
       + acc_phon
       + gender
       + (1 | speaker)
       + (1 | fin_phon)
       + (1 | prompt),
       data = m_corpus,
       control = lmerControl(
         optimizer = "optimx",
         calc.derivs = F,
         optCtrl = list(
           method = "nlminb",
           starttests = F,
           kkt = F
      )
    )
  )

h_t_mode_phon_mdl <- optimizeModel(h_t_mode_phon_mdl)

## Model needs trimmed.
h_t_mode_phon_mdl <- update(h_t_mode_phon_mdl,
                    data = m_corpus %>% filter(abs(scale(resid(
                      h_t_mode_phon_mdl
                    ))) <= 3.5))

h_t_mode_phon_mdl <- optimizeModel(h_t_mode_phon_mdl)                    


```
 
```{r summarise h_t_mode_phon model}
write = "output/h_t_mode_phon"

h_t_mode_phon.summary <- summariseLME(h_t_mode_phon_mdl,
                            run_step = F,
                            write = paste0(write, "_anova.csv"),
                            print_summary = F,
                            plot_resids = F)

``` 

##### Predictions

```{r h_t_mode_phon predictions}
#| fig.height=2.5,
#| fig.width=3.01,
#| results="asis"

h_t_mode_phon.tidy <- analyseModel(
  h_t_mode_phon_mdl,
  write="output/h_t_mode_phon",
  print_r2 = F,
  short_caption = T,
  y_lim = y_lim_t,
  y_lab = "t (ms)",
  caption_suffix = " (mode-and-phonology model)",
  plot_rounding = 0
)

printTidyPredictions(h_t_mode_phon_mdl,
                     digits = 0,
                     caption_suffix = " (mode-and-phonology model)")
```

##### Effect Sizes

```{r h_t_mode_phon effect sizes}
h_t_mode_phon.tidy$r2
h_t_mode_phon.summary$omega2
```

##### Pairwise comparisons

```{r h_t_mode_phon model Pairwise Comparisons}
#| fig.align="left"

charts <- getModelFixedFX(
  h_t_mode_phon_mdl,
  write = write
)

charts$slopes %>% tidyStatNumbers(digits = 1)

```

##### ANOVA

```{r print h_t_mode_phon model summary text info}
#| fig.align="left"

adjustP_posthoc("output/",
                p.value,
                suffix_id = "_anova")


read_csv(paste0(write, "_anova.csv")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(h_t_mode_phon_mdl)))
```

##### Details

```{r print h_t_mode_phon model summary}
#| fig.height=3,
#| fig.width=9

h_t_mode_phon.summary <- summariseLME(h_t_mode_phon_mdl,
                            run_step = F)
```

#### B.2. h_f0 model {.tabset}

- Using intercepts only model for parity with mode-and-phonology analysis. 
- (1 | prompt) works as random intercept here, but, again, retaining same model structure for parity across f0 models

```{r run initial model h_f0_mode_phon}
#| results="hide"

## model works with intercepts_phon model
## as per mode and acc_phon model

## run model
h_f0_mode_phon_mdl <-
  lmer(h_f0 ~
         mode
       + acc_phon
       + gender
       + (1 | speaker)
       + (1 | fin_phon)
       ,
       data = m_corpus,
       control = lmerControl(
         optimizer = "optimx",
         calc.derivs = F,
         optCtrl = list(
           method = "nlminb",
           starttests = F,
           kkt = F
      )
    )
  )

h_f0_mode_phon_mdl <- optimizeModel(h_f0_mode_phon_mdl)


```
 
```{r summarise h_f0_mode_phon model}
write = "output/h_f0_mode_phon"

h_f0_mode_phon.summary <- summariseLME(h_f0_mode_phon_mdl,
                            run_step = F,
                            write = paste0(write, "_anova.csv"),
                            print_summary = F,
                            plot_resids = F)

``` 

##### Predictions

```{r h_f0_mode_phon predictions}
#| fig.height=2.5,
#| fig.width=3.01,
#| results="asis"

h_f0_mode_phon.tidy <- analyseModel(
  h_f0_mode_phon_mdl,
  write = write,
  print_r2 = F,
  short_caption = T,
  y_lim = y_lim_f0,
  y_lab = "ST re speaker median",
  caption_suffix = " (mode-and-phonology model)"
)

printTidyPredictions(h_f0_mode_phon_mdl,
                     digits = 1,
                     caption_suffix = " (mode-and-phonology model)")
```

##### Effect Sizes

```{r h_f0_mode_phon effect sizes}
h_f0_mode_phon.tidy$r2
h_f0_mode_phon.summary$omega2
```

##### Pairwise comparisons

```{r h_f0_mode_phon model Pairwise Comparisons}
#| fig.align="left"

charts <- getModelFixedFX(
  h_f0_mode_phon_mdl,
  write = write
)

charts$slopes %>% tidyStatNumbers(digits = 1)

```

##### ANOVA

```{r print h_f0_mode_phon model summary text info}
#| fig.align="left"

adjustP_posthoc("output/",
                p.value,
                suffix_id = "_anova")


read_csv(paste0(write, "_anova.csv")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(h_f0_mode_phon_mdl)))
```

##### Details

```{r print h_f0_mode_phon model summary}
#| fig.height=3,
#| fig.width=9

h_f0_mode_phon.summary <- summariseLME(h_f0_mode_phon_mdl,
                            run_step = F)
```
