---
title: 'Chapter 07: Analysis of Function: Sentence Modes'
subtitle: '7.6.3 Mode and Global Phonetic Parameters'
author: "Antoin Rodgers"
output:
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r setup global phonetic, results="hide", message=F, warning=F}
knitr::opts_chunk$set(warning = F, message = F, tidy = "styler")

# Load Functions
source("../../2_R_Functions/myFunctions.R") 

# Load packages.
installMissingPackages(
  c(
    # Include statistical packages.
    "performance",
    "lmerTest",
    "lme4",
    "optimx",
    "MuMIn",
    # Include packages for tidy output.
    "tidyverse",
    "stringr",
    "broomExtra",
    "janitor",
    "sjPlot",
    "formattable",
    "knitr",
    "RColorBrewer",
    # Include packages for %in% %notin% syntactic notation
    "mefa4"
  )
)

# Load Corpus
m_corpus <- get_m_corpus("../../4_data/m_corpus.csv")  %>%
  # Keep only factors needed for analysis
  select(speaker,
         gender,
         mode,
         utt_mean_f0,
         utt_slope,
         phr_phon,
         acc_phon,
         fin_phon,
         prompt) %>%
  mutate(
    # h_start = phrases beginning with a high boundary or H* in first foot.
    h_start = if_else(!is.na((str_extract(phr_phon, "%H|\\^\\[%H|% H\\*"))),
                           "H start",
                           "no H start"),
    h_start = factor(h_start, levels = c("H start", "no H start")),
    # Remove register tier identifers from fin_phon. (See Chapter 7.5.2.2.) 
    fin_phon = factor(str_replace_all(fin_phon, "\\^|\\[|\\]", ""),
                           levels = c("%", "L%"))
         )

```

# 7.6.3 Global Phonetic Parameters

## A. Mode-Only Models

### A.1. Utterance mean _f_~0~ {.tabset}

```{r run utt_mean_f0_mode model, results="hide"}
utt_mean_f0_mode_mdl = lmer(
  utt_mean_f0 ~
    mode
  + gender
  + (1 + mode | speaker)
  + (1 | prompt),
  data = m_corpus,
  control = lmerControl(
    optimizer = "optimx",
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

utt_mean_f0_mode_mdl <- optimizeModel(utt_mean_f0_mode_mdl)

# model needs trimming
utt_mean_f0_mode_mdl = update(
  utt_mean_f0_mode_mdl,
  data = m_corpus %>% filter(abs(scale(resid(utt_mean_f0_mode_mdl))) <= 2.9))

utt_mean_f0_mode_mdl <- optimizeModel(utt_mean_f0_mode_mdl)
```

```{r summarise utt_mean_f0_mode model}
write = "output/utt_f0_mode_only"

utt_mean_f0_mode.summary <- summariseLME(utt_mean_f0_mode_mdl,
                            run_step = F,
                            write = paste0(write, "_anova.csv"),
                            print_summary = F,
                            plot_resids = F)

```

#### predictions

```{r utt_mean_f0_mode model predictions, fig.height=2.5, fig.width=3.01, results="asis"}
utt_mean_f0_mode.tidy <- analyseModel(
  utt_mean_f0_mode_mdl,
  write="output/utt_f0_mode_only",
  print_r2 = F,
  short_caption = T,
  y_lim = c(-2, 3),
  y_lab = "ST re speaker median",
  caption_suffix = " (mode only model)"
)
# utt_mean_f0.tidy$table %>% tidyNumbers(digits = 1) %>% tidyStatNumbers()
printTidyPredictions(utt_mean_f0_mode_mdl,
                     digits = 1,
                     caption_suffix = " (mode only model)")
```

#### Effect Sizes

```{r utt_mean_f0_mode effect sizes}
utt_mean_f0_mode.tidy$r2
utt_mean_f0_mode.summary$omega2
```

#### Pairwise comparisons

```{r utt_mean_f0_mode model Pairwise Comparisons, fig.align="left"}
# Get number of levels in target treatment variable.

charts <- getModelFixedFX(
  utt_mean_f0_mode_mdl,
  write="output/utt_f0_mode_only"
)

#charts$intercepts %>% tidyStatNumbers(digits = 1)
charts$slopes %>% tidyStatNumbers(digits = 1)

```

#### ANOVA

```{r print utt_mean_f0_mode model summary text info, fig.align="left"}
adjustP_posthoc("output/",
                p.value,
                suffix_id = "_anova")


read_csv(paste0(write, "_anova.csv")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(utt_mean_f0_mode_mdl)))
```

#### Details

```{r print utt_mean_f0_mode model model summary, fig.height=3, fig.width=9}
utt_mean_f0_mode.summary <- summariseLME(utt_mean_f0_mode_mdl,
                            run_step = F)
```

#### Random Effects

```{r utt_mean_f0_mode model random effect 1, fig.height=4, fig.width=6.1}
plot_model(utt_mean_f0_mode_mdl,
           type = "re", vline.color = "red", show.values = T)[[1]] +
  labs(caption = "A. Random effect: (1 + speaker | mode)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot"
  )
```

```{r utt_mean_f0_mode model random effect 3, fig.height=2, fig.width=3.05}
plot_model(utt_mean_f0_mode_mdl,
           type = "re", vline.color = "red", show.values = T)[[2]] +
  labs(caption = "B. Random effect: (1 | prompt)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot")
```

### A.2 Utterance slope _f_~0~(t) {.tabset}

```{r run utt_slope_mode model, results="hide"}
utt_slope_mode_mdl = lmer(
  utt_slope ~
    mode
  + gender
  + (1 + mode | speaker)
  + (1 | prompt),
  data = m_corpus,
  control = lmerControl(
    optimizer = "optimx",
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

utt_slope_mode_mdl <- optimizeModel(utt_slope_mode_mdl)

# model needs trimming
utt_slope_mode_mdl = update(
  utt_slope_mode_mdl,
  data = m_corpus %>% filter(abs(scale(resid(utt_slope_mode_mdl))) <= 2.5))

utt_slope_mode_mdl <- optimizeModel(utt_slope_mode_mdl)
```

```{r summarise utt_slope_mode model}
write = "output/utt_slope_mode_only"

utt_slope_mode.summary <- summariseLME(utt_slope_mode_mdl,
                            run_step = F,
                            write = paste0(write, "_anova.csv"),
                            print_summary = F,
                            plot_resids = F)

```

#### predictions

```{r utt_slope_mode model predictions, fig.height=2.5, fig.width=3.01, results="asis"}
utt_slope_mode.tidy <- analyseModel(
  utt_slope_mode_mdl,
  write="output/utt_slope_mode_only",
  print_r2 = F,
  short_caption = T,
  y_lim = c(-10, 8),
  y_lab = "ST/s",
  caption_suffix = " (mode only model)"
)
# utt_mean_f0.tidy$table %>% tidyNumbers(digits = 1) %>% tidyStatNumbers()
printTidyPredictions(utt_slope_mode_mdl,
                     digits = 1,
                     caption_suffix = " (mode only model)")
```

#### Effect Sizes

```{r utt_slope_mode effect sizes}
utt_slope_mode.tidy$r2
utt_slope_mode.summary$omega2
```

#### Pairwise comparisons

```{r utt_slope_mode model Pairwise Comparisons, fig.align="left"}
# Get number of levels in target treatment variable.

charts <- getModelFixedFX(
  utt_slope_mode_mdl,
  write="output/utt_slope_mode_only"
)

#charts$intercepts %>% tidyStatNumbers(digits = 1)
charts$slopes %>% tidyStatNumbers(digits = 1)

```

#### ANOVA

```{r print utt_slope_mode model summary text info, fig.align="left"}
adjustP_posthoc("output/",
                p.value,
                suffix_id = "_anova")


read_csv(paste0(write, "_anova.csv")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(utt_slope_mode_mdl)))
```

#### Details

```{r print utt_slope_mode model model summary, fig.height=3, fig.width=9}
utt_slope_mode.summary <- summariseLME(utt_slope_mode_mdl,
                            run_step = F)
```

#### Random Effects

```{r utt_slope_mode model random effect 1, fig.height=4, fig.width=6.1}
plot_model(utt_slope_mode_mdl,
           type = "re", vline.color = "red", show.values = T)[[1]] +
  labs(caption = "A. Random effect: (1 + speaker | mode)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot"
  )
```

```{r utt_slope_mode model random effect 3, fig.height=2, fig.width=3.05}
plot_model(utt_slope_mode_mdl,
           type = "re", vline.color = "red", show.values = T)[[2]] +
  labs(caption = "B. Random effect: (1 | prompt)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot")
```

## B Mode-and-Phonology Models

### B.1. Utterance mean _f_~0~ {.tabset}
```{r run model, results="hide"}
utt_mean_f0_full_mdl = lmer(
  utt_mean_f0  ~ mode + gender
  + (1 + mode | speaker)
  + (1 | prompt)
  + (1 | h_start)
  + (1 | acc_phon)
  + (1 | fin_phon),
  data = m_corpus,
  control = lmerControl(
    optimizer = "optimx",
    calc.derivs = F,
    optCtrl = list(
      method = "nlminb",
      starttests = F,
      kkt = F
    )
  )
)

utt_mean_f0_full_mdl <- optimizeModel(utt_mean_f0_full_mdl)
```

```{r summarise utt_mean_f0_full model}
write = "output/utt_f0_full_phon"

utt_mean_f0_full.summary <- summariseLME(utt_mean_f0_full_mdl,
                            run_step = F,
                            write = paste0(write, "_anova.csv"),
                            print_summary = F,
                            plot_resids = F)

```

#### predictions

```{r utt_mean_f0_full model predictions, fig.height=2.5, fig.width=3.01, results="asis"}
utt_mean_f0_full.tidy <- analyseModel(
  utt_mean_f0_full_mdl,
  write="output/utt_f0_full_phon",
  print_r2 = F,
  short_caption = T,
  y_lim = c(-2, 3),
  y_lab = "ST re speaker median",
  caption_suffix = " (mode + phon model)"
)
# utt_mean_f0_full.tidy$table %>% tidyNumbers(digits = 1) %>% tidyStatNumbers()
printTidyPredictions(utt_mean_f0_full_mdl,
                     digits = 1,
                     caption_suffix = " (mode + phonology model)")
```

#### Effect Sizes

```{r utt_mean_f0_full effect sizes}
utt_mean_f0_full.tidy$r2
utt_mean_f0_full.summary$omega2
```

#### Pairwise comparisons

```{r utt_mean_f0_full model Pairwise Comparisons, fig.align="left"}
# Get number of levels in target treatment variable.

charts <- getModelFixedFX(
  utt_mean_f0_full_mdl,
  write="output/utt_f0_full_phon"
)

#charts$intercepts %>% tidyStatNumbers(digits = 1)
charts$slopes %>% tidyStatNumbers(digits = 1)

```

#### ANOVA

```{r print utt_mean_f0_full model summary text info, fig.align="left"}
adjustP_posthoc("output/",
                p.value,
                suffix_id = "_anova")


read_csv(paste0(write, "_anova.csv")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(utt_mean_f0_full_mdl)))
```

#### Details

```{r print utt_mean_f0_full model model summary, fig.height=3, fig.width=9}
utt_mean_f0_full.summary <- summariseLME(utt_mean_f0_full_mdl,
                            run_step = F)
```

#### Random Effects

```{r utt_mean_f0_full model random effect 1, fig.height=4, fig.width=6.1}
plot_model(utt_mean_f0_full_mdl,
           type = "re", vline.color = "red", show.values = T)[[1]] +
  labs(caption = "A. Random effect: (1 + speaker | mode)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot"
  )
```
```{r utt_mean_f0_full model random effect 2, fig.height=4, fig.width=3.05}
plot_model(utt_mean_f0_full_mdl,
           type = "re", vline.color = "red", show.values = T)[[2]] +
  labs(caption = "B. Random effect: (1 | acc_phon)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot")
```
```{r utt_mean_f0_full model random effect 3, fig.height=2, fig.width=3.05}
plot_model(utt_mean_f0_full_mdl,
           type = "re", vline.color = "red", show.values = T)[[3]] +
  labs(caption = "C. Random effect: (1 | prompt)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot")
plot_model(utt_mean_f0_full_mdl,
           type = "re", vline.color = "red", show.values = T)[[4]] +
  labs(caption = "D. Random effect: (1 | fin_phon)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot")
plot_model(utt_mean_f0_full_mdl,
           type = "re", vline.color = "red", show.values = T)[[5]] +
  labs(caption = "E. Random effect: (1 | h_start)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot")
```

### B.2. Utterance slope _f_~0~(t) {.tabset}

```{r run utt_slope_full model, results="hide"}
utt_slope_full_mdl = lmer(
    utt_slope ~ mode
    + gender
    + (1 + mode | speaker)
    + (1 | prompt)
    + (1 | h_start)
    + (1 | acc_phon)
    + (1 | fin_phon),
    data = m_corpus,
    control = lmerControl(
      optimizer = "optimx",
      calc.derivs = F,
      optCtrl = list(
        method = "nlminb",
        starttests = F,
        kkt = F
        )
      )
    )

utt_slope_full_mdl <- optimizeModel(utt_slope_full_mdl, verbose = F)

utt_slope_full_mdl  <- update(
  utt_slope_full_mdl,
  data = m_corpus %>% filter(abs(scale(resid(utt_slope_full_mdl))) <= 3)
)

utt_slope_full_mdl <- optimizeModel(utt_slope_full_mdl, verbose = F)
```

```{r summarise utt_slope_full model, include=F}
write = "output/utt_slope_full_phon"

utt_slope_full.summary <- summariseLME(utt_slope_full_mdl,
                            run_step = F,
                            write = paste0(write, "_anova.csv"),
                            print_summary = F,
                            plot_resids = F)

```

#### predictions

```{r utt_slope_full model predictions, fig.height=2.5, fig.width=3.01, results="asis"}
utt_slope_full.tidy <- analyseModel(
  utt_slope_full_mdl,
  write="output/utt_slope_full_phon",
  plot_rounding = 2,
  print_r2 = F,
  short_caption = T,
  y_lim = c(-10, 8),
  y_lab = "ST/s",
  caption_suffix = " (mode + phon model)")

# utt_slope_full.tidy$table %>% tidyNumbers(digits = 1) %>% tidyStatNumbers()
printTidyPredictions(utt_slope_full_mdl,
                     digits = 2,
                     caption_suffix = " (mode + phonology model)")
```

#### Effect Sizes

```{r utt_slope_full model effect sizea}
utt_slope_full.tidy$r2
utt_slope_full.summary$omega2
```

#### Pairwise comparisons

```{r utt_slope_full model Pairwise Comparisons, fig.align="left"}
# Get number of levels in target treatment variable.

charts <- getModelFixedFX(
  utt_slope_full_mdl,
  write="output/utt_slope_full_phon"
)

#charts$intercepts %>% tidyStatNumbers(digits = 2)
charts$slopes %>% tidyStatNumbers(digits = 2)

```

#### ANOVA

```{r print sutt_slope_full summary text info, fig.align="left"}
adjustP_posthoc("output/",
                p.value,
                suffix_id = "_anova")


read_csv(paste0(write, "_anova.csv")) %>%
  tidyStatNumbers() %>%
formattable(caption = paste("ANOVA of model:",
                            getModelFormula(utt_slope_full_mdl)))
```

#### Details

```{r print utt_slope_full model summary, fig.height=3, fig.width=9}
mdl_summary <- summariseLME(utt_slope_full_mdl,
                            run_step = F)
```

#### Random Effects

```{r utt_slope_full model random effect 1, fig.height=4, fig.width=6.1}
plot_model(utt_slope_full_mdl,
           type = "re", vline.color = "red", show.values = T)[[1]] +
  labs(caption = "A. Random effect: (1 + speaker | mode)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot"
  )
```
```{r utt_slope_full model random effect 2, fig.height=4, fig.width=3.05}
plot_model(utt_slope_full_mdl,
           type = "re", vline.color = "red", show.values = T)[[2]] +
  labs(caption = "B. Random effect: (1 | acc_phon)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot")
```
```{r utt_slope_full model random effect 3, fig.height=2, fig.width=3.05}
plot_model(utt_slope_full_mdl,
           type = "re", vline.color = "red", show.values = T)[[3]] +
  labs(caption = "C. Random effect: (1 | prompt)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot")
plot_model(utt_slope_full_mdl,
           type = "re", vline.color = "red", show.values = T)[[4]] +
  labs(caption = "D. Random effect: (1 | fin_phon)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot")
plot_model(utt_slope_full_mdl,
           type = "re", vline.color = "red", show.values = T)[[5]] +
  labs(caption = "E. Random effect: (1 | h_start)") +
  theme(axis.title.y = element_blank(),
        plot.title = element_blank(),
        axis.title.x = element_blank(),
        plot.caption = element_text(hjust = 0, size = 10),
        plot.caption.position = "plot")
```

